---
title: "Matching EEZs to UN country names"
author: "Camilla Novaglio"
date: "2024-02-26"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Loading libraries adn defining forlders 
  
```{r, message=FALSE, warning=FALSE}
rm(list = ls())

library(tidyverse)

# Base folder for project (both inputs and outputs)
base_folder <- "/rd/gem/private/users/camillan/FAO_Report/Merged_FAO_EEZ/supporting_files/"
```

# EEZ data
Columns:
`Ocean` = FAO_fishing areas
`ID_mrgd` = key linked to each EEZ lat and lon
`nm_mrgd` = EEZ name 
`continent_map` = continent - comes from previous iteration of the EEZ & FAO data (using Stefania and Emmanuel' datasets). This combines continent from the original EEZ file with decisions taken by considering the FAO data and by looking at maps (Fig.2). 
`fa_ffcl` = FAO official country names - comes from previous iteration of the EEZ & FAO data (using Stefania and Emmanuel' datasets). Though the list of EEZ is the one from the original EEZ mask.   

```{r}
eez <- file.path(base_folder, 
                 "FAO_MajorAreas_200NM_cont_keys.csv") |> 
  read_csv()
```

# FAO data from https://www.fao.org/nocs/en (all countries, areas and territories) 
Columns:
`FULL NAME` = FAO official name for countries, territories and areas
`SHORT NAME` = FAO short name 
`ENGLISH_LIST NAME` = FAO short name option 2

```{r}
fao <- file.path(base_folder, "nocsDataExport_20240501-114742_English.csv") |> 
  #Select relevant columns
  read_csv(col_select = c(`FULL NAME`, `SHORT NAME`, `ENGLISH_LIST NAME`)) |> 
  #Update column names to something more descriptive
  rename(O_name_En_Website =`FULL NAME`, 
         Short_name_En_Website =`SHORT NAME`, 
         Short_name_En_Website2 =`ENGLISH_LIST NAME`) |> 
  # to join with Emmanuel and Stefania information
  mutate(O_Name_En = O_name_En_Website)

#Check result
head(fao)
```

# FAO data from Emmanuel Blondel 
Columns:
`CONTCD` = continent
`ISO_3` = ISO code for countries 
`O_Name_En` = official FAO name 
`ISO3CD` = ISO code for countries and territories

```{r}
fao_ema <- file.path(base_folder, 
                     "countries_un_fromEmmanuelShapeFiles_cleaned.csv") |> 
  read.csv() |>
  #Keep unique records only
  unique()
```

# Create a ISO_3 to Country name dataset

```{r}
# example: 
fao_ema |>
  # list of country code
  select(c(ISO_3, O_Name_En)) |> 
  unique() |> 
  # need to isolate GBR = United Kingdom
  filter(ISO_3 == "GBR")

# 1 list of country (excluding territories) code

key1 <- fao_ema |>
  select(ISO_3) |> 
  unique()

# 2 list of country and territory code and of country and territory names 

key2 <- fao_ema |>
  select(c(ISO3CD, O_Name_En)) |> 
  rename(ISO_3 = ISO3CD, Country = O_Name_En) |>
  unique()

filter(key2, ISO_3 == "GBR")
filter(key2, Country == "Bermuda")

# 3 left join to keep only list of country code as per 1 and add country 
#names as per 2
key3 <- key1 |>
  left_join(key2)

head(key3)
filter(key3, ISO_3 == "GBR")
filter(key3, Country == "Bermuda")
```

# Check data merging and official FAO names duplicates
Consider continent and official English name and add key 3 (country information) to Emmanuel's dataset. 

```{r}
fao2 <- fao_ema |> 
  select(CONTCD, O_Name_En, ISO_3) |>
  right_join(key3)
filter(fao2, ISO_3 == "GBR")

# Check for duplicates
duplicates <- fao2 |> 
  group_by(O_Name_En) |> 
  filter(n() > 1)

duplicates <- unique(duplicates$O_Name_En)
duplicates <- fao_ema |> 
  filter(O_Name_En %in% duplicates)

# check one by one - all checked and fixed below
duplicates <- split(duplicates, duplicates$O_Name_En)
length(duplicates)
# no official name 
# only coming up when using read.csv() instead of read_csv()
duplicates[1] 
# Au Ashmore - not in eez data 
duplicates[2] 
# Saba etc 
duplicates[3] 
# Clipperton 
duplicates[4] 
```

# ROMNAM & MAPLAB are identifying columns, keep them in the final dataset 

```{r}
fao_ema <- fao_ema |> 
  # keep also ISO3CD for merging with Stefania's data below 
  select(CONTCD, ROMNAM, MAPLAB, O_Name_En, ISO_3, ISO3CD) |> 
  right_join(key3) |> 
  # to join with FAO data from website and Stefania.
  mutate(O_Name_En_Emmanuel = O_Name_En) 

duplicated(fao_ema)
head(fao_ema)
```

# Add additional information from Stefania's FAO dataset, just for checking
Columns:
`ISO3_Code` = ISO code for countries and territories 
`Official_Name_En` = official FAO name 

```{r}
fao_st <- file.path(base_folder, "Country_Item_14022024.csv") |>
  read.csv() |>
  unique()

fao_st <- fao_st |> 
  select(c(ISO3_Code, Official_Name_En)) |> 
  unique() |> 
  rename(ISO3CD = ISO3_Code, O_Name_En_Stefania = Official_Name_En) |> 
  # to join with website and Emmanuel information.
  mutate(O_Name_En = O_Name_En_Stefania)

head(fao_st)
sort(unique(fao_ema$ISO3CD))
sort(unique(fao_st$ISO3CD))
```

# Merge all info into a final FAO dataset

```{r}
# Emmanuel + Stefania
fao_final <- fao_ema |> 
  full_join(fao_st)

# + website
fao_final <- fao_final |> 
  full_join(fao) |> 
  # not needed anymore as the merging is done. 
  select(-O_Name_En) 
```

# Adjust columns

```{r}
fao_final2 <- fao_final |> 
  rename(ISO3_country = ISO_3, ISO3_territory = ISO3CD, Country_name = Country, 
         Continent_ID = CONTCD, Romname_Emmanuel = ROMNAM, 
         Maplab_Emmanuel = MAPLAB) |> 
  mutate(Continent_name = case_when(Continent_ID == "ASI" ~ "Asia",
                                    Continent_ID == "AFR" ~ "Africa",
                                    Continent_ID == "AME" ~ "America",
                                    Continent_ID == "OCE" ~ "Oceania",
                                    Continent_ID == "EUR" ~ "Europe",
                                    .default = Continent_ID)) |> 
  relocate(O_name_En_Website, Short_name_En_Website, Country_name,
           Continent_name, Continent_ID, ISO3_country, ISO3_territory,
           O_Name_En_Emmanuel, O_Name_En_Stefania, Romname_Emmanuel,
           Maplab_Emmanuel, Short_name_En_Website2) |> 
  arrange(Country_name, O_name_En_Website)
```

# **PRINT 1**: Print datasets
Using the EEZ file as master, manually match FAO names, continent and sovereign country.

```{r, eval = F}
write.csv(eez, file.path(base_folder, "EEZ_key.csv"))
write.csv(fao_final, file.path(base_folder, "FAO_key.csv"))
```

# Load the new file and clean it 

```{r}
#EEZs not included in FAO data
not_fao <- c("Amsterdam Island & St. Paul Islands", "Crozet Islands", 
             "Europa Island", "Oecussi Ambeno", 
             "Protected Zone established under the Torres Strait Treaty")

new_key <- file.path(base_folder, "EEZ_FAO_key.csv") |> 
  read_csv() |> 
  # remove join and overlapping 
  filter(str_detect(fa_ffcl, "Overlapping|Joint", negate = T)) |> 
  # remove EEZs not included in FAO data and with no information on the country 
  #they belong to as per "assumptions" files 
  filter(!fa_ffcl %in% not_fao) |> 
  # remove FAO major fishing areas 
  filter(is.na(OCEAN))

head(new_key)
# filter(new_key, fa_ffcl == "Amsterdam Island & St. Paul Islands")
# filter(new_key, 
#        fa_ffcl == "Overlapping claim Falkland / Malvinas Islands: UK / Argentina")

# consider only columns needed 
new_key <- new_key |> 
  select(c(ID_mrgd, continent, fa_ffcl, Country_name)) |> 
  rename(fao_official = fa_ffcl, Country_FAO = Country_name)

## this was printed and called eez_keys_camilla_new.csv 
```

# Add short names as per new FAO regulation where they can be used in figs
new column considered: 
`LIST NAME` = Short Name for Lists and Tables

```{r}
fao <- file.path(base_folder, "nocsDataExport_20240501-114742_English.csv") |> 
  read_csv() |> 
  #Select relevant columns
  select(c(`FULL NAME`, `LIST NAME`)) |> 
  #Update column names to something more descriptive
  rename(O_name_En_Website =`FULL NAME`, 
         figure_name =`LIST NAME`) |> 
  rename(fao_official = O_name_En_Website)

new_key <- new_key |> 
  left_join(fao)
```

# **PRINT 2**: Print the dataset to: 
1. Add missing short names for those countries where fao_official was added in previous steps  
2. save the file as csv UTF-8 to adjust names with special characters  

```{r}
# missing short names 
filter(new_key, is.na(figure_name))
# e.g. fao_official: the Federative Republic of Brazil (Trindade) -> 
# List name: Brazil (Trindade)

# small adjustments: 
# the United States Minor Outlying Islands (Howland and Baker Islands) -> 
# Howland and Baker Islands
# the Commonwealth of the Northern Mariana Islands -> USA as sovereign country

# special characters 
# "Cura\x8dao"
sort(new_key$fao_official)[18] 
# "R\x8eunion"
sort(new_key$fao_official)[47] 
# "Saint Barth\x8elemy"
sort(new_key$fao_official)[49] 
# "the Republic of Colombia (Quitasueño)" 
# NOTE: changed to Quitasueño in csv before saving
sort(new_key$fao_official)[135] 
# "the Republic of C\x99te d'Ivoire"   
sort(new_key$fao_official)[141] 
# "the Republic of T\x9frkiye" 
sort(new_key$fao_official)[196] 

Encoding(sort(new_key$fao_official)[18])
Encoding(sort(new_key$figure_name)[135])

# print new final file 
# write_csv(new_key, file.path(base_folder, "EEZ_FAO_key_with_short_names.csv"))
```

# Load **FINAL** file to check 

```{r}
final <- file.path(base_folder, "EEZ_FAO_key_with_short_names_utf.csv") |> 
  read.csv(fileEncoding = "UTF-8")

# all good
sort(final$figure_name)[46]
sort(final$figure_name)[53]
sort(final$figure_name)[56]
sort(final$figure_name)[169]
sort(final$figure_name)[173]
sort(final$figure_name)[209]
```

# Preparing file to be used in masks
  
```{r}
eez <- eez |> 
  #Merge final dataset above with information from FAO-EEZ shapefile
  left_join(final, join_by(ID_mrgd)) |> 
  #Ensure any rows with now continent information are updated
  mutate(continent = case_when(is.na(continent) ~ CONTINENT,
                               T ~ continent)) |> 
  #Remove columns that are not relevant
  select(!c(F_CODE, OCEAN, MRGID:CONTINENT))

#Save results
eez |> 
  write_csv(file.path(base_folder, "FAO-Major-Areas_EEZ_continent_keys.csv"))
```



