---
title: "Creating masks to extract ESM data"
author: "Denisse Fierro Arcos"
date: "2024-05-13"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Introduction
In this notebook, we will create masks for GFDL and IPSL forced FishMIP models, which will include area of the grid cell, as well as FAO regions and country names.  
  
## Loading libraries
  
```{r}
library(terra)
library(sf)
library(readr)
library(dplyr)
library(tidyr)
```
  
## Setting up
  
```{r}
#Defining base directory
base_dir <- "/rd/gem/private/shared_resources/"

#Loading FAO-EEZ shapefile
fao_eez_shp <- list.files(base_dir, "FAO_MajorAreas_200NM.shp", 
                          recursive = T, full.names = T) |> 
  read_sf()

#Loading FAO-LME shapefile
fao_lme_shp <- list.files(base_dir, "FAO-MajorAreas_LMEs.shp", recursive = T,
                          full.names = T) |> 
  read_sf() 

#Loading GFDL area grid cell file
gfdl <- list.files(base_dir, 
                   "gfdl-esm4_areacello_w-fractions_60arcmin_global_fixed.nc",
                   recursive = T, full.names = T) |> 
  rast()

#Change grid cells with a value of 0 to NA
gfdl[gfdl == 0] <- NA

#Change variable name
names(gfdl) <- "cell_area"

#Loading IPSL area grid cell file
ipsl <- list.files(base_dir, 
                   "ipsl-cm6a-lr_areacello_wo-fractions_60arcmin_global_fixed.nc",
                   recursive = T, full.names = T) |> 
  rast()
```
  
## Defining functions
The first function will rasterize the shapefile and save the resulting raster. While the second function will transform the raster into a data frame and save it.  
  
```{r}
rast_save <- function(shp, ras, field, touches = T, fun = min, out_file = NULL){
  #Inputs:
  # - shp (sf object): shapefile containing limits of regions to be rasterised
  # - ras (SpatRaster): sample raster to be used for rasterising
  # - field (character): name of shp column to be used in rasterise function
  # - touches (boolean): default is TRUE - grid cells touched by polygon 
  #                      boundary to be included in rasterised region
  # - fun (function): default min. If multiple regions fall into a single cell,
  #                   define what function to be used by rasterise function
  # - out_file (character): OPTIONAL. File path to save rasterised shapefile
  
  #Applying rasterisation
  shp_ras <- rasterize(shp, ras, field = shp[[field]], touches = touches,
                       fun = fun)
  #Saving output
  if(!is.null(out_file)){
    writeCDF(shp_ras, out_file, varname = "region", 
           longname = "Fish-MIP regions", missval = NA, overwrite = T)
  }

  #Returning rasterised file
  return(shp_ras)
}

rast_df <- function(ras_mask, ras_area, metadata, out_file = NULL){
  #Inputs:
  # - ras_mask (SpatRaster): raster mask to be converted to data frame
  # - ras_area (SpatRaster): raster containing area of grid cell
  # - metadata (data frame): contains metadata to interpret mask
  # - out_file (character): OPTIONAL. File path to save data frame

  #Change raster mask to data frame
  mask_df <- ras_mask |>
    as.data.frame(xy = T) 
  
  #Change area file to data frame
  area_df <- ras_area |> 
    as.data.frame(xy = T) |> 
    #Join to raster mask
    left_join(mask_df, join_by(x, y)) |> 
    #Remove rows with no data in the "cell_area" column
    drop_na(cell_area) |> 
    #Join with metadata file
    left_join(metadata, join_by(ID_mrgd))
  
  #Save result
  if(!is.null(out_file)){
    area_df |> 
      write_csv(out_file)
  }
  
  #Return data frame of raster mask
  return(area_df)
}
```
  
## Applying rasterise function
  
```{r}
#Defining output folder
fao_dir <- "/rd/gem/private/users/camillan/FAO_Report/Merged_FAO_EEZ"

#Creating GFDL mask
gfdl_mask <- rast_save(fao_eez_shp, gfdl, "ID_mrgd",
                       out_file = file.path(fao_dir, 
                                            "gfdl-esm4_60arcmin_global_FAO-EEZ.nc"))

gfdl_lme <- rast_save(fao_lme_shp, gfdl, "ID_merged", 
                      out_file = "../data/gfdl-esm4_60arcmin_global_FAO-LME.nc")

#Creating IPSL mask
ipsl_mask <- rast_save(fao_eez_shp, ipsl, "ID_mrgd",
                       out_file = file.path(fao_dir, 
                                            "ipsl-cm6a_60arcmin_global_FAO-EEZ.nc"))

ipsl_lme <- rast_save(fao_lme_shp, ipsl, "ID_merged", 
                      out_file = "../data/ipsl-cm6a_60arcmin_global_FAO-LME.nc")

```
  
## Applying function to save raster as data frame
  
```{r}
#Loading metadata file
meta <- list.files(fao_dir, "FAO-Major-Areas_EEZ_continent_keys.csv",
                   recursive = T, full.name = T) |> 
  read_csv()

meta_lme <- list.files(base_dir, "FAO_LME_keys.csv", recursive = T, 
                       full.names = T) |> 
  read_csv() |> 
  rename("ID_mrgd" = "ID_merged")

#Changing the name of layer to match metadata file
names(gfdl_mask) <- "ID_mrgd"
names(ipsl_mask) <- "ID_mrgd"
names(gfdl_lme) <- "ID_mrgd"
names(ipsl_lme) <- "ID_mrgd"

#Saving GFDL raster as data frame
gfdl_df <- rast_df(gfdl_mask, gfdl, meta, 
                   file.path(fao_dir, "gfdl-esm4_60arcmin_global_FAO-EEZ.csv"))

gfdl_lme_df <- rast_df(gfdl_lme, gfdl, meta_lme, 
                       "../data/gfdl-esm4_60arcmin_global_FAO-LME.csv")

#Saving IPSL raster as data frame
ipsl_df <- rast_df(ipsl_mask, ipsl, meta, 
                   file.path(fao_dir, "ipsl-cm6a_60arcmin_global_FAO-EEZ.csv"))

ipsl_lme_df <- rast_df(ipsl_lme, ipsl, meta_lme, 
                       "../data/ipsl-cm6a_60arcmin_global_FAO-LME.csv")


# Check missing territories: 
head(meta) # names on file printed in 00 and including all countries and territories 
head(gfdl_df) # names on files just created

setdiff(sort(unique(meta$figure_name)), sort(unique(gfdl_df$figure_name)))
setdiff(sort(unique(meta$figure_name)), sort(unique(ipsl_df$figure_name)))

```
  
## Saving shapefile with updated continent names
  
```{r}
#Keep relevant columns from metadata
meta <- meta |> 
  select(ID_mrgd, continent)

fao_eez_shp |> 
  #Keep only relevant columns in shapefile
  select(NAME_EN, ID_mrgd, nm_mrgd) |> 
  #Replacing with updated continent column
  left_join(meta, join_by(ID_mrgd)) |> 
  #Saving result in FAO report folder
  write_sf(file.path(fao_dir, "FAO_MajorAreas_200NM.shp"), append = F)
```
  
Masks are now ready to be used in the workflow described in this repository.  
  