---
title: "Plotting biomass change under different SSPs"
author: "Denisse Fierro Arcos & Camilla Novaglio"
date: "2024-01-09"
output: pdf_document
---

# Introduction

In this notebook, we will calculate the mean percentage change in fish biomass between 2041 and 2050 under two climate emission scenarios: `SSP1-2.6` (low emissions) and `SSP5-8.5` (very high emissions). The last decade from the historical period (2005-2014) was used as the reference period.  
  
Note that outputs from all [global FishMIP models](https://fishmip.org/modellingteams.html) were used to calculate the mean percentage change for the entire ensemble.  
  
## Loading libraries
  
```{r, message=FALSE, warning=FALSE}
#Data wrangling
library(tidyverse)
library(data.table)
#Using more than one scale in plots
library(ggnewscale)
#Dealing with raster data
library(terra)
library(tidyterra)
#Dealing with vector data
library(sf)
#Base maps
library(rnaturalearth)
#Color palettes
library(cmocean)
library(scales)
#Combining plots
library(cowplot)
```

## Setting up notebook
This notebook uses a modified version of the Exclusive Economic Zones boundaries (version 11) from the [Flanders Marine Institute (2019)](https://doi.org/10.14284/386). Available [online](https://www.marineregions.org/). The following modifications were made:  
- Boundaries for South Georgia and Sandwich Islands were updated to boundaries provided by the UN  
- Boundaries around Bouvet Island (Norway) in the south Atlantic were added to this shapefile  
- Country boundaries were dissolved and polygons were reclassified by continent  
  
```{r}
#Base folder for project
base_folder <- "/rd/gem/private/users/camillan"

#Defining location of notebook outputs
out_folder <- file.path(base_folder, "FAO_Report")
if(!dir.exists(out_folder)){
  dir.create(out_folder)
}

#Defining location for plots
out_plot <- file.path(base_folder, "FAO_Report/maps_global_regions")
out_plot <- file.path("data/data_products")
if(!dir.exists(out_plot)){
  dir.create(out_plot)
}

#Getting GFDL grid area
area_gfdl_df <- list.files(out_folder, "gfdl-esm4_60arcmin_global_FAO-EEZ.csv",
                           recursive = T, full.names = T) |> 
  read_csv()

#Getting IPSL grid area
area_ipsl_df <- list.files(out_folder, "ipsl-cm6a_60arcmin_global_FAO-EEZ.csv",
                           recursive = T, full.names = T) |> 
  read_csv()
```
  
## Loading relevant shapefiles
We will apply a Robinson projection to all shapefiles before plotting them.  
  
```{r}
#Robinson projection
rob_proj <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs
+type=crs"

#Unprojected UN-approved world base map
world <- read_sf(list.files(out_folder, pattern = "country_bounds.shp",
                            full.names = T, recursive = T))

#Base map
world_proj <- world |> 
  st_transform(rob_proj)

#Boundaries provided by UN
world_bound <- read_sf(list.files(out_folder, pattern = "UN_intbnd.shp", 
                                        full.names = T, recursive = T)) |> 
  #Selecting boundary types to be included in the map
  filter(TYPE < 99) 

#Boundaries provided by UN
world_bound_proj <- world_bound |> 
  st_transform(rob_proj)

#FAO and EEZ boundaries - unprojected
fao_eez <- read_sf(list.files(out_folder, 
                               pattern = "FAO_MajorAreas_200NM_cont.shp", 
                               recursive = T, full.names = T))

#FAO - EEZ boundaries - projected
fao_eez_proj <- fao_eez |> 
   st_transform(rob_proj)

#EEZ boundaries - unprojected
eez <- fao_eez |> 
  filter(is.na(NAME_EN)) |> 
  #Removing empty columns
  select(!NAME_EN)

#EEZ boundaries - projected
eez_proj <- eez |> 
   st_transform(rob_proj)

#FAO regions 
fao_proj <- fao_eez_proj |>
  drop_na(NAME_EN) |> 
  #Removing empty columns
  select(!continent)
```
  
## Figure 2 FAO report - Regions used in report
FAO regions are coloured by ocean basin, and EEZs are classified by continent.  
  
```{r}
main <- ggplot()+
  #Plotting FAO classified by ocean basin
  geom_sf(data = fao_proj, aes(fill = nm_mrgd))+
  #Plotting EEZ classified by continent
  geom_sf(data = eez_proj, aes(fill = continent))+
  #Using colour blind friendly colours
  scale_fill_manual(values = c("Pacific" = "#222255", "Atlantic" = "#225555", 
                               "Indian" = "#225522", "Arctic" = "#666633", 
                               "Oceania" = "#bbccee", "Europe" = "#cceeff", 
                               "Americas" = "#ccddaa", "Africa" = "#eeeebb",
                               "Southern Ocean" = "#663333", 
                               "Asia" = "#ffcccc"))+
  #Plotting base map and country boundaries
  geom_sf(data = world_proj, color = "#dbdade", fill = "#dbdade")+
  #Adding coastlines and international boundaries
  geom_sf(data = world_bound_proj, lwd = 0.25, color = "black", show.legend = F, 
          aes(linetype = as.factor(TYPE)))+
  scale_linetype_manual(values = c("0" = "solid", "1" = "twodash", 
                                   "2" = "dotdash", "3" = "dashed", 
                                   "4" = "dotted"))+
  geom_sf(data = world_bound_proj[world_bound_proj$TYPE == 8,], 
          linetype = "longdash", 
          lwd = 0.25, color = "#838285", show.legend = F)+
  #Removing grey background
  theme_bw()+
  #Formatting legend title
  guides(fill = guide_legend(title = "FAO and EEZ classification", 
                             title.position = "top", title.hjust = 0.5))+
  #Formatting legend
  theme(legend.position = "bottom", panel.border = element_rect(colour = NA),
        legend.margin = margin(0, 0, 0, 0), 
        legend.box.margin = margin(0, 0, 0, 0),
        legend.title = element_text(size = 13), 
        legend.text = element_text(size = 13))
```
  
Saving map showing FAO regions and EEZs classified by continent.  
  
```{r, eval = F}
ggsave(filename = file.path(out_plot, "fao_eez_global_map.pdf"), main,
       device = "pdf", width = 14, height = 9)
```
  
## Figure 3 FAO report - Percentage change per decade and scenario.
In this figure, we will show percentage fish biomass change in the ocean, while yearly mean capture fisheries production per country (2012-2021) will be shown on land.

Capture fisheries production data comes from the [FAO's Fisheries and Aquaculture Data Portal](https://www.fao.org/fishery/statistics-query/en/global_production/global_production_quantity). 
  
```{r}
#Load fisheries data
fisheries_catch <- read_csv("data/world_mean_annual_catch.csv") |> 
  mutate(class_fish = factor(class_fish, 
                             levels = c("<0.25", "0.25-1", "1-2", "2-4", 
                                        "4-6", ">6"), ordered = T))

#Now we can add this information to the world maps
#Projected
world_proj <- world_proj |> 
  left_join(fisheries_catch, by = join_by(ISO3CD))

#Unprojected
world <- world |> 
  left_join(fisheries_catch, by = join_by(ISO3CD))
```
  
We will transform the data frame containing the percentage change into a multi-layer raster, and apply a Robinson projection.   
  
```{r}
#Loading ensemble biomass change
maps_data <- read_csv("data/ensemble_perc_bio_change_data_map.csv")

#Extract data for scenario, transform to raster and reproject
maps_data_proj <- maps_data |> 
  #Selecting coordinate columns and percentage change columns
  select(x, y, starts_with("rel_change")) |> 
  #Transform to multi-layer raster
  rast(type = "xyz", crs = "epsg:4326") |> 
  #Reproject to Robinson
  project(y = rob_proj) 
```
  
We will also create a custom-made colour palette for the maps.  
  
```{r}
#Create custom-made color palette
scale_fill_custom <- function(..., alpha = 1, begin = 0, end = 1, direction = 1,
                              option = "D", values = NULL, space = "Lab", 
                              na.value = "white", guide = "colourbar", 
                              aesthetics = "fill") {
  continuous_scale(aesthetics, scale_name = "custom",
                   palette = gradient_n_pal(c(cmocean("matter", start = 0.1, 
                                                      end = 0.8, 
                                                      direction = -1)(123),
                                              cmocean("delta", start = 0.49, 
                                                      end = 0.5)(20),
                                              cmocean("deep", start = 0.1, 
                                                      end = 0.8)(123)), values,
                                            space), 
                   na.value = na.value, guide = guide, ...)
}
```
  
We will plot each decade/scenario combination separately and then create a multi-panel figure with all maps.  
  
```{r}
#Define base steps for maps
base_gg <- list(scale_fill_binned(limits = c(-50, 50), n.breaks = 8,
                                  type = scale_fill_custom,
                                  oob = oob_squish),
                new_scale_fill(),
                geom_sf(inherit.aes = F, data = world_proj, color = NA, 
                        aes(fill = class_fish), show.legend = F),
                scale_fill_brewer(palette = "Blues", na.value = "#dcdcdc"),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = world_bound_proj, lwd = 0.25, 
                        color = "black", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = c("1" = "twodash", 
                                                 "2" = "dotdash", 
                                                 "3" = "dashed", 
                                                 "4" = "dotted")),
                geom_sf(inherit.aes = F, 
                        data = world_bound_proj[world_bound_proj$ISO3_CNT1 == "ATA",],
                        lwd = 0.25, linetype = "solid", color = "#5b5b5b", 
                        show.legend = F),
                geom_sf(inherit.aes = F, 
                        data = world_bound_proj[world_bound_proj$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                geom_sf(inherit.aes = F, data = eez_proj, fill = NA,
                        colour = "#5b5b5b", show.legend = F, linewidth = 0.25),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5),
                      legend.position = "none"))
```
  
Since we are using the same legend for data shown on land and the ocean, we will extract them once and apply them to all maps. To do this we will use the `get_legend` function from the `cowplot` package, but this option may not work with all operating system.  
  
If you get any errors, you can using the `get_plot_component` function instead. Replace line 275 with `get_plot_component(land, 'guide-box-bottom', return_all = TRUE)`, where `land` refers to the map with data plotted on land only. Make sure to do the same for the fish biomass map in line 288.   
  
```{r}
#Plotting land masses to get legend
land <- ggplot()+
  geom_sf(data = world_proj, aes(fill = class_fish))+
  scale_fill_brewer(palette = "Blues", na.value = "#dcdcdc", 
                    name = "Yearly landings between 2012-2021 (million tonnes)",
                    labels = c("<0.25", "0.25-1", "1-2", "2-4", "4-6", ">6",
                               "No data"))+
  guides(fill = guide_legend(title.position = "top", title.hjust = 0.5, 
                             barwidth = 15, 
                             label.position = "bottom", nrow = 1, 
                             keywidth = unit(1.25, "cm"), label.hjust = 0.5))+
  theme(legend.position = "bottom")

#Get legend for land
leg_land <- get_legend(land)
leg_land<-get_plot_component(land, 'guide-box-bottom', return_all = TRUE)

#Plotting fish biomass to get legend
fish <- ggplot()+
  geom_spatraster(data = maps_data_proj$rel_change_mean50_ssp126_mean)+
  scale_fill_binned(limits = c(-50, 50), n.breaks = 8, oob = oob_squish, 
                    type = scale_fill_custom, 
                    name = "Change in fish biomass (%)")+
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5, 
                               barwidth = 15))+
  theme(legend.position = "bottom")

#Get legend for fish biomass
leg_fish <- get_legend(fish)
leg_fish<-get_plot_component(fish, 'guide-box-bottom', return_all = TRUE)
```
  
We can now plot the maps for all scenario and period combinations. 
  
```{r}
#SSP1-2.6 2041-2050
p50_126_change <- ggplot()+ 
  geom_spatraster(data = maps_data_proj$rel_change_mean50_ssp126_mean)+
  labs(title = "SSP1-2.6: 2041-2050")+
  base_gg

#SSP5-8.5 2041-2050
p50_585_change <- ggplot()+
  geom_spatraster(data = maps_data_proj$rel_change_mean50_ssp585_mean)+
  labs(title = "SSP5-8.5: 2041-2050")+
  base_gg

#SSP1-2.6 2091-2100
p00_126_change <- ggplot()+
  geom_spatraster(data = maps_data_proj$rel_change_mean00_ssp126_mean)+
  labs(title = "SSP1-2.6: 2091-2100")+
  base_gg

#SSP5-8.5 2091-2100
p00_585_change <- ggplot()+
  geom_spatraster(data = maps_data_proj$rel_change_mean00_ssp585_mean)+
  labs(title = "SSP5-8.5: 2091-2100")+
  base_gg

#Plotting everything together
all_plots <- plot_grid(plot_grid(p50_126_change, p50_585_change, ncol = 2, 
                                 nrow = 1, labels = c("a", "b"), label_x = 0.1),
                       plot_grid(p00_126_change, p00_585_change, ncol = 2, 
                                 nrow = 1, labels = c("c", "d"), label_x = 0.1),
                       plot_grid(leg_land, leg_fish, nrow = 1, ncol = 2), 
                       ncol = 1, nrow = 3, rel_heights = c(1, 1, 0.4))
```
  
We can save the final figure including four maps.    
  
```{r}
# Saving multi-panel plot
ggsave(file.path(out_plot, "global_perc_change_map_50-00.pdf"), all_plots,
       device = "pdf", width = 14, height = 8)
```
  
## Figure 5 FAO report - Asia 
  
```{r}
sf_use_s2(F)

#Extracting data for Asia
asia_data <- maps_data |> 
  filter(continent == "Asia") 

#Selecting Asian countries
asia_cent <- world |> 
  st_crop(st_bbox(c(xmin = 20, ymin = -20, xmax = 165, ymax = 55))) |> 
  #Add empty row to add >6 "fish class"
  add_row(class_fish = "4-6", CONTCD = "ASI") |> 
  #Ensure classes are ordered prior to plotting
  mutate(class_fish = factor(class_fish, 
                             levels = c("<0.25", "0.25-1", "1-2", "2-4", 
                                        "4-6", ">6"), ordered = T))
```
    
We will plot each decade/scenario combination separately and then create a multi-panel figure with all maps.  
  
```{r}
#Define base steps for maps
base_gg <- list(scale_fill_binned(limits = c(-50, 50), n.breaks = 8,
                                  type = scale_fill_custom,
                                  oob = oob_squish),
                new_scale_fill(),
                geom_sf(inherit.aes = F, data = eez[eez$continent == "Asia",], 
                         fill = NA, colour = "#5b5b5b", show.legend = F,
                         linewidth = 0.25),
                geom_sf(inherit.aes = F, 
                        data = asia_cent[asia_cent$CONTCD == "ASI",], 
                        color = NA, aes(fill = class_fish), show.legend = F),
                geom_sf(inherit.aes = F, 
                        data = asia_cent[asia_cent$CONTCD != "ASI",], 
                        color = "#dbdade", fill = "#dbdade"),
                scale_fill_brewer(palette = "Blues", na.value = "#dcdcdc"),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = world_bound_proj, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = c("0" = "solid",
                                                 "1" = "twodash", 
                                                 "2" = "dotdash", 
                                                 "3" = "dashed", 
                                                 "4" = "dotted")),
                geom_sf(inherit.aes = F, 
                        data = world_bound_proj[world_bound_proj$ISO3_CNT1 == "ATA",],
                        lwd = 0.25, linetype = "solid", color = "#5b5b5b", 
                        show.legend = F),
                geom_sf(inherit.aes = F, 
                        data = world_bound_proj[world_bound_proj$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5),
                      legend.position = "none"),
                lims(x = c(25, 160), y = c(-15, 50)))
```

We will apply the plotting instructions to the four maps that we need in the figure for Asia.  

```{r}
#SSP1-2.6 2041-2050
p50_126 <- asia_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2041-2050")+
  base_gg

#SSP5-8.5 2041-2050
p50_585 <- asia_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2041-2050")+
  base_gg

#SSP1-2.6 2091-2100
p00_126 <- asia_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2091-2100")+
  base_gg
  
#SSP5-8.5 2091-2100
p00_585 <- asia_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2091-2100")+
  base_gg

#Plotting everything together
asia_plots <- plot_grid(plot_grid(p50_126, p50_585, ncol = 2, nrow = 1,
                                  labels = c("a", "b"), label_x = 0.1),
                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1,
                                  labels = c("c", "d"), label_x = 0.1),
                        #Using legends from Figure 3
                        plot_grid(leg_land, leg_fish, nrow = 1, ncol = 2), 
                        ncol = 1, nrow = 3, rel_heights = c(1, 1, 0.4))

#Removing variable
rm(asia_cent)
```
  
```{r}
# Saving multi-panel plot
ggsave(file.path(out_plot, "asia_perc_change_map_50-00.pdf"), asia_plots, 
       device = "pdf", width = 14, height = 8)
```
  
## Figure 7 FAO report - Americas
As we did with Europe, we would like the Americas to show in the middle of the map, so we will have to pre-process our data before plotting the maps.  
  
```{r}
#Split world in two hemispheres 
west <- st_crop(world, st_bbox(c(xmin = -180, ymin = -60, 
                                 xmax = 20, ymax = 90)))
east <- st_crop(world, st_bbox(c(xmin = 160, ymin = -60, 
                                 xmax = 180, ymax = 90)))
#Merge them together
amercent <- rbind(west, 
                  st_set_crs(st_transform(east, 
                                          "+proj=longlat +lon_wrap=-120"),
                             st_crs(west))) |> 
  #Add empty row to add >6 "fish class"
  add_row(class_fish = ">6", CONTCD = "AME") |> 
  #Ensure classes are ordered prior to plotting
  mutate(class_fish = factor(class_fish, 
                             levels = c("<0.25", "0.25-1", "1-2", "2-4", 
                                        "4-6", ">6"), ordered = T))

#Split world boundaries in two hemispheres 
west <- st_crop(world_bound, st_bbox(c(xmin = -180, ymin = -60, 
                                 xmax = 20, ymax = 90)))
east <- st_crop(world_bound, st_bbox(c(xmin = 160, ymin = -60, 
                                 xmax = 180, ymax = 90)))
#Merge them together
amercent_bound <- rbind(west, 
                        st_set_crs(st_transform(east, 
                                                "+proj=longlat +lon_wrap=-120"),
                                   st_crs(west))) 

#Split EEZ in two hemispheres
eez_west <- eez |> 
  filter(continent == "Americas") |> 
  st_crop(st_bbox(c(xmin = -180, ymin = -65, xmax = 20, ymax = 90)))
eez_east <- eez |> 
  filter(continent == "Americas") |> 
  st_crop(st_bbox(c(xmin = 160, ymin = -65, xmax = 180, ymax = 90)))
#Merge together
eez_amer <- rbind(eez_west,
                  st_set_crs(st_transform(eez_east, 
                                          "+proj=longlat +lon_wrap=-120"),
                             st_crs(eez_west))) 

#Removing variables that are not needed
rm(eez_east, eez_west, east, west)

#Extracting data for the Americas
americas_data <- maps_data |>
  filter(continent == "Americas") |> 
  #Selecting relevant columns
  select(x, y, starts_with("rel_change")) |> 
  #Converting to raster
  rast(crs = "epsg:4326") |> 
  #Reprojecting so raster is centered in the Americas
  project(rast(ext(-200, 20, -65, 90), res = 1)) |> 
  #Return to data frame
  as.data.frame(xy = T)
```
  
We will plot each decade/scenario combination separately and then create a multi-panel figure with all maps.  
  
```{r}
#Define linetypes to be used in plots
line_manual <- c("0" = "solid", "1" = "twodash", "2" = "dotdash", 
                 "3" = "dashed", "4" = "dotted")

#Define base steps for maps
base_gg <- list(scale_fill_binned(limits = c(-50, 50), n.breaks = 8, 
                                  type = scale_fill_custom, 
                                  oob = oob_squish),
                new_scale_fill(),
                geom_sf(inherit.aes = F, data = eez_amer, fill = NA, 
                        colour = "#5b5b5b", show.legend = F, linewidth = 0.25),
                geom_sf(inherit.aes = F, 
                         data = amercent[amercent$CONTCD == "AME",], 
                         aes(fill = class_fish), colour = NA),
                geom_sf(inherit.aes = F, 
                        data = amercent[amercent$CONTCD != "AME",],
                        color = "#dbdade", fill = "#dbdade"),
                scale_fill_brewer(palette = "Blues", na.value = "#dcdcdc"),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = amercent_bound, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = line_manual),
                geom_sf(inherit.aes = F, 
                        data = amercent_bound[amercent_bound$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5),
                      legend.position = "none"),
                coord_sf(ylim = c(-70, 87), expand = F))

#SSP1-2.6 2041-2050
p50_126 <- americas_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2041-2050")+
  base_gg

#SSP5-8.5 2041-2050
p50_585 <- americas_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2041-2050")+
  base_gg

#SSP1-2.6 2091-2100
p00_126 <- americas_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2091-2100")+
  base_gg
  
#SSP5-8.5 2091-2100
p00_585 <- americas_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2091-2100")+
  base_gg

#Plotting everything together
americas_plots <- plot_grid(plot_grid(p50_126, p50_585, ncol = 2, nrow = 1, 
                                      labels = c("a", "b"), label_x = 0.1),
                            plot_grid(p00_126, p00_585, ncol = 2, nrow = 1,
                                      labels = c("c", "d"), label_x = 0.1),
                            #Using legends from Figure 3
                            plot_grid(leg_land, leg_fish, nrow = 1, ncol = 2), 
                            ncol = 1, nrow = 3, rel_heights = c(1, 1, 0.4))

#Removing variables not needed
rm(amercent, amercent_bound, eez_amer)
```
  
```{r}
# Saving multi-panel plot
ggsave(file.path(out_plot, "americas_perc_change_map_50-00.pdf"), 
       americas_plots, device = "pdf", width = 14, height = 9)
```  
  
## Figure 9 FAO report - Europe
Since we want to show Europe in the middle of the map, we will have to project our data and split the base map in two hemispheres.  
  
```{r}
#Split world in two hemispheres
east <- st_crop(world, st_bbox(c(xmin = -50, ymin = 20, 
                                 xmax = 180, ymax = 90)))
west <- st_crop(world, st_bbox(c(xmin = -180, ymin = 20, 
                                 xmax = -150, ymax = 90)))
#Merge them together
eurocent <- rbind(st_set_crs(st_transform(west, "+proj=longlat +lon_wrap=180"), 
                             st_crs(east)), east) |> 
  #Add empty row to add >6 "fish class"
  add_row(class_fish = ">6", CONTCD = "EUR") |> 
  #Ensure classes are ordered prior to plotting
  mutate(class_fish = factor(class_fish, 
                             levels = c("<0.25", "0.25-1", "1-2", "2-4", 
                                        "4-6", ">6"), ordered = T))

#Split world boundaries in two hemispheres
east <- st_crop(world_bound, st_bbox(c(xmin = -50, ymin = 20, 
                                 xmax = 180, ymax = 90)))
west <- st_crop(world_bound, st_bbox(c(xmin = -180, ymin = 20, 
                                 xmax = -150, ymax = 90)))
#Merge them together
eurocent_bound <- rbind(st_set_crs(st_transform(west, 
                                                "+proj=longlat +lon_wrap=180"), 
                             st_crs(east)), east) 

#Split EEZ in two hemispheres
eez_east <- st_crop(eez[eez$continent == "Europe",], 
                    st_bbox(c(xmin = -50, ymin = 20, xmax = 180, ymax = 90)))
eez_west <- st_crop(eez[eez$continent == "Europe",], 
                    st_bbox(c(xmin = -180, ymin = 20, xmax = -150, ymax = 90)))
#Merge together
eez_euro <- rbind(st_set_crs(st_transform(eez_west, 
                                          "+proj=longlat +lon_wrap=180"), 
                             st_crs(eez_east)), eez_east) 

#Removing variables that are not needed
rm(eez_east, eez_west, east, west)

#Extracting data for Europe
europe_data <- maps_data |>
  filter(continent == "Europe") |> 
  #Selecting relevant columns
  select(x, y, starts_with("rel_change")) |> 
  #Converting to raster
  rast(crs = "epsg:4326") |> 
  #Reprojecting so raster is centered in Europe
  project(rast(ext(-50, 210, 20, 90), res = 1)) |> 
  #Return to data frame
  as.data.frame(xy = T)
```
  
We will plot each decade/scenario combination separately and then create a multi-panel figure with all maps.  
  
```{r}
#Define base steps for maps
base_gg <- list(scale_fill_binned(limits = c(-50, 50), n.breaks = 8, 
                                  type = scale_fill_custom, 
                                  oob = oob_squish),
                new_scale_fill(),
                geom_sf(inherit.aes = F, data = eez_euro, fill = NA, 
                        colour = "#5b5b5b", show.legend = F, linewidth = 0.25),
                 geom_sf(inherit.aes = F, 
                         data = eurocent[eurocent$CONTCD == "EUR",], 
                         aes(fill = class_fish), colour = NA),
                geom_sf(inherit.aes = F, 
                        data = eurocent[eurocent$CONTCD != "EUR",],
                        color = "#dbdade", fill = "#dbdade"),
                scale_fill_brewer(palette = "Blues", na.value = "#dcdcdc"),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = eurocent_bound, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = line_manual),
                geom_sf(inherit.aes = F, 
                        data = eurocent_bound[eurocent_bound$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5),
                      legend.position = "none"),
                coord_sf(ylim = c(23, 87), expand = F))

#SSP1-2.6 2041-2050
p50_126 <- europe_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2041-2050")+
  base_gg

#SSP5-8.5 2041-2050
p50_585 <- europe_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2041-2050")+
  base_gg

#SSP1-2.6 2091-2100
p00_126 <- europe_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2091-2100")+
  base_gg
  
#SSP5-8.5 2091-2100
p00_585 <- europe_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2091-2100")+
  base_gg

#Plotting everything together
europe_plots <- plot_grid(plot_grid(p50_126, p50_585, ncol = 2, nrow = 1,
                                    labels = c("a", "b"), label_x = 0.1),
                          plot_grid(p00_126, p00_585, ncol = 2, nrow = 1,
                                    labels = c("c", "d"), label_x = 0.1),
                          #Using legends from Figure 3
                          plot_grid(leg_land, leg_fish, nrow = 1, ncol = 2), 
                          ncol = 1, nrow = 3, rel_heights = c(1, 1, 0.4))

#Removing variables not needed
rm(eez_euro, eurocent, eurocent_bound)
```
  
```{r}
# Saving multi-panel plot
ggsave(file.path(out_plot, "europe_perc_change_map_50-00.pdf"),
       europe_plots, device = "pdf", width = 14, height = 8)
```
  
## Figure 11 FAO report - Africa
  
```{r}
#Extract data for scenario, transform to raster and reproject
africa_data <- maps_data |> 
  filter(continent == "Africa")

#Selecting African countries
africa_cent <- world |> 
  st_crop(st_bbox(c(xmin = -35, ymin = -60, xmax = 90, ymax = 45))) |> 
  #Add empty row to add 2-4 "fish class"
  add_row(class_fish = "2-4", CONTCD = "AFR") |> 
  #Add empty row to add 4-6 "fish class"
  add_row(class_fish = "4-6", CONTCD = "AFR") |> 
  #Add empty row to add >6 "fish class"
  add_row(class_fish = ">6", CONTCD = "AFR") |> 
  #Ensure classes are ordered prior to plotting
  mutate(class_fish = factor(class_fish, 
                             levels = c("<0.25", "0.25-1", "1-2", "2-4", 
                                        "4-6", ">6"), ordered = T))
```
  
We will plot each decade/scenario combination separately and then create a multi-panel figure with all maps.  
  
```{r}
#Define base steps for maps
base_gg <- list(scale_fill_binned(limits = c(-50, 50), n.breaks = 8, 
                                  type = scale_fill_custom, 
                                  oob = oob_squish),
                new_scale_fill(),
                geom_sf(inherit.aes = F, data = eez[eez$continent == "Africa",], 
                         fill = NA, colour = "#5b5b5b", show.legend = F,
                         linewidth = 0.25),
                geom_sf(inherit.aes = F, 
                        data = africa_cent[africa_cent$CONTCD == "AFR",], 
                        color = NA, aes(fill = class_fish), show.legend = F),
                geom_sf(inherit.aes = F, 
                        data = africa_cent[africa_cent$CONTCD != "AFR",], 
                        color = "#dbdade", fill = "#dbdade"),
                scale_fill_brewer(palette = "Blues", na.value = "#dcdcdc"),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = world_bound, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = line_manual),
                geom_sf(inherit.aes = F, 
                        data = world_bound[world_bound$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5),
                      legend.position = "none"),
                lims(x = c(-29, 85), y = c(-60, 40)))

#SSP1-2.6 2041-2050
p50_126 <- africa_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2041-2050")+
  base_gg

#SSP5-8.5 2041-2050
p50_585 <- africa_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2041-2050")+
  base_gg

#SSP1-2.6 2091-2100
p00_126 <- africa_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2091-2100")+
  base_gg
  
#SSP5-8.5 2091-2100
p00_585 <- africa_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2091-2100")+
  base_gg

#Plotting everything together
africa_plots <- plot_grid(plot_grid(p50_126, p50_585, ncol = 2, nrow = 1,
                                    labels = c("a", "b"), label_x = 0.1),
                          plot_grid(p00_126, p00_585, ncol = 2, nrow = 1,
                                    labels = c("c", "d"), label_x = 0.1),
                          #Using legends from Figure 3
                          plot_grid(leg_land, leg_fish, nrow = 1, ncol = 2), 
                          ncol = 1, nrow = 3, rel_heights = c(1, 1, 0.4))

#Removing variable not needed
rm(africa_cent)
```
  
```{r}
# Saving multi-panel plot
ggsave(file.path(out_plot, "africa_perc_change_map_50-00.pdf"),
       africa_plots, device = "pdf", width = 14, height = 9)
```
  
## Figure 13 FAO report - Oceania

```{r}
#Split world in two hemispheres 
west <- st_crop(world, st_bbox(c(xmin = -180, ymin = -60, 
                                 xmax = -100, ymax = 30)))
east <- st_crop(world, st_bbox(c(xmin = 60, ymin = -60, 
                                 xmax = 180, ymax = 30)))
#Merge them together
oceancent <- rbind(st_set_crs(st_transform(west, "+proj=longlat +lon_wrap=180"), 
                              st_crs(east)), east) |> 
  #Add empty row to add 2-4 "fish class"
  add_row(class_fish = "1-2", CONTCD = "OCE") |> 
  #Add empty row to add 2-4 "fish class"
  add_row(class_fish = "2-4", CONTCD = "OCE") |> 
  #Add empty row to add 4-6 "fish class"
  add_row(class_fish = "4-6", CONTCD = "OCE") |> 
  #Add empty row to add >6 "fish class"
  add_row(class_fish = ">6", CONTCD = "OCE") |> 
  #Ensure classes are ordered prior to plotting
  mutate(class_fish = factor(class_fish, 
                             levels = c("<0.25", "0.25-1", "1-2", "2-4", 
                                        "4-6", ">6"), ordered = T))

#Split world boundaries in two hemispheres 
west <- st_crop(world_bound, st_bbox(c(xmin = -180, ymin = -60, 
                                             xmax = -100, ymax = 30)))
east <- st_crop(world_bound, st_bbox(c(xmin = 60, ymin = -60, 
                                             xmax = 180, ymax = 30)))
#Merge them together
oceancent_bound <- rbind(st_set_crs(st_transform(west, 
                                                 "+proj=longlat +lon_wrap=180"),
                                    st_crs(east)), east)

#Split EEZ in two hemispheres
eez_west <- eez |> 
  filter(continent == "Oceania") |> 
  st_crop(st_bbox(c(xmin = -180, ymin = -60, xmax = -100, ymax = 30)))
eez_east <- eez |> 
  filter(continent == "Oceania") |> 
  st_crop(st_bbox(c(xmin = 60, ymin = -60, xmax = 180, ymax = 30)))
#Merge together
eez_ocean <- rbind(st_set_crs(st_transform(eez_west, 
                                           "+proj=longlat +lon_wrap=180"), 
                              st_crs(eez_east)), eez_east) 

#Removing variables that are not needed
rm(eez_east, eez_west, east, west)

#Extracting data for Oceania
ocean_data <- maps_data |>
  filter(continent == "Oceania") |>
  #Selecting relevant columns
  select(x, y, starts_with("rel_change")) |> 
  #Converting to raster
  rast(crs = "epsg:4326") |> 
  #Reprojecting so raster is centered in Oceania
  project(rast(ext(60, 260, -60, 30), res = 1)) |> 
  #Return to data frame
  as.data.frame(xy = T)
```
  
We will plot each decade/scenario combination separately and then create a multi-panel figure with all maps.  
  
```{r}
#Define base steps for maps
base_gg <- list(scale_fill_binned(limits = c(-50, 50), n.breaks = 8, 
                                  type = scale_fill_custom, 
                                  oob = oob_squish),
                new_scale_fill(),
                geom_sf(inherit.aes = F, data = eez_ocean, fill = NA, 
                        colour = "#5b5b5b", show.legend = F, linewidth = 0.25),
                 geom_sf(inherit.aes = F, 
                         data = oceancent[oceancent$CONTCD == "OCE",], 
                         aes(fill = class_fish), colour = NA),
                geom_sf(inherit.aes = F, 
                        data = oceancent[oceancent$CONTCD != "OCE",],
                        color = "#dbdade", fill = "#dbdade"),
                scale_fill_brewer(palette = "Blues", na.value = "#dcdcdc"),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = oceancent_bound, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = line_manual),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5),
                      legend.position = "none"))

#SSP1-2.6 2041-2050
p50_126 <- ocean_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2041-2050")+
  base_gg

#SSP5-8.5 2041-2050
p50_585 <- ocean_data |> 
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2041-2050")+
  base_gg

#SSP1-2.6 2091-2100
p00_126 <- ocean_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  labs(title = "SSP1-2.6: 2091-2100")+
  base_gg
  
#SSP5-8.5 2091-2100
p00_585 <- ocean_data |> 
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  labs(title = "SSP5-8.5: 2091-2100")+
  base_gg

#Plotting everything together
oceania_plots <- plot_grid(plot_grid(p50_126, p50_585, ncol = 2, nrow = 1, 
                                     labels = c("a", "b"), label_x = 0.1),
                           plot_grid(p00_126, p00_585, ncol = 2, nrow = 1,
                                     labels = c("c", "d"), label_x = 0.1),
                           #Using legends from Figure 3
                           plot_grid(leg_land, leg_fish, nrow = 1, ncol = 2), 
                           ncol = 1, nrow = 3, rel_heights = c(1, 1, 0.4))

#Removing variables not needed
rm(oceancent, oceancent_bound, eez_ocean)
```
  
```{r}
# Saving multi-panel plot
ggsave(file.path(out_plot, "oceania_perc_change_map_50-00.pdf"),
       oceania_plots, device = "pdf", width = 14, height = 8)
```
    
## Figure 15 FAO report - FAO regions
Finally, we will create maps for mean biomass change within the FAO regions.  
  
```{r}
#Listing all relevant files to calculate biomass projections
maps_data <- read_csv("data/ensemble_perc_bio_change_fao_map.csv")

#Add to FAO EEZ mask
choro_data_proj <- area_gfdl_df |> 
  left_join(maps_data, by = "ID_mrgd") |> 
  select(x, y, starts_with("rel_change")) |> 
  #Transform to multi-layer raster
  rast(type = "xyz", crs = "epsg:4326") |> 
  #Reproject to Robinson
  project(y = rob_proj) 
```

  
```{r}
#Define base steps for maps
base_gg <- list(scale_fill_binned(limits = c(-50, 50), n.breaks = 8,
                                  type = scale_fill_custom, 
                                  oob = oob_squish),
                geom_sf(inherit.aes = F, data = fao_proj, fill = NA, 
                        colour = "#5b5b5b",
                        show.legend = F, linewidth = 0.25),
                geom_sf(inherit.aes = F, data = eez_proj, fill = "white",
                        colour = "white", show.legend = F, linewidth = 0.25),
                geom_sf(inherit.aes = F, data = world_proj, show.legend = F,
                        color = "#dbdade", fill = "#dbdade"),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = world_bound_proj, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = line_manual),
                geom_sf(inherit.aes = F, 
                        data = world_bound_proj[world_bound_proj$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5),
                      legend.position = "none"))

#SSP1-2.6 2041-2050
p50_126 <- ggplot()+
  geom_spatraster(data = choro_data_proj$rel_change_mean50_ssp126_mean)+
  labs(title = "SSP1-2.6: 2041-2050")+
  base_gg

#SSP5-8.5 2041-2050
p50_585 <- ggplot()+
  geom_spatraster(data = choro_data_proj$rel_change_mean50_ssp585_mean)+
  labs(title = "SSP5-8.5: 2041-2050")+
  base_gg

#SSP1-2.6 2091-2100
p00_126 <- ggplot()+
  geom_spatraster(data = choro_data_proj$rel_change_mean00_ssp126_mean)+
  labs(title = "SSP1-2.6: 2091-2100")+
  base_gg

#SSP5-8.5 2091-2100
p00_585 <- ggplot()+
  geom_spatraster(data = choro_data_proj$rel_change_mean00_ssp585_mean)+
  labs(title = "SSP5-8.5: 2091-2100")+
  base_gg

#Plotting everything together
fao_reg_plots <- plot_grid(plot_grid(p50_126, p50_585, ncol = 2, nrow = 1,
                                     labels = c("a", "b"), label_x = 0.1),
                           plot_grid(p00_126, p00_585, ncol = 2, nrow = 1,
                                     labels = c("c", "d"), label_x = 0.1),
                           leg_fish, ncol = 1, nrow = 3, 
                           rel_heights = c(1, 1, 0.4))
```
  
```{r}
# Saving multi-panel plot
ggsave(file.path(out_plot, "fao_regions_perc_change_map_50-00.pdf"),
       fao_reg_plots, device = "pdf", width = 14, height = 9)
```
  
## Figure 16 FAO report - FAO sustainable fisheries - REMOVED from report
Finally, we will create a choropleth map for the proportion of sustainable fisheries within FAO regions for 2019.  
  
```{r}
# #Loading FAO fisheries data
# fao_fisheries <- read_csv("data/prop_sustainable_fisheries_fao.csv")
# 
# fao_fisheries <- fao_proj |> 
#   left_join(fao_fisheries, join_by(NAME_EN))
```
  
Finally, we can map this information.  
  
```{r}
# fao_fisheries_plot <- ggplot()+
#   #Plotting FAO classified by ocean basin
#   geom_sf(data = fao_fisheries, aes(fill = OBS_VALUE))+
#   geom_sf(data = eez_proj, fill = "white", colour = "white")+
#   scale_fill_distiller(palette = "YlGnBu", direction = 1)+
#   #Plotting base map
#   geom_sf(data = world_proj, color = "#dbdade", fill = "#dbdade")+
#   #Adding coastlines and international boundaries
#   geom_sf(data = world_bound_proj, lwd = 0.25, color = "#484848", 
#           show.legend = F, aes(linetype = as.factor(TYPE)))+
#   scale_linetype_manual(values = line_manual)+
#   geom_sf(data = world_bound_proj[world_bound_proj$TYPE == 8,], 
#           linetype = "longdash", lwd = 0.25, color = "#838285", 
#           show.legend = F)+
#   #Removing grey background
#   theme_bw()+
#   #Formatting legend title
#   guides(fill = guide_colorbar(title = "% sustainable fisheries", 
#                                title.position = "top", title.hjust = 0.5, 
#                                barwidth = 15))+
#   #Formatting legend
#   theme(legend.position = "bottom", panel.border = element_rect(colour = NA),
#         legend.margin = margin(0, 0, 0, 0),
#         legend.box.margin = margin(0, 0, 0, 0))
```
  
```{r}
# # Saving fisheries map
# ggsave(file.path(out_plot, "fao_regions_perc_sustainable_fisheries.pdf"),
#        fao_fisheries_plot, device = "pdf", width = 14, height = 9)
```
  
## Figure 18 FAO report - FAO standard deviation
  
```{r}
#Load standard deviation
maps_data_sd <- read_csv("data/ensemble_sd_bio_change_data_map.csv")

#Extract data for scenario, transform to raster and reproject
maps_data_sd_proj <- maps_data_sd |> 
  #Selecting coordinate columns and percentage change columns
  select(x, y, ends_with("_sd")) |> 
  #Transform to multi-layer raster
  rast(type = "xyz", crs = "epsg:4326") |> 
  #Reproject to Robinson
  project(y = rob_proj) 
```
  
```{r}
#Define base steps for maps
base_gg <- list(scale_fill_distiller(palette = "Blues", direction = 1, 
                                     na.value = NA, limits = c(0, 50), 
                                     oob = oob_squish),
                geom_sf(inherit.aes = F, data = world_proj, show.legend = F,
                        color = "#dbdade", fill = "#dbdade"),
                geom_sf(inherit.aes = F, data = eez_proj, fill = NA, 
                        colour = "#5b5b5b",
                        show.legend = F, linewidth = 0.25),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = world_bound_proj, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = line_manual),
                geom_sf(inherit.aes = F, 
                        data = world_bound_proj[world_bound_proj$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5),
                      legend.position = "none"))

# #SSP1-2.6 2041-2050
# p50_126_SD <- ggplot()+ 
#   geom_spatraster(data = maps_data_sd_proj$rel_change_mean50_ssp126_sd)+
#   labs(title = "SSP1-2.6: 2041-2050")+
#   base_gg
# 
# #SSP5-8.5 2041-2050
# p50_585_SD <- ggplot()+
#   geom_spatraster(data = maps_data_sd_proj$rel_change_mean50_ssp585_sd)+
#   labs(title = "SSP5-8.5: 2041-2050")+
#   base_gg

#SSP1-2.6 2091-2100
p00_126_SD <- ggplot()+ 
  geom_spatraster(data = maps_data_sd_proj$rel_change_mean00_ssp126_sd)+
  # labs(title = "SSP1-2.6: 2091-2100")+
  base_gg+
  theme(legend.position = "none")

#SSP5-8.5 2091-2100
p00_585_SD <- ggplot()+ 
  geom_spatraster(data = maps_data_sd_proj$rel_change_mean00_ssp585_sd)+
  # labs(title = "SSP5-8.5: 2091-2100")+
  base_gg+
  guides(fill = guide_colorbar(barwidth = 1.2, 
                               barheight = 7,
                               title = "Standard \ndeviation (%)",
                               title.position = "top", title.hjust = 0.5))+
  theme(legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))

# #Get legend for land
# legend_sd <- get_legend(p00_585_SD)
# 
# #Remove legend from last plot
# p00_585_SD <- p00_585_SD+
#   theme(legend.position = "none")
# 
# #Plotting everything together
# all_plots <- plot_grid(plot_grid(p50_126_SD, p50_585_SD, ncol = 2, nrow = 1, 
#                                  labels = c("a", "b"), label_x = 0.1),
#                        plot_grid(p00_126_SD, p00_585_SD, ncol = 2, nrow = 1,
#                                  labels = c("c", "d"), label_x = 0.1),
#                        legend_sd, ncol = 1, nrow = 3, 
#                        rel_heights = c(1, 1, 0.4))
```
  
```{r}
# # Saving multi-panel plot
# ggsave(file.path(out_plot, "global_sd_map_50-00.pdf"), all_plots, 
#        device = "pdf", width = 14, height = 9)
```

## Figure FAO report - Coefficient of variation - REMOVED from report

```{r}
# #Load coefficient of variation
# maps_data_cv <- read_csv("data/ensemble_cv_bio_change_data_map.csv")
# 
# #Extract data for scenario, transform to raster and reproject
# maps_data_cv_proj <- maps_data_cv |> 
#   #Selecting coordinate columns and percentage change columns
#   select(x, y, ends_with("_cv")) |> 
#   #Transform to multi-layer raster
#   rast(type = "xyz", crs = "epsg:4326") |> 
#   #Reproject to Robinson
#   project(y = rob_proj) 
```

```{r}
# #Define base steps for maps
# base_gg <- list(scale_fill_distiller(palette = "Blues", direction = 1, 
#                                      na.value = NA, limits = c(0, 200), 
#                                      oob = oob_squish),
#                 geom_sf(inherit.aes = F, data = world_proj, show.legend = F,
#                         color = "#dbdade", fill = "#dbdade"),
#                 geom_sf(inherit.aes = F, data = eez_proj, fill = NA, 
#                         colour = "white", show.legend = F, linewidth = 0.5),
#                 #Adding coastlines and international boundaries
#                 geom_sf(inherit.aes = F, data = world_bound_proj, lwd = 0.25, 
#                         color = "#484848", show.legend = F, 
#                         aes(linetype = as.factor(TYPE))),
#                 scale_linetype_manual(values = line_manual),
#                 geom_sf(inherit.aes = F, 
#                         data = world_bound_proj[world_bound_proj$TYPE == 8,], 
#                         linetype = "longdash", lwd = 0.25, color = "#838285",
#                         show.legend = F),
#                 theme_bw(),
#                 theme(axis.title = element_blank(), 
#                       panel.border = element_rect(colour = NA),
#                       plot.title = element_text(hjust = 0.5),
#                       legend.position = "none"))
# 
# #SSP1-2.6 2041-2050
# p50_126_CV <- ggplot()+
#   geom_spatraster(data = maps_data_cv_proj$rel_change_mean50_ssp126_cv)+
#   labs(title = "SSP1-2.6: 2041-2050")+
#   base_gg
# 
# #SSP5-8.5 2041-2050
# p50_585_CV <- ggplot()+
#   geom_spatraster(data = maps_data_cv_proj$rel_change_mean50_ssp585_cv)+
#   labs(title = "SSP5-8.5: 2041-2050")+
#   base_gg
# 
# #SSP1-2.6 2091-2100
# p00_126_CV <- ggplot()+
#   geom_spatraster(data = maps_data_cv_proj$rel_change_mean00_ssp126_cv)+
#   labs(title = "SSP1-2.6: 2091-2100")+
#   base_gg
# 
# #SSP5-8.5 2091-2100
# p00_585_CV <- ggplot()+
#   geom_spatraster(data = maps_data_cv_proj$rel_change_mean00_ssp585_cv)+
#   labs(title = "SSP5-8.5: 2091-2100")+
#   base_gg+
#   guides(fill = guide_colorbar(title = "Coefficient of variation (%)",
#                                title.position = "top", title.hjust = 0.5, 
#                                barwidth = 15))+
#   theme(legend.position = "bottom")
# 
# #Get legend for land
# legend_cv <- get_legend(p00_585_CV)
# 
# #Remove legend from last plot
# p00_585_CV <- p00_585_CV+
#   theme(legend.position = "none")
# 
# #Plotting everything together
# all_plots <- plot_grid(plot_grid(p50_126_CV, p50_585_CV, ncol = 2, nrow = 1, 
#                                  labels = c("a", "b"), label_x = 0.1),
#                        plot_grid(p00_126_CV, p00_585_CV, ncol = 2, nrow = 1,
#                                  labels = c("c", "d"), label_x = 0.1),
#                        legend_cv, ncol = 1, nrow = 3, 
#                        rel_heights = c(1, 1, 0.4))
```

```{r}
# # Saving multi-panel plot
# ggsave(file.path(out_plot, "global_cv_map_50-00.pdf"), all_plots, 
#        device = "pdf", width = 14, height = 9)
```

## Figure FAO report - Model agreement 
Combining maps of percentage change (`*_change`), standard deviation (`*_SD`) and model agreement (`*_agreement`).  

### Plot % cahnge but without land data (catches)

```{r}
#Loading ensemble biomass change
maps_data <- read_csv("data/ensemble_perc_bio_change_data_map.csv")

#Extract data for scenario, transform to raster and reproject
maps_data_proj <- maps_data |> 
  #Selecting coordinate columns and percentage change columns
  select(x, y, starts_with("rel_change")) |> 
  #Transform to multi-layer raster
  rast(type = "xyz", crs = "epsg:4326") |> 
  #Reproject to Robinson
  project(y = rob_proj) 

```

```{r}

#Define base steps for maps
base_gg <- list(scale_fill_binned(limits = c(-50, 50), n.breaks = 8,
                                  type = scale_fill_custom, 
                                  oob = oob_squish),
                geom_sf(inherit.aes = F, data = eez_proj, fill = NA,
                        colour = "#5b5b5b", show.legend = F, linewidth = 0.25),
                geom_sf(inherit.aes = F, data = world_proj, show.legend = F,
                        color = "#dbdade", fill = "#dbdade"),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = world_bound_proj, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = line_manual),
                geom_sf(inherit.aes = F, 
                        data = world_bound_proj[world_bound_proj$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5)))

p00_126_change2 <- ggplot()+
  geom_spatraster(data = maps_data_proj$rel_change_mean00_ssp126_mean)+
  labs(title = "SSP1-2.6: 2091-2100")+
  base_gg+
  theme(legend.position = "none")

p00_585_change2 <- ggplot()+
  geom_spatraster(data = maps_data_proj$rel_change_mean00_ssp585_mean)+
  labs(title = "SSP5-8.5: 2091-2100")+
  base_gg+
  guides(fill = guide_colorbar(barwidth = 1.2, 
                               barheight = 7,
                               title = "Change in \nfish biomass (%)",
                               title.position = "top", title.hjust = 0.5))+
  theme(legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))

```

### Plot agreement 
  
```{r}
# Listing all relevant files to calculate biomass projections
maps_data_agreement_proj <- read_csv("data/ensemble_model_agreement.csv") |>   
  #Transform to multi-layer raster
  rast(type = "xyz", crs = "epsg:4326") |>
  #Reproject to Robinson
  project(y = rob_proj)
```

```{r}
#Define base steps for maps
base_gg <- list(scale_fill_distiller(palette = "Blues", direction = 1, 
                                     na.value = NA, limits = c(50, 100), 
                                     oob = oob_squish),
                geom_sf(inherit.aes = F, data = world_proj, show.legend = F,
                        color = "#dbdade", fill = "#dbdade"),
                geom_sf(inherit.aes = F, data = eez_proj, fill = NA, 
                        colour = "#5b5b5b",
                        show.legend = F, linewidth = 0.25),
                #Adding coastlines and international boundaries
                geom_sf(inherit.aes = F, data = world_bound_proj, lwd = 0.25, 
                        color = "#484848", show.legend = F, 
                        aes(linetype = as.factor(TYPE))),
                scale_linetype_manual(values = line_manual),
                geom_sf(inherit.aes = F, 
                        data = world_bound_proj[world_bound_proj$TYPE == 8,], 
                        linetype = "longdash", lwd = 0.25, color = "#838285",
                        show.legend = F),
                theme_bw(),
                theme(axis.title = element_blank(), 
                      panel.border = element_rect(colour = NA),
                      plot.title = element_text(hjust = 0.5)))

#SSP1-2.6 2091-2100
p00_126_agreement <- ggplot()+ 
  geom_spatraster(data = maps_data_agreement_proj$rel_change_mean00_ssp126)+
  # labs(title = "SSP1-2.6: 2091-2100")+
  base_gg+
  theme(legend.position = "none")

#SSP5-8.5 2091-2100
p00_585_agreement <- ggplot()+
  geom_spatraster(data = maps_data_agreement_proj$rel_change_mean00_ssp585)+
  # labs(title = "SSP5-8.5: 2091-2100")+
  base_gg+
  guides(fill = guide_colorbar(barwidth = 1.2, 
                               barheight = 7,
                               title = "Model \nagreement (%)",
                               title.position = "top", title.hjust = 0.5))+
  theme(legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))

# #Get legend for land
# legend_agreement <- get_legend(p00_585_agreement)
# 
# #Remove legend from last plot
# p00_585_agreement <- p00_585_agreement+
#   theme(legend.position = "none")
# 
# #Plotting change maps and legend
# p_change <- plot_grid(p00_126_change, p00_585_change, ncol = 2, 
#                       labels = c("a", "b"), label_x = 0.1)
# #Adding legend
# p_change <- plot_grid(p_change, 
#                       plot_grid(leg_land, leg_fish, ncol = 2),
#                       nrow = 2, rel_heights = c(1, 0.4))
# 
# #Plotting standard deviation maps and legend
# p_sd <- plot_grid(p00_126_SD, p00_585_SD, ncol = 2, labels = c("c", "d"), 
#                   label_x = 0.1)
# #Adding legend
# p_sd <- plot_grid(p_sd, legend_sd, nrow = 2, rel_heights = c(1, 0.4))
# 
# #Plotting agreement maps and legend
# p_agree <- plot_grid(p00_126_agreement, p00_585_agreement, ncol = 2, 
#                      labels = c("e", "f"), label_x = 0.1)
# #Adding legend
# p_agree <- plot_grid(p_agree, legend_agreement, nrow = 2, 
#                      rel_heights = c(1, 0.4))
# 
# #Putting all plots together
# all_plots <- plot_grid(p_change, p_sd, p_agree, nrow = 3)
```

### Combine all plots using new layout 

```{r}

library(patchwork)

# ## trial 
# layout <- "
# AB
# "
# 
# all_plots<-p00_126_change2 + p00_585_change2 + 
#   plot_layout(design = layout)+
#   plot_annotation(tag_levels = "a") 

layout <- "
AB
CD
EF
"

all_plots<-p00_126_change2 + p00_585_change2 + 
  p00_126_SD + p00_585_SD +  
  p00_126_agreement + p00_585_agreement +  
  plot_layout(design = layout)+
  plot_annotation(tag_levels = "a") & theme(plot.tag = element_text(size = 15, face = 'bold'))

```

```{r}
ggsave(file.path(out_plot, "agreement_map_00.pdf"), all_plots, 
       device = "pdf", width = 14, height = 13.5)
```
