---
title: "Plotting temporal biomass change"
author: "Denisse Fierro Arcos & Camilla Novaglio"
date: "2024-04-30"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

## Loading libraries
  
```{r, message=FALSE, warning=FALSE}
rm(list = ls())
library(tidyverse)
library(data.table)
```

## Setting up notebook

```{r}
#Base folder for project
base_folder <- "/rd/gem/private/users/camillan/FAO_Report/"

# Ensemble data ----

#Ensemble percentage change in biomass by countries
count_bio <- list.files(base_folder, "ensemble_perc_bio_change_country.csv",
                        recursive = T, full.names = T) |> 
  read_csv()

#Ensemble percentage change in biomass by FAO regions
fao_bio <- list.files(base_folder, "ensemble_perc_bio_change_fao_region.csv",
                        recursive = T, full.names = T) |> 
  read_csv()

#Ensemble percentage change in biomass globally
global_bio <- list.files(base_folder, "ensemble_perc_bio_change_global.csv",
                         recursive = T, full.names = T) |> 
  read_csv()

#Loading data for top 7 countries
top_7_data <- list.files(base_folder,
                         "ensemble_perc_bio_change_top7countries.csv",
                         recursive = T, full.names = T) |> 
  read_csv()

# Data by model ----

#mem percentage change in biomass by countries
count_bio_by_model <- list.files(base_folder, 
                                 "mem_perc_bio_change_country.csv",
                                 recursive = T, full.names = T) |> 
  read_csv()

#mem percentage change in biomass by FAO regions
fao_bio_by_model <- list.files(base_folder, 
                               "mem_perc_bio_change_fao_region.csv",
                               recursive = T, full.names = T) |> 
  read_csv()

#mem percentage change in biomass globally
global_bio_by_model <- list.files(base_folder, 
                                  "mem_perc_bio_change_global.csv", 
                                  recursive = T, full.names = T) |> 
  read_csv()

#mem data for top 7 countries
top_7_data_by_model <- list.files(base_folder,
                         "mem_perc_bio_change_top7countries.csv",
                         recursive = T, full.names = T) |> 
  read_csv()

```

## Checks 

```{r}

# # Check NAs by model ----
# 
# # by country - either countries on the Caspian Sea for which data is avaialble only from a few IPSL-forced models or  countries with small coasts for which data is avaialble only from a few GFDL-forced models  
# a<-filter(count_bio_by_model, is.na(mean_bio))
# unique(a$fao_official) 
# 
# # [1] "Turkmenistan"                    "the Hashemite Kingdom of Jordan" "the Kingdom of Bahrain"
# # [4] "the Republic of Albania"         "the Republic of Azerbaijan"      "the Republic of Djibouti"
# # [7] "the Republic of Iraq"            "the Republic of Kazakhstan"      "the Republic of the Sudan"
# # [10] "the State of Kuwait"             "the Kingdom of Belgium"          "the Republic of Cameroon"
# 
# a<-filter(count_bio_by_model, fao_official == "Turkmenistan", !is.na(mean_bio))
# a$id<-paste(a$esm, a$mem)
# unique(a$id) # only models with Caspian Sea
# 
# a<-filter(count_bio_by_model, fao_official == "the Hashemite Kingdom of Jordan", !is.na(mean_bio))
# a$id<-paste(a$esm, a$mem)
# unique(a$id) # only gfdl model (but boats) # # almost landlocked
# 
# a<-filter(count_bio_by_model, fao_official == "the Kingdom of Bahrain", !is.na(mean_bio))
# a$id<-paste(a$esm, a$mem)
# unique(a$id) # only gfdl model (including boats)
# 
# # for global
# a<-filter(global_bio_by_model, is.na(mean_bio))
# a$id<-paste(a$mem, a$esm)
# unique(a$id)
# 
# # for fao regions
# a<-filter(fao_bio_by_model, is.na(mean_bio))
# a$id<-paste(a$mem, a$esm)
# unique(a$id)
# 
# # for top 7 
# a<-filter(top_7_data_by_model, is.na(mean_bio))
# a$id<-paste(a$mem, a$esm)
# unique(a$id)
# 
# # CHECK correct way of calculating % changes between periods for stat table ----
# 
# # calculate percentage change between 2 periods
# # Calculating reference period
# ref_bio <- count_bio_by_model |>
#   filter(year >= 2005 & year <= 2014) |>
#   #Calculate mean per ensemble member
#   group_by(mem, esm, fao_official) |>
#   dplyr::summarise(ref_bio = mean(mean_bio, na.rm = T))
# 
# # between decades
# last_decade_bio <- count_bio_by_model |>
#   filter(year >= 2041 & year <= 2050) |>
#   #Calculate mean per ensemble member
#   group_by(mem, esm, scenario, fao_official) |>
#   dplyr::summarise(last_decade_bio = mean(mean_bio, na.rm = T))
# 
# change_between_decades<-last_decade_bio %>%
#   full_join(ref_bio) %>%
#   arrange(fao_official) %>%
#   mutate(perc_change = ((last_decade_bio-ref_bio)/ref_bio)*100)
# 
# # check differences between averaging % changes across last decade and option above - No difference
# count_bio_by_model %>% filter(year %in%(2041:2050)) %>% group_by(mem, esm, scenario, fao_official) %>% summarise(mean = mean(perc_change)) %>% arrange(fao_official)
# change_between_decades
# 
# # check the 2 approaches with fake data - no difference
# # mean biomass across ref decade for each model, assuming one country
# ref<-data.frame(model = c("mod1", "mod2"), ref_bio = c(10,12))
# # biomass per year per model
# year<-data.frame(model= c("mod1", "mod1", "mod2", "mod2"), year = c(1,2,1,2), bio = c(5,10, 11,13))
# 
# # approach 1
# # A % change between mean biomass across ref decade per model and biomass per future years per model
# all<-ref %>%
#   full_join(year) %>%
#   mutate(per_change = ((bio-ref_bio)/ref_bio)*100) %>%
#   group_by(model) %>%
#   summarise(mean_change = mean(per_change)) # B mean percentage change across years
# 
# # approach 2
# # A mean biomass across ref decade and mean biomass across future years
# all2<-ref %>%
#   full_join(year) %>%
#   group_by(model, ref_bio) %>%
#   summarise(bio = mean(bio)) %>%
#   mutate(mean_change = ((bio-ref_bio)/ref_bio)*100) # B % change across decades

```

## Table of stats

```{r}

count_bio_by_model<-count_bio_by_model %>% 
  mutate(spatial_scale = "countries")

global_bio_by_model<-global_bio_by_model %>% 
  mutate(spatial_scale = "global", 
         figure_name = "global")

top_7_data_by_model<-top_7_data_by_model %>% 
  mutate(spatial_scale = "top_7_countries") %>% 
  rename(figure_name = Country_FAO) %>% # can use short names also for countries
  mutate(figure_name = case_when(figure_name == "the People's Republic of China" ~ "China",
                                  figure_name == "the Republic of India" ~ "India",
                                  figure_name == "the Republic of Indonesia" ~ "Indonesia",
                                  figure_name == "the Republic of Peru" ~ "Peru",
                                  figure_name == "the Russian Federation" ~ "Russian Federation",
                                  figure_name == "the Socialist Republic of Viet Nam" ~ "Viet Nam",
                                  figure_name == "the United States of America" ~ "United States of America", 
                                  .default = as.character(figure_name)))

fao_bio_by_model<-fao_bio_by_model %>% 
  mutate(spatial_scale = "FAO_area") %>% 
  rename(figure_name = NAME_EN)

all_scale_data_by_model<-count_bio_by_model %>% 
  full_join(global_bio_by_model) %>% 
  full_join(top_7_data_by_model) %>% 
  full_join(fao_bio_by_model) %>% 
  filter(year %in% c(2041:2050, 2091:2100)) %>% 
  mutate(decade = ifelse(year %in% c(2041:2050), "2041-2050", "2091-2100")) %>% 
  filter(!is.na(perc_change)) 

### mean, median, sd, min and max, agreement, wilcox ----
stats<-all_scale_data_by_model %>% 
  group_by(figure_name, mem, esm, scenario, decade, continent, spatial_scale) %>% 
  summarise(mean_across_decade = mean(perc_change, na.rm = T)) %>%  # across decade to have one (mean) value of change per model and exclude temporal variability from the calculations below
  ungroup() %>% 
  group_by(figure_name, scenario, decade, continent, spatial_scale) %>% 
  summarise(
    mean = mean(mean_across_decade), # then across model 
    median = median(mean_across_decade),
    sd = sd(mean_across_decade),
    min = min(mean_across_decade),
    max = max(mean_across_decade),
    n_model = n(), # to calculate agreement across models ### OR n_model = sum(!is.na(value)) as for maps??? 
    n_positive = sum(mean_across_decade >= 0), # non-negative, 0 included as per Tittensor. No difference as no true 0 in data.  #### OR na.rm = T as for maps ?? 
    n_negative = sum(mean_across_decade < 0), #### OR na.rm = T as for maps ?? 
    same_sign = max(n_positive,n_negative),
    agreement = (same_sign/n_model)*100) %>% 
  ungroup() %>% 
  select(-c(n_positive, n_negative, same_sign))

# # test on model agreement - OK
# data.frame(n_model = 10,
#            n_positive = 5,
#            n_negative = 5) %>%
#   mutate(same_sign = max(n_positive, n_negative),
#          agreement = (same_sign/n_model)*100)
# 
# # check by plotting top 7 countries - OK
# to_plot<-top_7_data_by_model %>%
#   filter(year %in% c(2041:2050),
#          scenario == "ssp126") %>%
#   mutate(id = paste(esm, mem, sep ="_")) %>% 
#   group_by(mem, esm, scenario, figure_name) %>% 
#   summarise(perc_change = mean(perc_change)) %>% 
#   ungroup()
# 
# unique(stats$spatial_scale)
# filter(stats, decade == "2041-2050", scenario == "ssp126", spatial_scale == "top_7_countries")
# 
# ggplot(to_plot, aes(x = mem, y = perc_change))+
#   # geom_line()+
#   geom_point()+
#   geom_hline(yintercept = 0, color = "black")+
#   facet_wrap(~figure_name)

#### wilcox test ----
# http://www.sthda.com/english/wiki/unpaired-two-samples-wilcoxon-test-in-r
# from example: The p-value of the test is 0.02712, which is less than the significance level alpha = 0.05. We can conclude that men’s median weight is significantly different from women’s median weight with a p-value = 0.02712.

w_test<-all_scale_data_by_model %>% 
  select(!c(mean_bio, ref_bio)) %>% 
  spread(scenario, perc_change) %>% 
  group_by(figure_name, decade, spatial_scale) %>% 
  summarise(wilcox_p = wilcox.test(ssp126, ssp585, alternative = "two.sided")$p.value) %>% 
  ungroup()

# # check warnings
# dplyr::last_dplyr_warnings() # warning for 4 countries

# filter(w_test, figure_name == "Bahrain", decade == "2041-2050", spatial_scale == "countries") # values are provided but might not be trustable is my interpretation.
#
# # check single test
# w_test_warnings<-all_scale_data_by_model %>%
#   select(!c(mean_bio, ref_bio)) %>% 
#   spread(scenario, perc_change) %>%
#   filter(figure_name == "Bahrain",
#          decade == "2041-2050",
#          spatial_scale == "countries")
# 
# wilcox.test(w_test_warnings$ssp126, w_test_warnings$ssp585, alternative = "two.sided")$p.value
# # Warning message:
# # In wilcox.test.default(w_test_warnings$ssp126, w_test_warnings$ssp585,  :
# #   cannot compute exact p-value with ties
# # where ties is when both set have the same value.
#
# # NAs non-influent
# a<-c(1,4,6,NA, 2,7,NA,8,9,20,NA)
# b<-c(NA,3,26, NA,28,40,33,NA,44,55,80)
# wilcox.test(a, b, alternative = "two.sided")$p.value
# 
# # check test
# # check if values from the table are the same as the below - yes
# filter(w_test, figure_name == "Viet Nam")
# 
# # test to understand wilcox
# # between scenarios in 2050 (test not significant)
# trial<-top_7_data_by_model %>%
#   filter(figure_name == "Viet Nam",
#          year >= 2041 & year <= 2050)
# 
# # boxplot
# ggplot(trial, aes(x=scenario, y=perc_change)) +
#     geom_boxplot() +
#     geom_jitter(color="black", size=0.4, alpha=0.9)
# 
# trial_test<-trial %>%
#   spread(scenario, perc_change)
# 
# wilcox.test(trial_test$ssp126, trial_test$ssp585, alternative = "two.sided")$p.value # 0.0590907 not significant
# 
# # between scenarios in 2100 (test significant)
# trial<-top_7_data_by_model %>%
#   filter(figure_name == "Viet Nam",
#          year >= 2091 & year <= 2100)
# 
# # boxplot
# ggplot(trial, aes(x=scenario, y=perc_change)) +
#     geom_boxplot() +
#     geom_jitter(color="black", size=0.4, alpha=0.9)
# 
# trial_test<-trial %>%
#   spread(scenario, perc_change)
# 
# wilcox.test(trial_test$ssp126, trial_test$ssp585, alternative = "two.sided")$p.value # 1.967129e-20 with raw values this is highly significant although sd of values overlap in trend figure and values overlap in boxplot.

### losses ----

losses<-stats %>% 
  select(!c(median, sd, min, max, n_model, agreement)) %>% 
  mutate(mean = round(mean, 1)) %>% # calculate with one decimal values. Results slightly change and should be consistent with the ones people can caclaulte using mean in the final table that are roundedd to one decimal. 
  pivot_wider(names_from = scenario, values_from = mean) %>% 
  group_by(figure_name, decade, spatial_scale) %>% 
  summarise(losses = (1-(ssp126/ssp585))*100,
            losses_direction = ifelse(ssp126 > ssp585, "reduction of losses under SSP1-2.6", "increase of losses under SSP1-2.6")) %>% 
  ungroup() %>%
  filter(decade == "2091-2100")

# # check losses - added column on direction (i.e. reduced or increased losses under ssp1) to understand signs             
# filter(stats, figure_name == "Algeria")  
# (1-(-4.5/-0.722))*100

# #### mitigation higher than high emission (they should all be negative) 
# a<-7 # mitigation 7% increase
# b<-3 # high emission 3% increase
# (1-(a/b))*100 # -133% # high emission results in more losses 
# 
# a<--1 
# b<--3 
# (1-(a/b))*100 # 66 # this should be negative 
# 
# a<-3 
# b<--1 
# (1-(a/b))*100 # 400 # this should be negative 
# 
# #### mitigation lower than high emission 
# a<-3 # mitigation 7% increase
# b<-7 # high emission 3% increase
# (1-(a/b))*100 # 57% # high emission results in less losses 
# 
# a<--3 
# b<-1 
# (1-(a/b))*100 # 400 
# 
# a<--3 
# b<--1 
# (1-(a/b))*100 # -200 # this should be positive 
# 
# # e.g global 
# a<--7
# b<--24
# (1-(a/b))*100 # 70

# ### historical sd REMOVE? ----
# 
# historical_sd<-count_bio_by_model %>% 
#   full_join(global_bio_by_model) %>% 
#   full_join(top_7_data_by_model) %>% 
#   full_join(fao_bio_by_model) %>% 
#   filter(year %in% c(2005:2014), !is.na(perc_change)) %>% 
#   group_by(figure_name, mem, esm, continent, spatial_scale) %>% 
#   summarise(mean_across_decade = mean(perc_change, na.rm = T)) %>%  # across decade to have one (mean) value of change per model and exclude temporal variability from the calculations below
#   ungroup() %>% 
#   group_by(figure_name, continent, spatial_scale) %>% 
#   summarise(
#     historical_sd = sd(mean_across_decade)) %>% 
#   ungroup() 

# # Check historical sd 
# filter(historical_sd, figure_name == "Viet Nam", spatial_scale == "top_7_countries")
# 
# # check VietNam 
# trial<-count_bio_by_model %>% 
#   full_join(global_bio_by_model) %>% 
#   full_join(top_7_data_by_model) %>% 
#   full_join(fao_bio_by_model) %>% 
#   filter(figure_name == "Viet Nam", spatial_scale == "top_7_countries") %>% 
#   select(!c(continent, fao_official, figure_name, spatial_scale))
# 
# to_plot<-trial %>% 
#   mutate(id = paste(mem, esm))
# 
# ggplot(filter(to_plot, scenario !="ssp126"), aes(x = year, y = perc_change, group = id, color = id))+
#   geom_line()
# 
# # calcualte sd 
# sd<-to_plot %>% 
#   mutate(decade = case_when(year %in% c(2005:2014) ~ "2005-2014",
#                             year %in% c(2041:2050) ~ "2041-2050"
#                             year %in% c(2091:2100) ~ "2091-2100",
#                             .default = NA)) %>% 
#   filter(!is.na(decade)) 
# 
# ggplot(filter(sd, scenario == "historical"), aes(x = year, y = mean, group = scenario, color = scenario))+
#   geom_line()+
#   geom_ribbon(aes(ymin = mean-sd, ymax = mean+sd, fill = scenario), alpha = 0.3, color = NA)
# 
# 
# sd<-sd %>% 
#   group_by(decade, scenario, mem, esm) %>% 
#   summarise(mean_across_decade = mean(perc_change)) %>% 
#   ungroup() %>% 
#   group_by(scenario) %>% 
#   summarise(mean = mean(mean_across_decade), 
#             sd = sd(mean_across_decade)) %>% 
#   ungroup()
#  
# sd
# 
# ggplot(to_plot2, aes(x = year, y = mean, group = scenario, color = scenario))+
#   geom_line()+
#   geom_ribbon(aes(ymin = mean-sd, ymax = mean+sd, fill = scenario), alpha = 0.3, color = NA)

#### combine tables ----
table<-stats %>% 
  full_join(w_test) %>% 
  full_join(losses) 

# # print tables 
fwrite(table, paste0(base_folder,"table_stats.csv")) 

# ##### check that tables are the same -----
# old<-read_csv("data/table_stats_with_continent.csv") # older version - saved with different name in google drive, then uploaded from Julia after she calculated some stat and worked on it in excel.
# new<-read_csv(file.path(base_folder,"table_stats.csv"))
# 
# old<-old %>%
#   arrange(fa_ffcl_shrt, scenario, decade, contnnt, spatial_scale)
# new<-new %>%
#   arrange(figure_name, scenario, decade, continent, spatial_scale)
# 
# # check countries included - only difference is Bouvet Island - to add and Colombia (Serranilla) to remove. Colombia (Serranilla) not even in the formatted table below so was deleted in previous rounds.
# setdiff(sort(unique(old$fa_ffcl_shrt)), sort(unique(new$figure_name))) # plus russia that should have been russian federation
# setdiff(sort(unique(new$figure_name)), sort(unique(old$fa_ffcl_shrt)))
# 
# # difference in values
# old<-old %>%
#   filter(!fa_ffcl_shrt == "Colombia (Serranilla)") %>%
#   mutate(fa_ffcl_shrt = case_when(fa_ffcl_shrt == "C√¥te d'Ivoire" ~ "Côte d'Ivoire",
#                                   fa_ffcl_shrt == "Cura√ßao" ~ "Curaçao",
#                                   fa_ffcl_shrt == "R√©union" ~ "Réunion",
#                                   fa_ffcl_shrt == "Russia" ~ "Russian Federation",
#                                   fa_ffcl_shrt == "T√ºrkiye" ~ "Türkiye",
#                                   .default = as.character(fa_ffcl_shrt)))
# new<-new %>%
#   filter(!figure_name == "Bouvet Island")
# 
# old<-old %>%
#   arrange(fa_ffcl_shrt, scenario, decade, contnnt, spatial_scale) %>%
#   select(fa_ffcl_shrt, scenario, decade, contnnt, spatial_scale, mean)
# 
# new<-new %>%
#   arrange(figure_name, scenario, decade, continent, spatial_scale) %>%
#   rename(mean_new = mean, fa_ffcl_shrt = figure_name, contnnt = continent) %>%
#   select(fa_ffcl_shrt, scenario, decade, contnnt, spatial_scale, mean_new)
# 
# comapre_values<- old %>%
#   full_join(new) %>%
#   mutate(mean = round(mean, 1),
#          mean_new = round(mean_new, 1)) %>%
#   mutate(compare = ifelse(mean == mean_new, "ok", "wrong")) %>%
#   filter(compare == "wrong") %>%
#   mutate(diff = abs(abs(mean) - abs(mean_new))) %>%
#   arrange(-diff)
# 
# View(comapre_values)
# very different = Canary Island, Mauritania, Puerto Rico, Romania, Turkia
# different for some scenarios ~0.5 = Comoros, Cuba, Fiji

# # check Romania - all models present and none showing huge changes fro ssp126 as previously done.
# rom<-count_bio_by_model %>%
#   filter(figure_name == "Romania") %>%
#   mutate(id = paste(esm, mem, sep = "-"))
#
# ggplot(rom, aes(x = year, y = perc_change, group = id, color = id))+
#   geom_line()+
#   facet_wrap(~scenario)
#
# rom<-rom %>%
#   group_by(scenario, year) %>%
#   summarise(mean = mean(perc_change),
#             sd = sd(perc_change))
#
# ggplot(rom, aes(x = year, y = mean, group = scenario, color = scenario))+
#   geom_line()+
#   geom_ribbon(aes(ymin = mean-sd,
#                   ymax = mean+sd,
#                   fill = scenario), alpha = 0.3)
#
# # print the diffences and update the report.
# fwrite(comapre_values, "data/data_products/table_stats_differences.csv")

##### format table for report ----

table_formatted<-table %>% 
  mutate(id = paste(scenario, decade, sep = "_")) %>% 
  select(!c(scenario, decade, median, min, max, n_model)) %>% 
  mutate(mean = round(mean, 1), 
         sd = round(sd, 1),
         agreement = round(agreement),
         wilcox_p = ifelse(wilcox_p < 0.05, "*", NA), 
         losses = round(losses, 1)) %>% 
  pivot_wider(names_from = id, values_from = c(mean, sd, agreement, wilcox_p, losses, losses_direction)) %>%
  relocate(spatial_scale, continent, figure_name,
           `mean_ssp126_2041-2050`,`sd_ssp126_2041-2050`,`agreement_ssp126_2041-2050`,`wilcox_p_ssp126_2041-2050`,
           `mean_ssp585_2041-2050`,`sd_ssp585_2041-2050`,`agreement_ssp585_2041-2050`,`wilcox_p_ssp585_2041-2050`,
           `mean_ssp126_2091-2100`,`sd_ssp126_2091-2100`,`agreement_ssp126_2091-2100`,`wilcox_p_ssp126_2091-2100`,`losses_ssp126_2091-2100`,`losses_direction_ssp126_2091-2100`,
           `mean_ssp585_2091-2100`,`sd_ssp585_2091-2100`,`agreement_ssp585_2091-2100`,`wilcox_p_ssp585_2091-2100`,`losses_ssp585_2091-2100`,`losses_direction_ssp585_2091-2100`)  %>% 
  select(!c(`wilcox_p_ssp126_2041-2050`,`wilcox_p_ssp126_2091-2100`,
            `losses_ssp126_2041-2050`,`losses_ssp126_2091-2100`,
            `losses_direction_ssp126_2041-2050`,`losses_direction_ssp126_2091-2100`)) %>% # wilcox and losses are the same across scenarios but different across decades so i keep only ssp585 for both decades. historical_sd is unique across scenarios and decade, so keep only one
  rename(`wolcox_2041-2050` = `wilcox_p_ssp585_2041-2050`, `wolcox_2091-2100` =`wilcox_p_ssp585_2091-2100`, `losses_2091-2100` =`losses_ssp585_2091-2100`, `losses_direction_2091-2100` = `losses_direction_ssp585_2091-2100`) %>% 
  mutate(continent = case_when(spatial_scale == "top_7_countries" & figure_name == "China" ~ "Asia",
                               spatial_scale == "top_7_countries" & figure_name == "India" ~ "Asia",
                               spatial_scale == "top_7_countries" & figure_name == "Indonesia" ~ "Asia",
                               spatial_scale == "top_7_countries" & figure_name == "Peru" ~ "Americas",
                               spatial_scale == "top_7_countries" & figure_name == "Russian Federation" ~ "Europe",
                               spatial_scale == "top_7_countries" & figure_name == "United States of America" ~ "Americas",
                               spatial_scale == "top_7_countries" & figure_name == "Viet Nam" ~ "Asia",
                               .default = as.character(continent))) %>% 
  mutate(spatial_scale = factor(spatial_scale, levels = c("global", "top_7_countries", "countries", "FAO_area"))) %>% 
  arrange(spatial_scale, continent, figure_name) 

fwrite(table_formatted, paste0(base_folder,"table_stats_formatted.csv")) 

```

## Defining basic plot design
  
```{r}
base_gg <- list(geom_line(linewidth = 0.5),
                #Adding no change line at 0 for reference 
                geom_hline(yintercept = 0, color = "grey80", linewidth = 0.65, 
                           linetype = 2),
                #Adding line dividing historical period and future projections
                geom_vline(xintercept = 2015, color = "grey80", 
                           linewidth = 0.65),
                #Adding SD as shading 
                geom_ribbon(aes(ymin = mean_change-sd_change,
                                ymax = mean_change+sd_change, fill = scenario),
                            alpha = 0.3, color = NA),
                #Manually setting colours to be used in plots
                scale_color_manual(values = c("historical" = "black",
                                              "ssp126" = "#33bbee", 
                                              "ssp585" = "#ee3377"), 
                                   name = "Scenarios",
                                   labels = c("Historical", "SSP1-2.6", 
                                              "SSP5-8.5")),
                scale_fill_manual(values = c("historical" = "black", 
                                             "ssp126" = "#33bbee", 
                                             "ssp585" = "#ee3377"), 
                                  name = "Scenarios",
                                  labels = c("Historical", "SSP1-2.6", 
                                             "SSP5-8.5")),
                guides(color = guide_legend(nrow = 1, title.position = "left")),
                theme_classic(),
                labs(y = "Change in exploitable\nfish biomass (%)"),
                theme(legend.position = "top", legend.justification = "center", 
                      legend.text = element_text(size = 10),
                      panel.grid.minor.y = element_blank(), 
                      axis.title.x = element_blank(),
                      axis.title.y = element_text(size = 12),
                      axis.text.x = element_text(angle = 45, vjust = 0.765, 
                                                 hjust = 0.65)))
```

## Trends globally and for top 7 countries with highest catches
  
```{r}
#Merging global and top 7 countries data

top_7_data<-top_7_data |>
  mutate(Country_FAO = case_when(Country_FAO == "the People's Republic of China" ~ "China",
                             Country_FAO == "the Republic of India" ~ "India",
                             Country_FAO == "the Republic of Indonesia" ~ "Indonesia",
                             Country_FAO == "the Republic of Peru" ~ "Peru",
                             Country_FAO == "the Russian Federation" ~ "Russian Federation",
                             Country_FAO == "the Socialist Republic of Viet Nam" ~ "Viet Nam",
                             Country_FAO == "the United States of America" ~ "United States of America",
                             .default = as.character(Country_FAO)))
  
top_7<-global_bio |> 
  #Add an extra column to match top 7 countries
  mutate(Country_FAO = "Global", .after = "scenario") |> 
  #Joining to top7 data
  bind_rows(top_7_data) |> 
  #Create an order category for SOVEREIGN1
  mutate(Country_FAO = factor(Country_FAO, levels = c("Global", 
                                              unique(top_7_data$Country_FAO)), 
                             ordered = T)) |> 
  #Plotting
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  #Applying base plot design
  base_gg+
  #Plotting by country
  facet_wrap(~Country_FAO, scales = "free_y", nrow = 4)+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 11, face = "bold"))+ 
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
```{r, eval = F}
pdf(file.path(base_folder, "top_7_global_trends.pdf"), width=9, height= 10)
top_7
dev.off()
```

## Plotting trends for single country or the globe - REMOVED from report

```{r}
# #Selecting country from trends per country variable
# plot_country <- count_bio |> 
#   filter(fao_official == "Australia") |> 
#   #Plot
#   ggplot(aes(x = year, y = mean_change, colour = scenario))+
#   #Apply base plotting scheme
#   base_gg
# 
# #Saving output
# png("data/data_products/Au_trends.png", width = 5, height = 3.5, units = "in", res = 1200)
# #Check country plot
# plot_country
# #Remove figure from memory
# dev.off()
```

## Figure 2 - Asia
In this figure, we will exclude marine areas under the jurisdiction of countries in Asia as classified by the FAO.  
  
```{r}

#Filtering data
asia<-count_bio |> 
  #Extract EEZs in Asia
  filter(continent == "Asia") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
```{r, eval = F}
pdf(file.path(base_folder, "asia_perc_change.pdf"), width=9, height= 10)
asia
dev.off()
```
  
## Figure 3 - Africa
In this figure, we will exclude marine areas under the jurisdiction of countries in Africa as classified by the FAO.  
  
```{r}
#Filtering data
africa<-count_bio |> 
  #Extract EEZs in Africa
  filter(continent == "Africa") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5, 
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
```{r, eval = F}
pdf(file.path(base_folder, "africa_perc_change.pdf"), width=9, height= 12)
africa
dev.off()
```
  
## Figure 4 - Europe
In this figure, we will exclude marine areas under the jurisdiction of countries in Europe as classified by the FAO.  
  
```{r}
#Filtering data

sort(unique(count_bio$figure_name))

count_bio_europe<-count_bio %>% 
  mutate(figure_name = ifelse(figure_name == "United Kingdom of Great Britain and Northern Ireland", "United Kingdom" , figure_name))

europe<-count_bio_europe |> 
  #Extract EEZs in Europe
  filter(continent == "Europe") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
```{r, eval = F}
pdf(file.path(base_folder, "europe_perc_change.pdf"), width=9, height= 9)
europe
dev.off()
```
  
## Figure 5 - Americas
In this figure, we will exclude marine areas under the jurisdiction of countries in the Americas as classified by the FAO.  
  
```{r}
#Filtering data
americas<-count_bio |> 
  #Extract EEZs in the Americas
  filter(continent == "Americas") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
   theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
```{r, eval = F}
pdf(file.path(base_folder, "americas_perc_change.pdf"), width=9, height= 13)
americas
dev.off()
```
  
## Figure 6 - Oceania
In this figure, we will exclude marine areas under the jurisdiction of countries in Oceania as classified by the FAO.  
  
```{r}
#Filtering data
oceania<-count_bio |> 
  #Extract EEZs in Oceania
  filter(continent == "Oceania") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
```{r, eval = F}
pdf(file.path(base_folder, "oceania_perc_change.pdf"), width=9, height= 9)
oceania
dev.off()
```
  
## Figure 7 - FAO regions
  
```{r}
#Filtering data
fao_regions<-fao_bio |> 
  #Plotting data
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by FAO sector
  facet_wrap(~NAME_EN, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(NAME_EN = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
```{r, eval = F}
pdf(file.path(base_folder, "fao_regions_perc_change.pdf"), width=9, height= 6)
fao_regions
dev.off()
```
  
## Uncertaity plot - REMOVED from report

```{r}

# # use mean and sd calculated above 
# 
# # sort values
# sorting<-table %>% 
#   filter(spatial_scale == "FAO_area") %>% 
#   filter(scenario == "ssp585", decade == "2091-2100") %>% 
#   arrange(-mean)
# 
# sorting<-unique(sorting$figure_name)
# sorting<-factor(sorting, levels=unique(sorting))
# 
# bio_dataOCD <-table %>% 
#   filter(spatial_scale == "FAO_area") %>% 
#   mutate(figure_name = factor(figure_name, levels = sorting))
# 
# uncert_decade<-ggplot(filter(bio_dataOCD, figure_name != "Arctic Sea"), aes(x=figure_name, y=mean, group=scenario, color=scenario)) + 
#   geom_point()+ 
#   geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
#                 position=position_dodge(0.05))+
#   theme_classic() +
#   geom_hline(yintercept = 0, linetype = "dashed", color = "grey80", linewidth = 1)+
#   scale_color_manual(name = "Scenario", labels = c("SSP1-2.6", "SSP5-8.5"), values=c("#33bbee", "#ee3377"))+
#   labs(x="FAO Major Fishing Area", y = "Change in exploitable fish biomass (%)") +
#   coord_flip()+
#   theme(axis.text=element_text(size=14),
#         legend.title = element_text(size = 14),
#         legend.text = element_text(size = 14),
#         axis.title=element_text(size=14)) + # face = "bold"
#   facet_wrap(~decade)
# 
# png("data/data_products/FAOarea_uncertity.png", width=10, height=5, units="in",res=1200)
# uncert_decade
# dev.off()

```
  