---
title: "Plotting temporal biomass change"
author: "Denisse Fierro Arcos & Camilla Novaglio"
date: "2024-04-30"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

## Loading libraries
  
```{r, message=FALSE, warning=FALSE}
rm(list = ls())
library(tidyverse)
library(data.table)
```

## Setting up notebook

```{r}
#Base folder for project
base_folder <- "/rd/gem/private/users/camillan/FAO_Report/"

# Ensemble data ----

#Ensemble percentage change in biomass by countries
count_bio <- list.files(base_folder, "ensemble_perc_bio_change_country.csv",
                        recursive = T, full.names = T) |> 
  read_csv()

### correct territory name: 
count_bio<-count_bio %>% 
  mutate(figure_name = ifelse(figure_name == "Prince Edward Island","Prince Edward Islands",figure_name))

#Ensemble percentage change in biomass by FAO regions
fao_bio <- list.files(base_folder, "ensemble_perc_bio_change_fao_region.csv",
                        recursive = T, full.names = T) |> 
  read_csv()

#Ensemble percentage change in biomass globally
global_bio <- list.files(base_folder, "ensemble_perc_bio_change_global.csv",
                         recursive = T, full.names = T) |> 
  read_csv()

#Loading data for top 7 countries
top_7_data <- list.files(base_folder,
                         "ensemble_perc_bio_change_top7countries.csv",
                         recursive = T, full.names = T) |> 
  read_csv()

# Data by model ----

#mem percentage change in biomass by countries
count_bio_by_model <- list.files(base_folder, 
                                 "mem_perc_bio_change_country.csv",
                                 recursive = T, full.names = T) |> 
  read_csv()

### correct territory name: 
count_bio_by_model<-count_bio_by_model %>% 
  mutate(figure_name = ifelse(figure_name == "Prince Edward Island","Prince Edward Islands",figure_name))

#mem percentage change in biomass by FAO regions
fao_bio_by_model <- list.files(base_folder, 
                               "mem_perc_bio_change_fao_region.csv",
                               recursive = T, full.names = T) |> 
  read_csv()

#mem percentage change in biomass globally
global_bio_by_model <- list.files(base_folder, 
                                  "mem_perc_bio_change_global.csv", 
                                  recursive = T, full.names = T) |> 
  read_csv()

#mem data for top 7 countries
top_7_data_by_model <- list.files(base_folder,
                         "mem_perc_bio_change_top7countries.csv",
                         recursive = T, full.names = T) |> 
  read_csv()

```

## Table of stats

```{r}

count_bio_by_model<-count_bio_by_model %>% 
  mutate(spatial_scale = "countries")

global_bio_by_model<-global_bio_by_model %>% 
  mutate(spatial_scale = "global", 
         figure_name = "global")

top_7_data_by_model<-top_7_data_by_model %>% 
  mutate(spatial_scale = "top_7_countries") %>% 
  rename(figure_name = Country_FAO) %>% # can use short names also for countries
  mutate(figure_name = case_when(figure_name == "the People's Republic of China" ~ "China",
                                  figure_name == "the Republic of India" ~ "India",
                                  figure_name == "the Republic of Indonesia" ~ "Indonesia",
                                  figure_name == "the Republic of Peru" ~ "Peru",
                                  figure_name == "the Russian Federation" ~ "Russian Federation",
                                  figure_name == "the Socialist Republic of Viet Nam" ~ "Viet Nam",
                                  figure_name == "the United States of America" ~ "United States of America", 
                                  .default = as.character(figure_name)))

fao_bio_by_model<-fao_bio_by_model %>% 
  mutate(spatial_scale = "FAO_area") %>% 
  rename(figure_name = NAME_EN)

all_scale_data_by_model<-count_bio_by_model %>% 
  full_join(global_bio_by_model) %>% 
  full_join(top_7_data_by_model) %>% 
  full_join(fao_bio_by_model) %>% 
  filter(year %in% c(2041:2050, 2091:2100)) %>% 
  mutate(decade = ifelse(year %in% c(2041:2050), "2041-2050", "2091-2100")) %>% 
  filter(!is.na(perc_change)) 

### mean, median, sd, min and max, agreement, wilcox ----
stats<-all_scale_data_by_model %>% 
  group_by(figure_name, mem, esm, scenario, decade, continent, spatial_scale) %>% 
  summarise(mean_across_decade = mean(perc_change, na.rm = T)) %>%  # across decade to have one (mean) value of change per model and exclude temporal variability from the calculations below
  ungroup() %>%
  group_by(figure_name, scenario, decade, continent, spatial_scale) %>% 
  summarise(
    mean = mean(mean_across_decade), # then across model 
    median = median(mean_across_decade),
    sd = sd(mean_across_decade),
    min = min(mean_across_decade),
    max = max(mean_across_decade),
    n_model = n(), # to calculate agreement across models (i.e. number of rows)
    n_positive = sum(mean_across_decade >= 0), # non-negative, 0 included as per Tittensor. (there are no NAs in mean_across_decade) 
    n_negative = sum(mean_across_decade < 0), 
    same_sign = max(n_positive,n_negative),
    agreement = (same_sign/n_model)*100) %>% 
  ungroup() %>% 
  select(-c(n_positive, n_negative, same_sign))

#### wilcox test ----
# http://www.sthda.com/english/wiki/unpaired-two-samples-wilcoxon-test-in-r
# from example: The p-value of the test is 0.02712, which is less than the significance level alpha = 0.05. We can conclude that men’s median weight is significantly different from women’s median weight with a p-value = 0.02712.

w_test<-all_scale_data_by_model %>% 
  select(!c(mean_bio, ref_bio)) %>% 
  spread(scenario, perc_change) %>% 
  group_by(figure_name, decade, spatial_scale) %>% 
  summarise(wilcox_p = wilcox.test(ssp126, ssp585, alternative = "two.sided")$p.value) %>% 
  ungroup()

## Warnings on Bahrain both decades, Djibouti 2041-205, Kuwait 2041-2051. These countries/decades will not be considered significant in final table because of warning: "cannot compute exact p-value with ties". 

### losses ----

losses<-stats %>% 
  select(!c(median, sd, min, max, n_model, agreement)) %>% 
  mutate(mean = round(mean, 1)) %>% # calculate with one decimal values. Results are sensitive to decimal places, these will be consistent with the ones that can be calcualted from the final, published table (where mean is rounded to 1 decimal) 
  pivot_wider(names_from = scenario, values_from = mean) %>% 
  group_by(figure_name, decade, spatial_scale) %>% 
  summarise(losses = (1-(ssp126/ssp585))*100,
            losses_direction = ifelse(ssp126 > ssp585, "reduction of losses under SSP1-2.6", "increase of losses under SSP1-2.6")) %>% 
  ungroup() %>%
  filter(decade == "2091-2100")

#### combine tables ----
table<-stats %>% 
  full_join(w_test) %>% 
  full_join(losses) 

### print tables ----
fwrite(table, paste0(base_folder,"table_stats.csv")) 

##### format table A4 for report ----

table_formatted<-table %>% 
  mutate(id = paste(scenario, decade, sep = "_")) %>% 
  select(!c(scenario, decade, median, min, max, n_model)) %>% 
  mutate(mean = round(mean, 1), 
         sd = round(sd, 1),
         agreement = round(agreement),
         wilcox_p = ifelse(wilcox_p < 0.05, "*", NA), 
         losses = round(losses, 1)) %>% 
  pivot_wider(names_from = id, values_from = c(mean, sd, agreement, wilcox_p, losses, losses_direction)) %>%
  select(!c(`wilcox_p_ssp126_2041-2050`,`wilcox_p_ssp126_2091-2100`,
            `losses_ssp126_2041-2050`,`losses_ssp126_2091-2100`,
            `losses_ssp585_2041-2050`,`losses_direction_ssp126_2041-2050`,`losses_direction_ssp126_2091-2100`,`losses_direction_ssp585_2041-2050`)) %>% # wilcox and losses are the same across scenarios but different across decades so i keep only ssp585 for both decades. I also delete losses in 2041 as not considered in text.  
  relocate(spatial_scale, continent, figure_name,
           `mean_ssp126_2041-2050`,`sd_ssp126_2041-2050`,`agreement_ssp126_2041-2050`,
           `mean_ssp585_2041-2050`,`sd_ssp585_2041-2050`,`agreement_ssp585_2041-2050`,`wilcox_p_ssp585_2041-2050`,
           `mean_ssp126_2091-2100`,`sd_ssp126_2091-2100`,`agreement_ssp126_2091-2100`,
           `mean_ssp585_2091-2100`,`sd_ssp585_2091-2100`,`agreement_ssp585_2091-2100`,`wilcox_p_ssp585_2091-2100`,`losses_ssp585_2091-2100`,`losses_direction_ssp585_2091-2100`)  %>% 
  rename(`wolcox_2041-2050` = `wilcox_p_ssp585_2041-2050`, `wolcox_2091-2100` =`wilcox_p_ssp585_2091-2100`, `losses_2091-2100` =`losses_ssp585_2091-2100`, `losses_direction_2091-2100` = `losses_direction_ssp585_2091-2100`) %>% 
  mutate(continent = case_when(spatial_scale == "top_7_countries" & figure_name == "China" ~ "Asia",
                               spatial_scale == "top_7_countries" & figure_name == "India" ~ "Asia",
                               spatial_scale == "top_7_countries" & figure_name == "Indonesia" ~ "Asia",
                               spatial_scale == "top_7_countries" & figure_name == "Peru" ~ "Americas",
                               spatial_scale == "top_7_countries" & figure_name == "Russian Federation" ~ "Europe",
                               spatial_scale == "top_7_countries" & figure_name == "United States of America" ~ "Americas",
                               spatial_scale == "top_7_countries" & figure_name == "Viet Nam" ~ "Asia",
                               .default = as.character(continent))) %>% 
  mutate(spatial_scale = factor(spatial_scale, levels = c("global", "top_7_countries", "countries", "FAO_area"))) %>% 
  arrange(spatial_scale, continent, figure_name) 

fwrite(table_formatted, paste0(base_folder,"table_stats_formatted.csv")) 

```

## Defining basic plot design
  
```{r}
base_gg <- list(geom_line(linewidth = 0.5),
                #Adding no change line at 0 for reference 
                geom_hline(yintercept = 0, color = "grey80", linewidth = 0.65, 
                           linetype = 2),
                #Adding line dividing historical period and future projections
                geom_vline(xintercept = 2015, color = "grey80", 
                           linewidth = 0.65),
                #Adding SD as shading 
                geom_ribbon(aes(ymin = mean_change-sd_change,
                                ymax = mean_change+sd_change, fill = scenario),
                            alpha = 0.3, color = NA),
                #Manually setting colours to be used in plots
                scale_color_manual(values = c("historical" = "black",
                                              "ssp126" = "#33bbee", 
                                              "ssp585" = "#ee3377"), 
                                   name = "Scenarios",
                                   labels = c("Historical", "SSP1-2.6", 
                                              "SSP5-8.5")),
                scale_fill_manual(values = c("historical" = "black", 
                                             "ssp126" = "#33bbee", 
                                             "ssp585" = "#ee3377"), 
                                  name = "Scenarios",
                                  labels = c("Historical", "SSP1-2.6", 
                                             "SSP5-8.5")),
                guides(color = guide_legend(nrow = 1, title.position = "left")),
                theme_classic(),
                labs(y = "Change in exploitable\nfish biomass (%)"),
                theme(legend.position = "top", legend.justification = "center", 
                      legend.text = element_text(size = 10),
                      panel.grid.minor.y = element_blank(), 
                      axis.title.x = element_blank(),
                      axis.title.y = element_text(size = 12),
                      axis.text.x = element_text(angle = 45, vjust = 0.765, 
                                                 hjust = 0.65)))
```

## Trends globally and for top 7 countries with highest catches
  
```{r}
#Merging global and top 7 countries data

top_7_data<-top_7_data |>
  mutate(Country_FAO = case_when(Country_FAO == "the People's Republic of China" ~ "China",
                             Country_FAO == "the Republic of India" ~ "India",
                             Country_FAO == "the Republic of Indonesia" ~ "Indonesia",
                             Country_FAO == "the Republic of Peru" ~ "Peru",
                             Country_FAO == "the Russian Federation" ~ "Russian Federation",
                             Country_FAO == "the Socialist Republic of Viet Nam" ~ "Viet Nam",
                             Country_FAO == "the United States of America" ~ "United States of America",
                             .default = as.character(Country_FAO)))
  
top_7<-global_bio |> 
  #Add an extra column to match top 7 countries
  mutate(Country_FAO = "Global", .after = "scenario") |> 
  #Joining to top7 data
  bind_rows(top_7_data) |> 
  #Create an order category for SOVEREIGN1
  mutate(Country_FAO = factor(Country_FAO, levels = c("Global", 
                                              unique(top_7_data$Country_FAO)), 
                             ordered = T)) |> 
  #Plotting
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  #Applying base plot design
  base_gg+
  #Plotting by country
  facet_wrap(~Country_FAO, scales = "free_y", nrow = 4)+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 11, face = "bold"))+ 
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
## saving   
  
```{r, eval = F}
pdf(file.path(base_folder, "top_7_global_trends.pdf"), width=9, height= 10)
top_7
dev.off()
```

## Figure - Asia
In this figure, we will exclude marine areas under the jurisdiction of countries in Asia as classified by the FAO.  
  
```{r}

#Filtering data
asia<-count_bio |> 
  #Extract EEZs in Asia
  filter(continent == "Asia") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
## Saving   
  
```{r, eval = F}
pdf(file.path(base_folder, "asia_perc_change.pdf"), width=9, height= 10)
asia
dev.off()

```
  
## Figure - Africa
In this figure, we will exclude marine areas under the jurisdiction of countries in Africa as classified by the FAO.  
  
```{r}
#Filtering data
africa<-count_bio |> 
  #Extract EEZs in Africa
  filter(continent == "Africa") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5, 
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
## Saving   
  
```{r, eval = F}
pdf(file.path(base_folder, "africa_perc_change.pdf"), width=9, height= 12)
africa
dev.off()
```
  
## Figure - Europe
In this figure, we will exclude marine areas under the jurisdiction of countries in Europe as classified by the FAO.  
  
```{r}
#Filtering data

sort(unique(count_bio$figure_name))

count_bio_europe<-count_bio %>% 
  mutate(figure_name = ifelse(figure_name == "United Kingdom of Great Britain and Northern Ireland", "United Kingdom" , figure_name))

europe<-count_bio_europe |> 
  #Extract EEZs in Europe
  filter(continent == "Europe") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```

## Saving 

```{r, eval = F}
pdf(file.path(base_folder, "europe_perc_change.pdf"), width=9, height= 9)
europe
dev.off()
```
  
## Figure - Americas
In this figure, we will exclude marine areas under the jurisdiction of countries in the Americas as classified by the FAO.  
  
```{r}
#Filtering data
americas<-count_bio |> 
  #Extract EEZs in the Americas
  filter(continent == "Americas") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
   theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
## Saving  
  
```{r, eval = F}
pdf(file.path(base_folder, "americas_perc_change.pdf"), width=9, height= 13)
americas
dev.off()
```
  
## Figure - Oceania
In this figure, we will exclude marine areas under the jurisdiction of countries in Oceania as classified by the FAO.  
  
```{r}
#Filtering data
oceania<-count_bio |> 
  #Extract EEZs in Oceania
  filter(continent == "Oceania") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~figure_name, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(figure_name = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
## Saving   
  
```{r, eval = F}
pdf(file.path(base_folder, "oceania_perc_change.pdf"), width=9, height= 9)
oceania
dev.off()
```
  
## Figure - FAO regions
  
```{r}
#Filtering data
fao_regions<-fao_bio |> 
  #Plotting data
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by FAO sector
  facet_wrap(~NAME_EN, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(NAME_EN = label_wrap_gen(25)))+
  theme(strip.background = element_blank(),
        strip.text.x = element_text(margin = margin(0.08,0,0.08,0, "cm"), size= 9))+ # face = "bold"
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

```
  
## Saving   
  
```{r, eval = F}
pdf(file.path(base_folder, "fao_regions_perc_change.pdf"), width=9, height= 6)
fao_regions
dev.off()
```
  
## Uncertaity plot

```{r}

# use mean and sd calculated above

# sort values
sorting<-table %>%
  filter(spatial_scale == "FAO_area") %>%
  filter(scenario == "ssp585", decade == "2091-2100") %>%
  arrange(-mean)

sorting<-unique(sorting$figure_name)
sorting<-factor(sorting, levels=unique(sorting))

bio_dataOCD <-table %>%
  filter(spatial_scale == "FAO_area") %>%
  mutate(figure_name = factor(figure_name, levels = sorting))

uncert_decade<-ggplot(filter(bio_dataOCD, figure_name != "Arctic Sea"), aes(x=figure_name, y=mean, group=scenario, color=scenario)) +
  geom_point()+
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                position=position_dodge(0.05))+
  theme_classic() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey80", linewidth = 1)+
  scale_color_manual(name = "Scenario", labels = c("SSP1-2.6", "SSP5-8.5"), values=c("#33bbee", "#ee3377"))+
  labs(x="FAO Major Fishing Area", y = "Change in exploitable fish biomass (%)") +
  coord_flip()+
  theme(axis.text=element_text(size=14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 14),
        axis.title=element_text(size=14)) + # face = "bold"
  facet_wrap(~decade)

pdf(file.path(base_folder, "FAOarea_uncertity.png"), width=10, height=5, units="in",res=1200)
uncert_decade
dev.off()

```
  