---
title: "Plotting biomass change under different SSPs"
author: "Denisse Fierro Arcos"
date: "2024-01-09"
output: github_document
---

# Introduction

In this notebook, we will calculate the mean percentage change in fish biomass between 2041 and 2050 under two climate emission scenarios: `SSP1-2.6` (low emissions) and `SSP5-8.5` (very high emissions). The last decade from the historical period (2005-2014) was used as the reference period.  
  
Note that outputs from all [global FishMIP models](https://fishmip.org/modellingteams.html) were used to calculate the mean percentage change for the entire ensemble.  
  
## Loading libraries
  
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(rnaturalearth)
library(data.table)
library(cowplot)
library(sf)
```

## Setting up notebook
  
```{r}
#FAO_EEZ key - Contains continent information
fao_eez <- read_csv("/rd/gem/private/users/ldfierro/Merged_FAO_EEZ/FAO-MajorAreas_EEZs_keys.csv")


#FAO_EEZ mask - Contains IDs for FAO and EEZ areas
fao_eez_mask <- read_csv("/rd/gem/private/users/ldfierro/Merged_FAO_EEZ/Outputs/masked_area_1deg.csv") |> 
   rename(x = Lon, y = Lat, ID_mrgd = mask)

#Defining location of notebook outputs
out_folder <- "/rd/gem/private/users/camillan/FAO_Report"
if(!dir.exists(out_folder)){
  dir.create(out_folder)
}

#Defining location of model outputs 10g-100kg
data_dir <- "/rd/gem/private/users/camillan/Extract_tcblog10_Data/Output/sumSize_annual/sizeConsidered10g_100kg/EEZsummaries/gridded_outputs/" 

#Getting a list of files containing biomass data
global_files <- list.files(data_dir, full.names = T)

#Getting a list of models
members <- str_extract(global_files, "outputs//(.*)_(h|s)", group = 1) |> 
  unique()

#Define Robinson projection to be applied to shapefile and basemap
#rob_proj <- "+proj=robin +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"

```
  
## Calculating percentage change in fish biomass estimates from global FishMIP models
We will go through each FishMIP model and calculate the mean fish biomass for the decade between 2005 and 2014 (last decade of `historical` period), and for the period between 2041 and 2050 (for the two emission scenarios). Finally, we will calculate the percentage change between these two decades.  
  
```{r}
#Looping through each FishMIP model
for(m in members){
  #Load all data available for a single FishMIP model
  df_model <- str_subset(global_files, m) |> 
    #Ignore columns SOVEREIGN1-3 - not needed here
    map_df(~fread(., drop = c(paste0("SOVEREIGN", 1:3), "area_m", "GEONAME", "eez"))) |> 
    #Extract data only for years to be used in maps
    filter(year >= 2005 & year <= 2014 | year >= 2041 & year <= 2050 | year >= 2091 & year <= 2100) |> 
    #Create new group column to calculate means
    mutate(group = case_when(year <= 2014 ~ "reference",
                             year >= 2041 & year <= 2050 ~ "mean50",
                             year >= 2091 & year <= 2100 ~ "mean00"),
           #The new group column also need to have the scenario as part of the label
           group = case_when(group != "reference" ~ str_c(group, scenario, sep = "_"),
                             T ~ group)) |> 
    #Apply mask
    left_join(fao_eez_mask, by = join_by(x, y)) |> 
    #Calculate mean per ensemble member
    group_by(x, y, mem, esm, ID_mrgd, group) |> 
    summarise(mean_bio = mean(biomass, na.rm = T)) |> 
    #Reorganise table to facilitate calculations
    pivot_wider(names_from = group, values_from = mean_bio) |> 
    ungroup() |> 
    #Calculate % change in fish biomass for the two emissions scenarios
    mutate(rel_change_mean50_ssp126 = ((mean50_ssp126-reference)/reference)*100,
           rel_change_mean50_ssp585 = ((mean50_ssp585-reference)/reference)*100,
           rel_change_mean00_ssp126 = ((mean00_ssp126-reference)/reference)*100,
           rel_change_mean00_ssp585 = ((mean00_ssp585-reference)/reference)*100)
  
  #Create name to save file  
  f_out <- file.path(out_folder, str_c(m, "_perc_bio_change_data_map.csv"))
  
  #Saving results for each model
  df_model |> 
    fwrite(f_out)
}
  
```
  
### Calculating ensemble mean
We will now load the percentage change in biomass for all global models and calculate an ensemble mean.  
  
```{r}
#Listing all relevant files to calculate biomass projections
maps_data <- list.files(out_folder, pattern = "_perc_bio_change_data_map.csv", full.names = T) |> 
  map_df(~fread(.)) |> 
  #Calculations performed by year and EEZ
  group_by(x, y, ID_mrgd) |> 
  #Apply calculations to biases only
  summarise(across(reference:rel_change_mean00_ssp585, 
                   #Listing statistics to be calculated
                   list(mean = mean), 
                   #Setting column names
                   .names = "{.col}_{.fn}")) |> 
  ungroup()
```
## Loading information about continents
We will add information about the continents where each EEZ is located.  
  
```{r}
#Merge with biomass data
maps_data <- maps_data |> 
  inner_join(fao_eez, by = join_by(ID_mrgd)) |> 
  #If a grid cell is not assigned to an EEZ, then change to NA
  mutate(across(starts_with("rel_change"), ~ case_when(is.na(ID_mrgd) ~ NA,
                                                      T ~ .x))) 

```

## Plotting maps
In this section, we will plot individual maps for fish biomass change per decade/scenario combination. We will merge them together in a single plot.  
  
### Loading relevant shapefiles
  
```{r}
#Base map of the world
world <- ne_countries(returnclass = "sf") 

#EEZ boundaries
eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |> 
  #Removing Antarctic EEZ
  filter(str_detect(GEONAME, "Antarctic", negate = T))

```
  
### Creating individual maps
All maps share the same colour palette, so we will not include legends in any of them except for the last one.

Grey or black cells are areas with no data in the models.  
  
```{r}
# theme for Cami's plot_tcblog10()
my_theme_map<-theme_bw()+ 
  theme(
    text = element_text(size = 9), 
    plot.title = element_text(size = 9),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.title=element_text(size = 8), 
    legend.text=element_text(size = 7),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    plot.background = element_rect(fill="white"),
    panel.border = element_blank(),
    legend.key.height = unit(0.2, "cm"), # this becomes width if you choose bottom position
    legend.key.width = unit(0.7, "cm"),
    legend.position= 'bottom' # c(0.5,0.6) # in middle
)
```


```{r}
#Using some settings from Cami's plot_tcblog10() code:
min=-50
max=50
breaks_vector_n=10
breaks_vector = seq(min, max, length.out = breaks_vector_n)
labels_vector = format(round(seq(min, max, length.out = breaks_vector_n), digits = 0), nsmall = 0)
 


#SSP1-2.6 2041-2050
p50_126 <- maps_data |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  # scale_fill_distiller(palette = "RdYlBu", direction = 1,
  #                      na.value = "white", 
  #                      rescaler = ~scales::rescale_mid(.x, from = c(-50, 50), 
  #                                                      # to = c(-50, 50),
  #                                                      mid = 0))+
    scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         breaks = breaks_vector,
                                         labels = labels_vector,
                                         limits=c(min, max),
                                         oob = scales::squish,
                                         na.value = "black") +
  geom_sf(inherit.aes = F, data = world)+
  geom_sf(inherit.aes = F, data = eez, fill = NA,
          show.legend = F, linewidth = 0.25)+
  labs(title = "SSP1-2.6: 2041-2050")+
  # theme(axis.title = element_blank(), 
  #       plot.title = element_text(hjust = 0.5),
  #       legend.position = "none") +
  my_theme_map +
  theme(legend.position = "none")

#SSP5-8.5 2041-2050
p50_585 <- maps_data |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  # scale_fill_distiller(palette = "RdYlBu", direction = 1,
  #                      na.value = "white", 
  #                      rescaler = ~scales::rescale_mid(.x, from = c(-50, 50), 
  #                                                      # to = c(-50, 50),
  #                                                      mid = 0))+
    scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         breaks = breaks_vector,
                                         labels = labels_vector,
                                         limits=c(min, max),
                                         oob = scales::squish,
                                         na.value = "black") +
  geom_sf(inherit.aes = F, data = world)+
  geom_sf(inherit.aes = F, data = eez, fill = NA,
          show.legend = F, linewidth = 0.25)+
  theme_bw()+
  labs(title = "SSP5-8.5: 2041-2050")+
  # theme(axis.title = element_blank(), 
  #       plot.title = element_text(hjust = 0.5),
  #       legend.position = "none")
  my_theme_map +
  theme(legend.position = "none")

#SSP1-2.6 2091-2100
p00_126 <- maps_data |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_raster()+
  # scale_fill_distiller(palette = "RdYlBu", direction = 1,
  #                      na.value = "white", 
  #                      rescaler = ~scales::rescale_mid(.x, from = c(-50, 50), 
  #                                                      # to = c(-50, 50),
  #                                                      mid = 0))+
    scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         breaks = breaks_vector,
                                         labels = labels_vector,
                                         limits=c(min, max),
                                         oob = scales::squish,
                                         na.value = "black") +
  geom_sf(inherit.aes = F, data = world)+
  geom_sf(inherit.aes = F, data = eez, fill = NA,
          show.legend = F, linewidth = 0.25)+
  theme_bw()+
  labs(title = "SSP1-2.6: 2091-2100")+
  # theme(axis.title = element_blank(), 
  #        plot.title = element_text(hjust = 0.5),
  #        legend.position = "none")
  my_theme_map +
  theme(legend.position = "none")

#SSP5-8.5 2091-2100

p00_585 <- maps_data |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_raster()+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         # limits=c(min, max),
                                         # oob = scales::squish,
                                         na.value = "black") +
  
  # scale_fill_distiller(palette = "RdYlBu", direction = 1,
  #                      na.value = "white", limits = c(-50, 50),
  #                      breaks = c(seq(-50, 0, length.out = 5),
  #                                 seq(32.5, 260, length.out = 8)),
  #                      rescaler = ~scales::rescale_mid(.x, from = c(-50, 50), 
  #                                                      # to = c(-50, 50),
  #                                                      mid = 0))+
  geom_sf(inherit.aes = F, data = world)+
  geom_sf(inherit.aes = F, data = eez, fill = NA,
          show.legend = F, linewidth = 0.25)+
  labs(title = "SSP5-8.5: 2091-2100")+
  guides(fill = guide_legend(title = "% change", title.position = "top",
                             title.hjust = 0.5,
                             label.position = "bottom", nrow = 1))+
  theme(axis.title = element_blank(),
        legend.spacing.x = unit(0, "cm"), legend.position = "bottom",
        legend.direction = "horizontal", legend.key.width = unit(1.5, "cm"),
        legend.key.height = unit(0.75, "cm"))
  my_theme_map 

#Get legend
legend <- get_legend(p00_585)

#Remove legend from last plot
p00_585 <- p00_585+theme(legend.position = "none")

#Create title for 
# title <- ggdraw()+
#   draw_label("Mean % change in exploitable fish biomass from 2005-2014")+
#   theme(plot.margin = margin(0, 0, 0, 0, unit = "cm"))

#Plotting everything together
all_plots <- plot_grid(plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
                       plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
                       legend, ncol = 1, nrow = 4, 
                       rel_heights = c(0.15, 1, 1, 0.4))

#checking patchwork layout
# library(patchwork)
# all_plots<-p50_126+p50_585 +p00_126+p00_585

all_plots
```
  
### GLOBAL - Saving final global plot
  
```{r}
#ggsave(file.path(out_folder, "global_perc_change_map_50-00.pdf"), device = "pdf", width = 14, height = 9)

# name_Fig = paste("figures/", Sys.Date(),"_" ,"global_perc_change_map.pdf", sep = "")
# 
# pdf(name_Fig)
# all_plots
# dev.off()

```



### ASIA

In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_asia <- maps_data |>
  filter(region=="Asia")

#Base map of the world
world <- ne_countries(returnclass = "sf") |> 
  st_shift_longitude() |>
  filter(continent %in% c("Asia")) 

asia_fao_eez<-filter(fao_eez,region=="Asia")

#EEZ boundaries
eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |> 
  #Removing Antarctic EEZ
  filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
  filter(SOVEREIGN1 %in%  asia_fao_eez$SOVEREIGN1) 

```

colour palette
```{r}
library(cmocean)
pal <- c(cmocean("matter", start = 0.1, end = 0.8, direction = -1)(179), 
         cmocean("delta", start = 0.49, end = 0.5)(20),
         cmocean("deep", start = 0.1, end = 0.3)(57))
```

Try and map this to see if it works for region

```{r}
p50_126 <- maps_data_asia  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))

p50_126
```
Maximum losses for region around -20%
Most EEZs show losses, exception of small EEZ in northwest (?)

```{r}
p50_585 <- maps_data_asia  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))

p50_585
```
Under ssp585 max loss is - 30%
Greater fraction of area < -20%
By 2050

Next by 2100

```{r}
p00_126 <- maps_data_asia  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))

p00_126
```


Under strong mitigation- now the losses are much reduced relative to reference period by 2100 

```{r}
p00_585 <- maps_data_asia  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  # theme_bw()  + 
  # theme(axis.title = element_blank(), 
  #       plot.title = element_text(hjust = 0.5))

p00_585
```

Without  mitigation-  the losses are great max -50% by 2100 


#### ASIA PLOT
```{r}
#Plotting everything together
all_plots <-(p50_126+  labs(title = "SSP1-2.6: 2050"))+
  (p50_585 +  labs(title = "SSP5-8.5: 2050"))+ 
  (p00_126+  labs(title = "SSP1-2.6: 2100"))+ 
  (p00_585+  labs(title = "SSP5-8.5: 2100"))
# all_plots <- plot_grid(title, plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
#                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
#                        legend, ncol = 1, nrow = 4, 
#                        rel_heights = c(0.15, 1, 1, 0.4))

ggsave(file.path(out_folder, "ASIA_perc_change_map_50-00.pdf"), device = "pdf",
       width = 14, height = 9)
```



### AFRICA

In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(region=="Africa")

#Base map of the world
world <- ne_countries(returnclass = "sf") |> 
  filter(continent %in% c("Africa"))#|>
#  st_shift_longitude() 
 

region_fao_eez<-filter(fao_eez,region=="Africa")

#EEZ boundaries
eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |> 
  #Removing Antarctic EEZ
  filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
  filter(GEONAME %in%  region_fao_eez$nm_mrgd) # |> 
  # #Ensure ID and name are factors
  # mutate(ID_1 = as.factor(ID_1)) |> 
  # #Ensuring shapefile crosses the international dateline
  #st_shift_longitude() 
```


Let's look at 2050 first


```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126
```

Max losses - 20% 
Declines most of continent except norther Africa some increases up to 30%

```{r}
p50_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_585
```
More pronouced losses by 2050 under ssp585 - up to -30%, though some additional (modest) gains

```{r}
p00_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_126
```

No further reduction from 2050 and some recovery/gains

```{r}
p00_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_585
```
Losses more widespread under ssp585 for most EEZ up to - 30%
Isolated regions where increases projected, but smaller

#### AFRICA PLOTS

```{r}
#Plotting everything together
all_plots <-(p50_126+  labs(title = "SSP1-2.6: 2050"))+
  (p50_585 +  labs(title = "SSP5-8.5: 2050"))+ 
  (p00_126+  labs(title = "SSP1-2.6: 2100"))+ 
  (p00_585+  labs(title = "SSP5-8.5: 2100"))
# all_plots <- plot_grid(title, plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
#                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
#                        legend, ncol = 1, nrow = 4, 
#                        rel_heights = c(0.15, 1, 1, 0.4))

ggsave(file.path(out_folder, "AFRICA_perc_change_map_50-00.pdf"), device = "pdf",
       width = 14, height = 9)
```



### EUROPE

In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(region=="Europe")

#Base map of the world
world <- ne_countries(returnclass = "sf") |> 
  filter(continent %in% c("Europe"))#|>
#  st_shift_longitude() 
 

region_fao_eez<-filter(fao_eez,region=="Europe")

#EEZ boundaries
eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |> 
  #Removing Antarctic EEZ
  filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
  filter(GEONAME %in%  region_fao_eez$nm_mrgd) # |> 
  # #Ensure ID and name are factors
  # mutate(ID_1 = as.factor(ID_1)) |> 
  # #Ensuring shapefile crosses the international dateline
  #st_shift_longitude() 


```


Let's look at 2050 first


```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126


# need to get area weighted means:
europe_summaries<-maps_data_region |>
    group_by(nm_mrgd) %>%
    summarise(spatialMeanChange50_ssp126=mean(rel_change_mean50_ssp126_mean,na.rm=T),
      spatialMeanChange00_ssp126=mean(rel_change_mean00_ssp126_mean,na.rm=T), 
      spatialMeanChange50_ssp585=mean(rel_change_mean50_ssp585_mean,na.rm=T), spatialMeanChange00_ssp585=mean(rel_change_mean00_ssp585_mean),na.rm=T )


europe_summaries[which(europe_summaries$spatialMeanChange50_ssp126<0),]
length(europe_summaries$spatialMeanChange50_ssp126)
europe_summaries[which(europe_summaries$spatialMeanChange50_ssp126>=0),]



```

Max losses - 30% 
Large increases in polar regions & also eastern mediterranean

```{r}
p50_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_585

min(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
min(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
```
More pronouced losses by 2050 under ssp585 - up to -30%, less extreme gains

```{r}
p00_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_126
```

Drastic reduction in losses from 2050 and MASSIVE gains in polar regions... hmmm would be good to look at uncertainty , perhaps include maps of across model SD ?

```{r}
p00_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_585
min(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
```
Dominated by large gains (+243) in polar, ocmpared to reductions (-39% still pretty bad though for some EEZs), highly uncertain region
Starting to think we show focus on 2050...


#### EUROPE PLOT

```{r}
#Plotting everything together
all_plots <-(p50_126+  labs(title = "SSP1-2.6: 2050"))+
  (p50_585 +  labs(title = "SSP5-8.5: 2050"))+ 
  (p00_126+  labs(title = "SSP1-2.6: 2100"))+ 
  (p00_585+  labs(title = "SSP5-8.5: 2100"))
# all_plots <- plot_grid(title, plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
#                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
#                        legend, ncol = 1, nrow = 4, 
#                        rel_heights = c(0.15, 1, 1, 0.4))

ggsave(file.path(out_folder, "AFRICA_perc_change_map_50-00.pdf"), device = "pdf",
       width = 14, height = 9)
all_plots






```



### AMERICAS

In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(region=="Americas")

#Base map of the world
world <- ne_countries(returnclass = "sf") |> 
  filter(continent %in% c("Americas"))#|>
#  st_shift_longitude() 
 

region_fao_eez<-filter(fao_eez,region=="Americas")

#EEZ boundaries
eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |> 
  #Removing Antarctic EEZ
  filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
  filter(GEONAME %in%  region_fao_eez$nm_mrgd) # |> 
  # #Ensure ID and name are factors
  # mutate(ID_1 = as.factor(ID_1)) |> 
  # #Ensuring shapefile crosses the international dateline
  #st_shift_longitude() 
```


Let's look at 2050 first


```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126
min(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
```

Area dominated by losses (max loss - 23%), but scale skewed towards high increases (up to 98% in localised area) in Arctic (uncertain, need to show SD somewhere)



```{r}
p50_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_585
min(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
```
More pronouced losses by 2050 under ssp585 - up to -32%, less extreme gains but still high

```{r}
p00_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_126
min(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp125_mean,na.rm=T)

```
Again...
Drastic reduction in losses from 2050 and MASSIVE gains in polar regions... hmmm would be good to look at uncertainty , perhaps include maps of across model SD ?

```{r}
p00_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_585
min(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
```
ALSO : Dominated by large gains (+244) in polar, ocmpared to reductions (-57% still pretty bad though for some EEZs), highly uncertain region
Starting to think we show focus on 2050...

#### AMERICAS MAP

```{r}
#Plotting everything together
all_plots <-(p50_126+  labs(title = "SSP1-2.6: 2050"))+
  (p50_585 +  labs(title = "SSP5-8.5: 2050"))+ 
  (p00_126+  labs(title = "SSP1-2.6: 2100"))+ 
  (p00_585+  labs(title = "SSP5-8.5: 2100"))
# all_plots <- plot_grid(title, plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
#                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
#                        legend, ncol = 1, nrow = 4, 
#                        rel_heights = c(0.15, 1, 1, 0.4))

ggsave(file.path(out_folder, "AFRICA_perc_change_map_50-00.pdf"), device = "pdf",
       width = 14, height = 9)
all_plots
```


### OCEANIA


In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(region=="Oceania")

#Base map of the world
world <- ne_countries(returnclass = "sf") |> 
  filter(continent %in% c("Oceania"))|>
  st_shift_longitude() 
 

region_fao_eez<-filter(fao_eez,region=="Oceania")

#EEZ boundaries
eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |>
  #Removing Antarctic EEZ
  filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
  filter(GEONAME %in%  region_fao_eez$nm_mrgd)  |> 
  # #Ensure ID and name are factors
  # mutate(ID_1 = as.factor(ID_1)) |> 
  # #Ensuring shapefile crosses the international dateline
  st_shift_longitude() 
```


Let's look at 2050 first


```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  geom_sf(inherit.aes = F, data = eez, fill = NA, 
          show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126
min(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
```



### FAO ATLANTIC


In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(OCEAN=="Atlantic")

#Base map of the world
world <- ne_countries(returnclass = "sf") 
 

region_fao_eez<-filter(fao_eez,OCEAN=="Atlantic")

# #EEZ boundaries
# eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |>
#   #Removing Antarctic EEZ
#   filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
#   filter(OCEAN %in%  region_fao_eez$nm_mrgd)  |> 
#   # #Ensure ID and name are factors
#   # mutate(ID_1 = as.factor(ID_1)) |> 
#   # #Ensuring shapefile crosses the international dateline
#   st_shift_longitude() 
```


Let's look at 2050 first


#### ATLANTIC MAPS

```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126
min(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
```


########## WARNING - MANUALLY CHANGING THESE

### FAO INDIAN


In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(OCEAN=="Indian")

#Base map of the world
world <- ne_countries(returnclass = "sf") 
 

region_fao_eez<-filter(fao_eez,OCEAN=="Indian")

# #EEZ boundaries
# eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |>
#   #Removing Antarctic EEZ
#   filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
#   filter(OCEAN %in%  region_fao_eez$nm_mrgd)  |> 
#   # #Ensure ID and name are factors
#   # mutate(ID_1 = as.factor(ID_1)) |> 
#   # #Ensuring shapefile crosses the international dateline
#   st_shift_longitude() 
```


Let's look at 2050 first


```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126
min(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
```
Need to fix map

Like most regions - area is dominated by losses range:
[1] -32.96508
[1] 49.2035


```{r}
p50_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_585
min(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
```
More pronouced losses by 2050 under ssp585 - up to -30%, minimal gains
[1] -35.14472
[1] 51.78325

```{r}
p00_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
   geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_126
min(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)

```
Range
[1] -36.81024
[1] 58.52815
Again area dominated by losses

```{r}
p00_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_585
min(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
```
[1] -70.03788
[1] 229.2175

#### INDIAN HIGH SEAS MAP

```{r}
#Plotting everything together
all_plots <-(p50_126+  labs(title = "SSP1-2.6: 2050"))+
  (p50_585 +  labs(title = "SSP5-8.5: 2050"))+ 
  (p00_126+  labs(title = "SSP1-2.6: 2100"))+ 
  (p00_585+  labs(title = "SSP5-8.5: 2100"))
# all_plots <- plot_grid(title, plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
#                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
#                        legend, ncol = 1, nrow = 4, 
#                        rel_heights = c(0.15, 1, 1, 0.4))

ggsave(file.path(out_folder, "Inidan_perc_change_map_50-00.pdf"), device = "pdf",
       width = 14, height = 9)
all_plots
```



### FAO PACIFIC


In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(OCEAN=="Pacific")

#Base map of the world
world <- ne_countries(returnclass = "sf") 
 

region_fao_eez<-filter(fao_eez,OCEAN=="Pacific")

# #EEZ boundaries
# eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |>
#   #Removing Antarctic EEZ
#   filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
#   filter(OCEAN %in%  region_fao_eez$nm_mrgd)  |> 
#   # #Ensure ID and name are factors
#   # mutate(ID_1 = as.factor(ID_1)) |> 
#   # #Ensuring shapefile crosses the international dateline
#   st_shift_longitude() 
```


Let's look at 2050 first


```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126
min(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
```
Need to fix map

Like most regions - area is dominated by losses range:
[1] -32.96508
[1] 49.2035


```{r}
p50_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_585
min(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
```
More pronouced losses by 2050 under ssp585 - up to -30%, minimal gains
[1] -35.14472
[1] 51.78325

```{r}
p00_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
   geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_126
min(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)

```
Range
[1] -36.81024
[1] 58.52815
Again area dominated by losses

```{r}
p00_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_585
min(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
```
[1] -70.03788
[1] 229.2175

#### PACIFIC HIGH SEAS MAP

```{r}
#Plotting everything together
all_plots <-(p50_126+  labs(title = "SSP1-2.6: 2050"))+
  (p50_585 +  labs(title = "SSP5-8.5: 2050"))+ 
  (p00_126+  labs(title = "SSP1-2.6: 2100"))+ 
  (p00_585+  labs(title = "SSP5-8.5: 2100"))
# all_plots <- plot_grid(title, plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
#                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
#                        legend, ncol = 1, nrow = 4, 
#                        rel_heights = c(0.15, 1, 1, 0.4))

ggsave(file.path(out_folder, "Inidan_perc_change_map_50-00.pdf"), device = "pdf",
       width = 14, height = 9)
all_plots
```

### FAO Southern Ocean


In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(OCEAN=="Southern Ocean")

#Base map of the world
world <- ne_countries(returnclass = "sf") 
 

region_fao_eez<-filter(fao_eez,OCEAN=="Southern Ocean")

# #EEZ boundaries
# eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |>
#   #Removing Antarctic EEZ
#   filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
#   filter(OCEAN %in%  region_fao_eez$nm_mrgd)  |> 
#   # #Ensure ID and name are factors
#   # mutate(ID_1 = as.factor(ID_1)) |> 
#   # #Ensuring shapefile crosses the international dateline
#   st_shift_longitude() 
```


Let's look at 2050 first


```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126
min(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
```
Need to fix map

Like most regions - area is dominated by losses range:
[1] -32.96508
[1] 49.2035


```{r}
p50_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_585
min(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
```
More pronouced losses by 2050 under ssp585 - up to -30%, minimal gains
[1] -35.14472
[1] 51.78325

```{r}
p00_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
   geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_126
min(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)

```
Range
[1] -36.81024
[1] 58.52815
Again area dominated by losses

```{r}
p00_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_585
min(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
```
[1] -70.03788
[1] 229.2175

#### SOUTHERN OCEAN HIGH SEAS MAP

```{r}
#Plotting everything together
all_plots <-(p50_126+  labs(title = "SSP1-2.6: 2050"))+
  (p50_585 +  labs(title = "SSP5-8.5: 2050"))+ 
  (p00_126+  labs(title = "SSP1-2.6: 2100"))+ 
  (p00_585+  labs(title = "SSP5-8.5: 2100"))
# all_plots <- plot_grid(title, plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
#                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
#                        legend, ncol = 1, nrow = 4, 
#                        rel_heights = c(0.15, 1, 1, 0.4))

ggsave(file.path(out_folder, "Inidan_perc_change_map_50-00.pdf"), device = "pdf",
       width = 14, height = 9)
all_plots
```

### FAO Arctic


In this section the maps will focus on a particular region and EEZs in that region

```{r}
# FishMIP data
maps_data_region <- maps_data |>
  filter(OCEAN=="Arctic")

#Base map of the world
world <- ne_countries(returnclass = "sf") 
 

region_fao_eez<-filter(fao_eez,OCEAN=="Arctic")

# #EEZ boundaries
# eez <- read_sf("/rd/gem/private/users/ldfierro/World_EEZ_v11_20191118/eez_short.shp") |>
#   #Removing Antarctic EEZ
#   filter(str_detect(GEONAME, "Antarctic", negate = T)) |>
#   filter(OCEAN %in%  region_fao_eez$nm_mrgd)  |> 
#   # #Ensure ID and name are factors
#   # mutate(ID_1 = as.factor(ID_1)) |> 
#   # #Ensuring shapefile crosses the international dateline
#   st_shift_longitude() 
```


Let's look at 2050 first


```{r}
p50_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_126
min(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp126_mean,na.rm=T)
```
Need to fix map

Like most regions - area is dominated by losses range:
[1] -32.96508
[1] 49.2035


```{r}
p50_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean50_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p50_585
min(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean50_ssp585_mean,na.rm=T)
```
More pronouced losses by 2050 under ssp585 - up to -30%, minimal gains
[1] -35.14472
[1] 51.78325

```{r}
p00_126 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp126_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
   geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_126
min(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp126_mean,na.rm=T)

```
Range
[1] -36.81024
[1] 58.52815
Again area dominated by losses

```{r}
p00_585 <- maps_data_region  |>
  ggplot(aes(x, y, fill = rel_change_mean00_ssp585_mean))+
  geom_tile()+
  #scale_fill_cmocean(name="balance")+
  # scale_fill_stepsn(colors = pal, n.breaks = 11, 
  #                   limits = c(-75, 15), show.limits = T)+
  scale_fill_continuous_divergingx(palette = 'Geyser', # ArmyRose
                                         mid = 1,
                                         rev = TRUE,
                                         guide = guide_colorbar(
                                           title = "% change",
                                           title.position = "top",
                                           title.hjust = 0.5),
                                         # breaks = breaks_vector,
                                         # labels = labels_vector,
                                         #limits=c(min, max),
                                         #oob = scales::squish,
                                         na.value = "grey") +
  # geom_sf(inherit.aes = F, data = eez, fill = NA, 
  #         show.legend = F, linewidth = 0.25)+
  geom_sf(inherit.aes = F, data = world)+
  theme_bw()  + 
  theme(axis.title = element_blank(), 
        plot.title = element_text(hjust = 0.5))


p00_585
min(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
max(maps_data_region$rel_change_mean00_ssp585_mean,na.rm=T)
```
[1] -70.03788
[1] 229.2175

#### ARCTIC HIGH SEAS MAP

```{r}
#Plotting everything together
all_plots <-(p50_126+  labs(title = "SSP1-2.6: 2050"))+
  (p50_585 +  labs(title = "SSP5-8.5: 2050"))+ 
  (p00_126+  labs(title = "SSP1-2.6: 2100"))+ 
  (p00_585+  labs(title = "SSP5-8.5: 2100"))
# all_plots <- plot_grid(title, plot_grid(p50_126, p50_126, ncol = 2, nrow = 1),
#                        plot_grid(p00_126, p00_585, ncol = 2, nrow = 1),
#                        legend, ncol = 1, nrow = 4, 
#                        rel_heights = c(0.15, 1, 1, 0.4))

ggsave(file.path(out_folder, "Inidan_perc_change_map_50-00.pdf"), device = "pdf",
       width = 14, height = 9)
all_plots
```

