---
title: "PLotting biomass vs catches"
author: "Camilla Novaglio"
date: "2024-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# dowload BOATS data

# tcb nat
scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/historical/boats_ipsl-cm6a-lr_nobasd_historical_nat_default_tcb_global_monthly_1950_2014.nc ./ 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/future/boats_ipsl-cm6a-lr_nobasd_ssp585_nat_default_tcb_global_monthly_2015_2100.nc ./ 

# tcb histsoc / 2015soc
scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/historical/boats_ipsl-cm6a-lr_nobasd_historical_histsoc_default_tcb_global_monthly_1950_2014.nc ./ 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/future/boats_ipsl-cm6a-lr_nobasd_ssp585_2015soc-from-histsoc_default_tcb_global_monthly_2015_2100.nc ./ 

# tc histsoc / 2015soc
scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/historical/boats_ipsl-cm6a-lr_nobasd_historical_histsoc_default_tc_global_monthly_1950_2014.nc ./ 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/future/boats_ipsl-cm6a-lr_nobasd_ssp585_2015soc-from-histsoc_default_tc_global_monthly_2015_2100.nc ./ 


```{r cars}

base_folder <- "/rd/gem/private/users/camillan/FAO_Report/BOATS_Data_for_catch_plot"

```
