---
title: "Temporal biomass change"
author: "Denisse Fierro Arcos"
date: "2024-01-17"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Introduction - Forcing file divergence

In this notebook, we will calculate the mean percentage change in fish biomass between 2041 and 2050 under two climate emission scenarios: `SSP1-2.6` (low emissions) and `SSP5-8.5` (very high emissions). The last decade from the historical period (2005-2014) was used as the reference period.  
  
Note that outputs from all [global FishMIP models](https://fishmip.org/modellingteams.html) were used to calculate the mean percentage change for the entire ensemble.  
  
## Loading libraries
  
```{r, message=FALSE, warning=FALSE}

rm(list=ls())
#Data wrangling
library(tidyverse)
library(data.table)
#Raster data
library(terra)
#Combining plots
library(cowplot)

```

## Setting up notebook
This notebook uses a modified version of the Exclusive Economic Zones boundaries (version 11) from the [Flanders Marine Institute (2019)](https://doi.org/10.14284/386). Available [online](https://www.marineregions.org/). The following modifications were made:  
- Boundaries for South Georgia and Sandwich Islands were updated to boundaries provided by the UN  
- Boundaries around Bouvet Island (Norway) in the south Atlantic were added to this shapefile  
- Country boundaries were dissolved and polygons were reclassified by continent  
  
```{r}
#Base folder for project
base_folder <- "/rd/gem/private/users/camillan"

#Defining location of notebook outputs
out_folder <- file.path(base_folder, "FAO_Report")
if(!dir.exists(out_folder)){
  dir.create(out_folder)
}

#EEZ keys
eez_keys <- read.csv(list.files(out_folder, 
                           pattern = "EEZ_keys_updated.csv", 
                           full.names = T, recursive = T), 
                     encoding = "UTF-8") |> 
  #Keeping relevant columns
  select(ID_mrgd, Name_EEZ, region, Country_FAO, 
         Official_Name_En_Stefania_NEW) |> 
  #Renaming columns
  rename("continent" = "region", 
         "fao_official" = "Official_Name_En_Stefania_NEW") |> 
  #Removing jointly managed areas and overlapping claims from original EEZ file
  filter(str_detect(Name_EEZ, "Overlapping|Joint", negate = T)) |> 
  #Removing unaccepted EEZ names
  select(!Name_EEZ)


### CN work in progress ----

## problem when merging (left_join) line 168 because of multiple matches
# need to check for records with same FAO official name and different Country_FAO. lots 
eez_keys[order(eez_keys$fao_official,eez_keys$continent),]
# need to understand why 2 Australia and why 2 French Polynesia
filter(eez_keys, fao_official == "French Polynesia")

#### Need to load, inspect and clean the original file following steps in email (remove ISO3CD = NA, and remove all ovelapping/joint territories in our EEZ dataset) :
eez_keys <- read.csv(list.files(out_folder, 
                           pattern = "EEZ_keys_updated.csv", 
                           full.names = T, recursive = T), 
                     encoding = "UTF-8")

filter(eez_keys, is.na(ISO3CD) & is.na(Official_Name_En_Stefania_NEW)) # WARNING - some like South African (Prince Edward Islands) Exclusive Economic Zone could/should(?) be included into e.g. South Africa as done for e.g. Chilean (Easter Island) Exclusive Economic Zone. But all of these are very small territories and can be done later. 

# remove ISO3CD = NA AND Official_Name_En_Stefania_NEW = NA (this includes jointly managed areas and overlapping claims from original EEZ file)
remove<-filter(eez_keys, is.na(ISO3CD) & is.na(Official_Name_En_Stefania_NEW))
eez_keys <- eez_keys %>% filter(!Name_EEZ %in% unique(remove$Name_EEZ))

# check 
filter(eez_keys, is.na(ISO3CD))
# remove copy-paste mistake in the original table where this is named "the Republic of Colombia"
eez_keys<-eez_keys %>% filter(!Name_EEZ == "Overlapping claim: Venezuela / Colombia / Dominican Republic")
filter(eez_keys, is.na(Official_Name_En_Stefania_NEW))

filter(eez_keys, Official_Name_En_Stefania_NEW %in% c("the United States of America (Alaska)", "the United States of America (Hawaii)"))
# correct for a decision taken on Hawaii and Alaska - these are considered US in FAO but as separate EEZ in our EEZ file. I wanted to keep this separation but changed my mind as we should fully follow FAO classification. 

eez_keys<-eez_keys %>% mutate(
  Country_FAO = ifelse(Official_Name_En_Stefania_NEW %in% c("the United States of America (Alaska)", "the United States of America (Hawaii)"), "United States of America" , Country_FAO),
  Official_Name_En_Stefania_NEW = ifelse(Official_Name_En_Stefania_NEW %in% c("the United States of America (Alaska)", "the United States of America (Hawaii)"), "the United States of America" , Official_Name_En_Stefania_NEW))

# check 
filter(eez_keys, Official_Name_En_Stefania_NEW == "the United States of America")
filter(eez_keys, Country_FAO == "the United States of America")

## do the same for Indian (Andaman and Nicobar Islands) Exclusive Economic Zone as this is part of one of the top 7 nations so need to be included into India even if not very influent
eez_keys<-eez_keys %>% mutate(
  Country_FAO = ifelse(Official_Name_En_Stefania_NEW %in% c("Indian (Andaman and Nicobar Islands) Exclusive Economic Zone"), "India" , Country_FAO),
  Official_Name_En_Stefania_NEW = ifelse(Official_Name_En_Stefania_NEW %in% c("Indian (Andaman and Nicobar Islands) Exclusive Economic Zone"), "the Republic of India" , Official_Name_En_Stefania_NEW))

# check if you still have duplicates: 
head(eez_keys)
trial<-eez_keys %>% 
  select(Name_EEZ,region, Official_Name_En_Stefania_NEW) %>% 
  arrange(Official_Name_En_Stefania_NEW,region)

## still some ovelapping claim in the data: these are records classified as territories by FAO and as ovelapping claim in the EEZ dataset. Becasue we don't know whether the boundaries of these territories include waters under other jurisdiction than the one specific by FAO, we remove them.    
eez_keys<-eez_keys %>% filter(str_detect(Name_EEZ, "Overlapping|Joint", negate = T))

## Clipperton problem: named french polynesia in both dataset, given as Oceania in FAO dataset and as Europe in EEZ dataset. 
# WARNING - this means we need to check continent classification but this can be done later as there are no other visible missmatches 
# change to Oceania 
eez_keys<-eez_keys %>% 
  mutate(region = ifelse(Name_EEZ == "Clipperton Exclusive Economic Zone", "Oceania", region))

# clean and change column names as per above: 
eez_keys<-eez_keys %>% 
  #Keeping relevant columns
  select(ID_mrgd, Name_EEZ, region, Country_FAO, 
         Official_Name_En_Stefania_NEW) |> 
  #Renaming columns
  rename("continent" = "region", 
         "fao_official" = "Official_Name_En_Stefania_NEW") %>% 
  #Removing column with unaccepted EEZ names
  select(!Name_EEZ) 

### end work in progress ----

#FAO keys
fao <- read_csv(list.files(out_folder, 
                           pattern = "FAO_MajorAreas_200NM_cont_keys.csv", 
                           full.names = T, recursive = T)) |> 
  #Keep FAO areas only
  drop_na(OCEAN) |> 
  #Removing duplicates
  distinct() |> 
  #Keep relevant columns only
  select(ID_mrgd, OCEAN, NAME_EN)

#Merging FAO and EEZ keys
fao_eez_keys <- fao |>
  bind_rows(eez_keys) |> 
  #Keep all official names in a single column for FAO and EEZ areas
  mutate(fao_official = case_when(is.na(fao_official) ~ NAME_EN, 
                                  T ~ fao_official)) |> 
  #Remove duplicate column
  select(!NAME_EN)
  
#Removing individual keys
rm(eez_keys, fao)

#FAO_EEZ mask - Contains IDs for FAO and EEZ areas
fao_eez_mask <- read_csv(list.files(out_folder, 
                                    pattern = "FAO_Major_200NM_masked_area_1deg.csv",
                                    full.names = T, recursive = T)) |> 
  #Renaming ID column
  rename("ID_mrgd" = "mask") |> 
  #Adding official names to be used in plots
  left_join(fao_eez_keys[,c("ID_mrgd", "fao_official")], 
            by = join_by(ID_mrgd))

#Loading area per grid cell
area_grid <- rast(list.files(out_folder, pattern = "^area_1deg.nc", 
                             recursive = T, full.names = T)) |> 
  #Transform to data frame
  as.data.frame(xy = T) |> 
  #Join with FAO keys
  left_join(fao_eez_mask, by = join_by(x, y))

#Removing mask without area values
rm(fao_eez_mask)

#Getting a list of files containing biomass data
global_files <- list.files(base_folder, pattern = "global_10g-100kg.rds", 
                           full.names = T, recursive = T)

#Getting a list of models
members <- str_extract(global_files, "annual_(.*)_(h|s)", group = 1) |> 
  unique()

```
  
## Calculating percentage change in fish biomass estimates from global FishMIP models
We will go through each FishMIP model and calculate the mean fish biomass for the decade between 2005 and 2014 (last decade of `historical` period), and for the period between 2041 and 2050 (for the two emission scenarios), and between 2091 and 2100 (two scenarios). Finally, we will calculate the percentage change between these two decades.  
  
```{r, eval = F}
# #Looping through each FishMIP model
# for(m in members){
#   #Load all data available for a single FishMIP model
#   df_model <- str_subset(global_files, m) |> 
#     map(~readRDS(.)) |> 
#     map_df(~bind_rows(.)) |> 
#     #Add area for each grid cell and area ID (FAO and EEZ)
#     left_join(area_grid, by = join_by(x, y)) |> 
#     #Removing areas not recognised in FAO official name list
#     drop_na(fao_official) |> 
#     #Calculate mean per ensemble member
#     group_by(year, mem, esm, scenario, fao_official) |> 
#     summarise(mean_bio = weighted.mean(biomass, area_m, na.rm = T))
#   
#   #Create name to save file  
#   f_out <- file.path(out_folder, str_c(m, "_yearly_perc_bio_change_data.csv"))
# 
#   #Saving results for each model
#   df_model |>
#     fwrite(f_out)
# }

```
  
## Calculating mean for reference period
Plots created here will show percentage change in biomass over time under two emissions scenarios: `SSP1-2.6` and `SSP5-8.5`. Our reference period is 2005-2014.  
  
```{r eval = F}
#Listing all relevant files to calculate biomass projections
bio_data <- list.files(out_folder, pattern = "_yearly_perc_bio_change_data.csv",
                       full.names = T) |> 
  map_df(~fread(.))

# filter(bio_data, fao_official == "the Federative Republic of Brazil")

#Calculating reference period
ref_bio <- bio_data |> 
  filter(year >= 2005 & year <= 2014) |> 
  #Calculate mean per ensemble member
  group_by(mem, esm, fao_official) |> 
  summarise(ref_bio = mean(mean_bio, na.rm = T))

#Getting list of unique keys
fao_eez <-  fao_eez_keys |> 
  select(!ID_mrgd) |> 
  distinct()

bio_data <- bio_data |> 
  #Add reference period to yearly biomass data
  left_join(ref_bio, by = c("mem", "esm", "fao_official")) |> 
  #Calculate percentage change per model/year(scenario)/region
  mutate(perc_change = ((mean_bio-ref_bio)/ref_bio)*100) |> 
  #Calculate mean and standard deviation for percentage change 
  #per year(scenario)/region
  group_by(year, scenario, fao_official) |> 
  summarise(mean_change = mean(perc_change, na.rm = T),
            sd_change = sd(perc_change, na.rm = T)) %>% 
  #Add names in FAO-EEZ keys
  left_join(fao_eez, by = "fao_official") |> 
  ungroup()

#Check results
bio_data
```
  
We can now save the results for future reference.  
  
```{r}
# #Create name to save file  
# f_out <- file.path(out_folder, "ensemble_perc_bio_change_data.csv")
# 
# #Saving results for each model
# bio_data |>
#   write_csv(f_out)
```
  
## Defining basic plot design
  
```{r}
base_gg <- list(geom_line(linewidth = 0.5),
                #Adding no change line at 0 for reference 
                geom_hline(yintercept = 0, color = "#709fcc", linewidth = 0.65, 
                           linetype = 2),
                #Adding line dividing historical period and future projections
                geom_vline(xintercept = 2015, color = "#709fcc", 
                           linewidth = 0.65),
                #Adding SD as shading 
                geom_ribbon(aes(ymin = mean_change-sd_change,
                                ymax = mean_change+sd_change, fill = scenario),
                            alpha = 0.3, color = NA),
                #Manually setting colours to be used in plots
                scale_color_manual(values = c("historical" = "black",
                                              "ssp126" = "#33bbee", 
                                              "ssp585" = "#ee3377"), 
                                   name = "Scenarios",
                                   labels = c("Historical", "SSP1-2.6", 
                                              "SSP5-8.5")),
                scale_fill_manual(values = c("historical" = "black", 
                                             "ssp126" = "#33bbee", 
                                             "ssp585" = "#ee3377"), 
                                  name = "Scenarios",
                                  labels = c("Historical", "SSP1-2.6", 
                                             "SSP5-8.5")),
                guides(color = guide_legend(nrow = 1, title.position = "left")),
                theme_bw(),
                labs(y = "Change in exploitable\nfish biomass (%)"),
                theme(legend.position = "top", legend.justification = "center", 
                      legend.text = element_text(size = 10),
                      panel.grid.minor.y = element_blank(), 
                      axis.title.x = element_blank(),
                      axis.title.y = element_text(size = 12),
                      axis.text.x = element_text(angle = 45, vjust = 0.765, 
                                                 hjust = 0.65)))
```

  
## Figure 1 - Trends globally and for top 7 countries with highest catches
Note that data used here correspond to EEZs linked to each country main landmass where areas are not disputed. Overseas territories and disputed areas are not included here. This means that high seas (i.e., areas beyond a country's jurisdiction) were not included in this calculation.  

### Calculating global trends
  
```{r eval = F}
# #Looping through each FishMIP model
# for(m in members){
#   #Load all data available for a single FishMIP model
#   df_model <- str_subset(global_files, m) |> 
#     map(~readRDS(.)) |> 
#     map_df(~bind_rows(.)) |> 
#     #Apply mask
#     left_join(area_grid, by = join_by(x, y)) |> 
#     #Removing areas not recognised in FAO official name list
#     drop_na(fao_official) |> 
#     #Removing FAO Major Areas (2 digit numbers) instead of 4-5 digits for EEZs
#     filter(ID_mrgd > 999) |> 
#     #Calculate mean per ensemble member
#     group_by(year, mem, esm, scenario) |> 
#     summarise(mean_bio = weighted.mean(biomass, area_m, na.rm = T))
#   
#   #Create name to save file  
#   f_out <- file.path(out_folder, 
#                      str_c(m, "_yearly_global_perc_bio_change_data.csv"))
# 
#   #Saving results for each model
#   df_model |>
#     fwrite(f_out)
# }
```
  
Now we can calculate the percentage change in global exploitable fish biomass.  
  
```{r eval = F}
#Listing all relevant files to calculate biomass projections
global_data <- list.files(out_folder, pattern = "_yearly_global_perc_bio_change_data.csv",
                       full.names = T) |> 
  map_df(~fread(.))

#Calculating global reference period
global_ref_bio <- global_data |> 
  filter(year >= 2005 & year <= 2014) |> 
  #Calculate mean per ensemble member
  group_by(mem, esm) |> 
  summarise(ref_bio = mean(mean_bio, na.rm = T))

global_data <- global_data |> 
  #Add reference period to yearly biomass data
  left_join(global_ref_bio, ref_bio, by = c("mem", "esm")) |> 
  #Calculate percentage change per model/year(scenario)
  mutate(perc_change = ((mean_bio-ref_bio)/ref_bio)*100) |> 
  #Calculate annual mean and standard deviation for percentage change 
  group_by(year, scenario) |> 
  summarise(mean_change = mean(perc_change, na.rm = T),
            sd_change = sd(perc_change, na.rm = T))

#Check results
global_data
```
  
### Calculating trends for top 7 countries
The top 7 countries by fishing catches are China (including Taiwan), India, Indonesia, Peru, Russia, United States and Viet Nam. Note that all marine areas identified to be under the jurisdiction of these countries were included in the calculation of trends.  
  
```{r}
#Defining top 7 countries by catch
top_7 <- c("China", "India", "Indonesia", "Peru", "Russian", "United States",
           "Viet")

#Identifying areas to keep from the FAO_EEZ keys
keys_top7 <- fao_eez_keys |> 
  #Keep only top 7 fishing countries
  filter(str_detect(Country_FAO, paste0(top_7, collapse = "|"))) |> 
  #Drop columns that are not relevant
  select(continent:fao_official)

keys_top7<-keys_top7 %>% 
  distinct()

# #Looping through each FishMIP model
# for(m in members){
#   
#   #Load all data available for a single FishMIP model
#   df_model <- str_subset(global_files, m) |> 
#     map(~readRDS(.)) |> 
#     map_df(~bind_rows(.)) |> 
#     #Apply mask
#     left_join(area_grid, by = join_by(x, y)) |> 
#     #Select top 7 producers
#     filter(fao_official %in% keys_top7$fao_official) |> 
#     #Add names in top 7 keys
#     left_join(keys_top7, by = "fao_official") |>
#     #Keep same name for all EEZ that for part of China and the US
#     mutate(fao_official = case_when(str_detect(Country_FAO, "United") ~ 
#                                      "the United States of America",
#                                    Country_FAO == "China" ~ 
#                                      "the People's Republic of China",
#                                    T ~ fao_official)) |> 
#     #Calculate mean per ensemble member
#     group_by(year, mem, esm, scenario, fao_official) |> 
#     summarise(mean_bio = weighted.mean(biomass, area_m, na.rm = T))
#   
#   #Create name to save file  
#   f_out <- file.path(out_folder, 
#                      str_c(m, "_yearly_top7_perc_bio_change_data.csv"))
# 
#   #Saving results for each model
#   df_model |>
#     fwrite(f_out)
# }
```
  
Now we can calculate the percentage change in exploitable fish biomass from the reference period (2005-2014) for the top 7 countries by fisheries catches.  
  
```{r}
#Listing all relevant files to calculate biomass projections
top7_data <- list.files(out_folder, 
                        pattern = "_yearly_top7_perc_bio_change_data.csv",
                       full.names = T) |> 
  map_df(~fread(.))

#Calculating top 7 reference period
top7_ref_bio <- top7_data |> 
  #Select reference period
  filter(year >= 2005 & year <= 2014) |> 
  #Calculate mean per country
  group_by(mem, esm, fao_official) |> 
  summarise(ref_bio = mean(mean_bio, na.rm = T))

top7_base_data <- top7_data |> 
  #Add reference period to yearly biomass data
  left_join(top7_ref_bio, by = c("mem", "esm", "fao_official")) |> 
  #Calculate percentage change per model/year(scenario)/region
  mutate(perc_change = ((mean_bio-ref_bio)/ref_bio)*100) |> 
  #Calculate mean and standard deviation for percentage change 
  #per year(scenario)/region
  group_by(year, scenario, fao_official) |> 
  summarise(mean_change = mean(perc_change, na.rm = T),
            sd_change = sd(perc_change, na.rm = T)) |> 
  ungroup()
```
  
### print table of changes 

```{r}

table_global<-global_data %>% 
  filter(year %in% c(2050, 2100)) %>% 
  select(-sd_change) %>% 
  arrange(scenario, year) %>% 
  relocate(scenario, year, mean_change)

table_top7<-top7_base_data %>% 
  filter(year %in% c(2050, 2100)) %>% 
  select(-sd_change) %>% 
  arrange(fao_official, scenario, year) %>% 
  relocate(fao_official, scenario, year, mean_change)

table_countries<-bio_data %>% 
  filter(year %in% c(2050, 2100)) %>% 
  select(-c(Country_FAO, sd_change)) %>% 
  arrange(continent, fao_official, scenario, year) %>% 
  relocate(continent, OCEAN, fao_official, scenario, year, mean_change)

# # print - change location of files as no access to terminal right now 
# fwrite(table_global, paste0(out_folder,"/table_global.csv"))
# fwrite(table_top7, paste0(out_folder,"/table_top7.csv"))
# fwrite(table_countries, paste0(out_folder,"/table_countries.csv"))
# 
# out_folder
# # Warning - repetition of records in table_countries for e.g. the Federative Republic of Brazil + Chile, Colombia, Equador, USA. No problems when platting.  
# filter(table_countries, fao_official == "the Federative Republic of Brazil")
# filter(bio_data, fao_official == "the Federative Republic of Brazil") # because of Country_FAO column. need to check -> in original table 2 records for Brazil: one is Trinidade. Other places where country not specified.  
# filter(fao_eez_keys, fao_official == "the Federative Republic of Brazil") # calculations are OK as consider fao_official as key, the record gets repeated in line 230 when bio_data get re-merged with fao_eez to consider continent.
# # this coul be problematic at country level (7 top producers) - checked on table and no Country_FAO = NA is part of the 7 top producers besides Hawaii and Alaska (fixed above). though need to adjust table for future analysis. 
# 

```

### PLotting trends for FishMIP Ocean Decede Conference 

```{r}

# copy-paste from above
ODCdata<-global_data |>
  #Add an extra column to match top 7 countries
  mutate(fao_official = "Global", .after = "scenario") |>
  #Joining to top7 data
  bind_rows(top7_base_data) |>
  #Create an order category for SOVEREIGN1
  mutate(fao_official = factor(fao_official,
                               levels = c("Global",
                                          unique(top7_base_data$fao_official)),
                             ordered = T))

ggplot(filter(ODCdata, fao_official == "Global"), aes(x = year, y = mean_change, colour = scenario))+
  #Applying base plot design
  base_gg

ggsave(filename = "figures/global_trends_ODC.pdf",
       device = "pdf", width = 6, height = 3.5)

ggplot(filter(ODCdata, !fao_official %in% c("Global", "the United States of America")), aes(x = year, y = mean_change, colour = scenario))+
  #Applying base plot design
  base_gg+
  facet_wrap(~fao_official, scales = "free_y", nrow = 2,
             labeller = labeller(fao_official = label_wrap_gen(40)))

ggsave(filename = "figures/top_6_global_trends_ODC.pdf",
       device = "pdf", width = 8, height = 4)


### plot error bars FAO ASIA 

bio_dataOCD<-bio_data %>% 
  filter(!is.na(continent),
         year == 2050) %>% 
  select(-OCEAN, -Country_FAO) 

ggplot(filter(bio_dataOCD, continent == "Asia"), aes(x=fao_official, y=mean_change, group=scenario, color=scenario)) + 
  geom_point()+
  geom_errorbar(aes(ymin=mean_change-sd_change, ymax=mean_change+sd_change), width=.2,
                 position=position_dodge(0.05))+
  theme_classic() +
  scale_color_manual(values=c("#33bbee", "#ee3377"))+ 
  labs(x="Country", y = "Change in exploitable fish biomass (%)")+
  coord_flip()

ggsave(filename = "figures/Asia_global_trends_ODC.pdf",
       device = "pdf", width = 8, height = 6)


## Do not run further than this. 

```

### Plotting trends globally and for top 7 countries
  
```{r}
#Merging global and top 7 countries data
global_data |> 
  #Add an extra column to match top 7 countries
  mutate(fao_official = "Global", .after = "scenario") |> 
  #Joining to top7 data
  bind_rows(top7_base_data) |> 
  #Create an order category for SOVEREIGN1
  mutate(fao_official = factor(fao_official, 
                               levels = c("Global", 
                                          unique(top7_base_data$fao_official)), 
                             ordered = T)) |> 
  #Plotting
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  #Applying base plot design
  base_gg+
  #Plotting by country
  facet_wrap(~fao_official, scales = "free_y", nrow = 8)

```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "top_7_global_trends.pdf"), 
       device = "pdf", width = 6, height = 12)
```
  
## Figure 2 - Asia
In this figure, we will exclude marine areas under the jurisdiction of countries in Asia as classified by the FAO.  
  
```{r}
#Filtering data
bio_data |> 
  #Extract EEZs in Asia
  filter(continent == "Asia") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~fao_official, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(fao_official = label_wrap_gen(20)))
```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "asia_perc_change.pdf"), 
       device = "pdf", width = 10, height = 14.35)
```
  
## Figure 3 - Africa
In this figure, we will exclude marine areas under the jurisdiction of countries in Africa as classified by the FAO.  
  
```{r}
#Filtering data
bio_data |> 
  #Extract EEZs in Africa
  filter(continent == "Africa") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~fao_official, scales = "free_y", ncol = 5, 
             #Ensure labels wrap at ~30 characters
             labeller = labeller(fao_official = label_wrap_gen(20)))
```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "africa_perc_change.pdf"), 
       device = "pdf", width = 10, height = 14.35)
```
  
## Figure 4 - Europe
In this figure, we will exclude marine areas under the jurisdiction of countries in Europe as classified by the FAO.  
  
```{r}
#Filtering data
bio_data |> 
  #Extract EEZs in Europe
  filter(continent == "Europe") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~fao_official, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(fao_official = label_wrap_gen(25)))
```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "europe_perc_change.pdf"),
       device = "pdf", width = 12, height = 16)
```
  
## Figure 5 - Americas
In this figure, we will exclude marine areas under the jurisdiction of countries in the Americas as classified by the FAO.  
  
```{r}
#Filtering data
bio_data |> 
  #Extract EEZs in the Americas
  filter(continent == "Americas") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~fao_official, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(fao_official = label_wrap_gen(25)))
```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "americas_perc_change.pdf"), 
       device = "pdf", width = 12, height = 17.22)
```
  
## Figure 6 - Oceania
In this figure, we will exclude marine areas under the jurisdiction of countries in Oceania as classified by the FAO.  
  
```{r}
#Filtering data
bio_data |> 
  #Extract EEZs in Oceania
  filter(continent == "Oceania") |> 
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by country
  facet_wrap(~fao_official, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(fao_official = label_wrap_gen(25)))
```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "oceania_perc_change.pdf"),
       device = "pdf", width = 10, height = 9)
```
  
## Figure 7 - FAO regions
  
```{r}
#Filtering data
bio_data |> 
  #Removing EEZs
  drop_na(OCEAN) |> 
  #Plotting data
  ggplot(aes(x = year, y = mean_change, colour = scenario))+
  base_gg+
  #Plotting by FAO sector
  facet_wrap(~fao_official, scales = "free_y", ncol = 5,
             #Ensure labels wrap at ~30 characters
             labeller = labeller(official = label_wrap_gen(30)))
```
  
```{r, eval = F}
#Saving plot
ggsave(filename = file.path(out_folder, "fao_regions_perc_change.pdf"), 
       device = "pdf", width = 10, height = 6)
```
  

  