---
title: "PLotting biomass vs catches"
author: "Camilla Novaglio"
date: "2024-05-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# dowload BOATS data from DKRZ (06/05/2024)

## tcb nat
scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/historical/boats_ipsl-cm6a-lr_nobasd_historical_nat_default_tcb_global_monthly_1950_2014.nc ./ 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/future/boats_ipsl-cm6a-lr_nobasd_ssp585_nat_default_tcb_global_monthly_2015_2100.nc ./ 

## tcb histsoc / 2015soc
scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/historical/boats_ipsl-cm6a-lr_nobasd_historical_histsoc_default_tcb_global_monthly_1950_2014.nc ./ 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/future/boats_ipsl-cm6a-lr_nobasd_ssp585_2015soc-from-histsoc_default_tcb_global_monthly_2015_2100.nc ./ 

## tc histsoc / 2015soc
scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/historical/boats_ipsl-cm6a-lr_nobasd_historical_histsoc_default_tc_global_monthly_1950_2014.nc ./ 

scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/OutputData/marine-fishery_global/BOATS/ipsl-cm6a-lr/future/boats_ipsl-cm6a-lr_nobasd_ssp585_2015soc-from-histsoc_default_tc_global_monthly_2015_2100.nc ./ 

# Extract data 

```{r cars}

rm(list=ls())
library(raster)
library(ncdf4) 
library(stringr)
library(lubridate)
library(tictoc)
library(dplyr)
library(parallel)

dir <- "/rd/gem/private/users/camillan/FAO_Report/BOATS_Data_for_catch_plot"

unfished_hist = "boats_ipsl-cm6a-lr_nobasd_historical_nat_default_tcb_global_monthly_1950_2014.nc"
unfished_fur = "boats_ipsl-cm6a-lr_nobasd_ssp585_nat_default_tcb_global_monthly_2015_2100.nc"
fished_hist = "boats_ipsl-cm6a-lr_nobasd_historical_histsoc_default_tcb_global_monthly_1950_2014.nc"
fished_fut = "boats_ipsl-cm6a-lr_nobasd_ssp585_2015soc-from-histsoc_default_tcb_global_monthly_2015_2100.nc"
catch_hist = "boats_ipsl-cm6a-lr_nobasd_historical_histsoc_default_tc_global_monthly_1950_2014.nc"
catch_fut = "boats_ipsl-cm6a-lr_nobasd_ssp585_2015soc-from-histsoc_default_tc_global_monthly_2015_2100.nc"

extract<-function(netcdf){
  
  # trial 
  # netcdf = unfished_hist
  
  # extract data 
  brick_data<-brick(file.path(dir, netcdf)) 
  
  # define variable = either tcb_unfished, tcb_fished or tc based on file name 
  if(str_detect(netcdf, "nat.*_tcb_", negate = FALSE)){
    variable = "tcb_unfished"
  }else if (str_detect(netcdf, "histsoc.*_tcb_", negate = FALSE)){
    variable = "tcb_fished"
  }else if (str_detect(netcdf, "_tc_", negate = FALSE)){
    variable = "tc"
  }

  # check units 
  nc_data <- nc_open(file.path(dir, netcdf))

  if(variable %in% c("tcb_unfished", "tcb_fished")){
    units<-ncatt_get(nc_data, "tcb", "units")$value
  }else{
    units<-ncatt_get(nc_data, "tc", "units")$value # SHOULDN'T THIS HAVE A TIME COMPONENT? 
  }
  
  nc_close(nc_data)

  # define time step of output based on file name
  if(str_detect(netcdf, "monthly", negate = FALSE)){
    time_step = "monthly"
  }else if (str_detect(netcdf, "annual", negate = FALSE)){
    time_step = "annual"
  }
  
  # get time vector based on file name
  if(str_detect(netcdf, "1850", negate = FALSE)){
    stTime = "1850-1-1"
    enTime = "2014-12-31"
  }else if (str_detect(netcdf, "1950", negate = FALSE)){
    stTime = "1950-1-1"
    enTime = "2014-12-31"
  }else if (str_detect(netcdf, "2015", negate = FALSE)){
    stTime = "2015-1-1"
    enTime = "2100-12-31"
  }
  
  if(time_step == "monthly"){
    time_step_vector = "month"
  }else if(time_step == "annual"){
    time_step_vector = "year"
  }
  
  t<- as.character(seq(ymd(stTime), ymd(enTime), by = time_step_vector))
  # print(paste("Start time from file name ",t[1], sep = ""))
  # print(paste("End time from file name ",t[length(t)], sep = ""))
  
  # IF TCB DO MEAN ACROSS MONTHS, IF TC DO SUM ACROSS MONTHS?  
  indices<-as.Date(t)
  indices2<-format(indices, format = "%Y")
  indices2<-as.numeric(indices2)
  
  if(variable %in%c("tcb_unfished", "tcb_fished")){
    brick_data_annual<-stackApply(brick_data, indices=indices2, fun=mean)
  }else if (variable == "tc"){
    brick_data_annual<-stackApply(brick_data, indices=indices2, fun=sum)
  }
  
  # GLOBAL SUMS OF TCB AND TC OR WEIGHTED MEANS? HERE GLOBAL SUMS
  # grid cell area in m
  w2 <- raster::area(brick_data_annual[[1]], na.rm = TRUE)*1e6 # could do the IPSL file provided by Matthis but here it is the same
  # value (tcb or tc) * area
  x <- brick_data_annual * w2
  # sum across grid cells 
  global_sum<-cellStats(x, sum, na.rm = TRUE) 
  # transform in daframe
  global_sum<-data.frame(Year = unique(indices2), sum = global_sum) %>%
      mutate(
        Year = as.numeric(Year),
        file = netcdf,
        variable = variable)
  rownames(global_sum)<-NULL
  
  # save result
  return(global_sum = global_sum)

}

# apply function across netcdfs 
netcdf = c(unfished_hist, unfished_fur, fished_hist, fished_fut, catch_hist, catch_fut)
all_bricks<-mclapply(netcdf, function(x) extract(x), mc.cores = detectCores()-2) 
names(all_bricks) <- netcdf

# extract outputs from function above 
global_sum<-do.call("rbind", all_bricks)
rownames(global_sum)<-NULL

library(data.table)
fwrite(global_sum, "data/BOATS_data.csv")

```

# Plot data 

```{r}

library(data.table)
library(ggplot2)

global_sum<-read_csv("data/BOATS_data.csv")

global_sum<-global_sum %>% 
  mutate(sum = (sum/1e+6)/1000000) %>% # from g to tonnes & to million tonnes
  dplyr::select(-file)

# # identify the stable peak in catches 
 peak<-global_sum %>% 
   filter(variable == "tc") %>% 
   arrange(-sum) %>% # 2002-2009
   filter(Year %in% c(2002:2009)) %>% 
   summarise(mean_tc = mean(sum))

# calculate % exploitable biomass at unfished conditions relative to the years of peak in catches
 ref_unfished<-global_sum %>% 
   filter(variable == "tcb_unfished") %>% 
   filter(Year %in% c(2000:2009)) %>% 
   summarise(mean_tcb_unfished = mean(sum))

global_sum2<-global_sum %>% 
  pivot_wider(names_from = variable, values_from = sum) %>% 
  mutate(tcb_unfished_perc = tcb_unfished/unlist(ref_unfished)) 

# multiply peak catch (e.g MSY) by the % change in exploitable fish biomass 

#global_sum2$tc_potential<-global_sum2$tcb_unfished
global_sum2<-global_sum2 %>%
  mutate(tc_potential = unlist(peak)*tcb_unfished_perc)  %>%
  pivot_longer(!Year)


df1<-as.data.frame(filter(global_sum2, name %in% c("tcb_unfished","tcb_fished")))
                   
df2<-as.data.frame(filter(global_sum2, name %in% c("tc","tc_potential")))


p1<-ggplot(df1, aes(x = Year, y = value, group = name, color = name))+
  geom_line()


p2<-ggplot(df2, aes(x = Year, y = value, group = name, color = name))+
  geom_line() + base_gg

# plot catches and biomass. 
base_gg1 <- list(geom_line(linewidth = 0.5),
                #Adding no change line at 0 for reference 
                #geom_hline(yintercept = 0, color = "grey80", linewidth = 0.65, 
                 #          linetype = 2),
                #Adding line dividing historical period and future projections
                geom_vline(xintercept = 2015, color = "grey80", 
                           linewidth = 0.65),
                #Manually setting colours to be used in plots
                scale_color_manual(values = c("tcb_fished" = "#fc8d62", 
                                              "tcb_unfished" = "#8da0cb"), 
                                   name = "Variable",
                                   labels = c("Biomass fished", 
                                              "Biomass unfished")),
                guides(color = guide_legend(nrow = 1, title.position = "left")),
                theme_classic(),
                labs(y = "Million tonnes", 
                     x = ""),
                theme(legend.position = "top", legend.justification = "center", 
                      legend.text = element_text(size = 10),
                      panel.grid.minor.y = element_blank(), 
                      axis.title.x = element_blank(),
                      axis.title.y = element_text(size = 12),
                      axis.text.x = element_text(angle = 45, vjust = 0.765, 
                                                 hjust = 0.65)))

# plot catches and biomass. 
base_gg2 <- list(geom_line(linewidth = 0.5),
                #Adding no change line at 0 for reference 
                #geom_hline(yintercept = 0, color = "grey80", linewidth = 0.65, 
                 #          linetype = 2),
                #Adding line dividing historical period and future projections
                geom_vline(xintercept = 2015, color = "grey80", 
                           linewidth = 0.65),
                #Manually setting colours to be used in plots
                scale_color_manual(values = c("tc" = "#fc8d62", 
                                              "tc_potential" = "#8da0cb"), 
                                   name = "Variable",
                                   labels = c("Catches", 
                                              "Maximum Potential Catches")),
                guides(color = guide_legend(nrow = 1, title.position = "left")),
                theme_classic(),
                labs(y = "Million tonnes per year", 
                     x = ""),
                theme(legend.position = "top", legend.justification = "center", 
                      legend.text = element_text(size = 10),
                      panel.grid.minor.y = element_blank(), 
                      axis.title.x = element_blank(),
                      axis.title.y = element_text(size = 12),
                      axis.text.x = element_text(angle = 45, vjust = 0.765, 
                                                 hjust = 0.65)))


# p<-ggplot(global_sum, aes(x = Year, y = sum, group = variable, color = variable))+
#   geom_line()+
#   base_gg
p1<-p1 + base_gg1
p2<-p2 +base_gg2

library(patchwork)
 combined <- p1 + p2 +
   plot_layout(guides = "collect") & theme(legend.position = 'bottom') 

# 
# # combine all in one figure
# fig_scen <- plot_grid(plot_grid(p1, p2, ncol = 2, nrow = 1), 
#                       plot_grid(NULL, NULL, nrow = 1, rel_widths = c(0.5, 1, 0.5)),
#                       ncol = 1, nrow = 2, rel_heights = c(1, 0.1))

#### NEED TO FIX LEGENF & ADD LABELS
 
 
 
# png("figures/BOATS_trends.png", width=5, height=3.5, units="in", res=1200)
# p
# dev.off()

```
