---
title: "Temporal biomass change at the country level (country AND territories combined)"
author: "Gage"
date: "2024-01-17"
output: pdf_document
---

## Loading libraries
  
```{r, message=FALSE, warning=FALSE}
#Data wrangling
library(tidyverse)
library(data.table)
#Raster data
library(terra)
#Combining plots
library(cowplot)

# filepaths
library(here)
# getting iso3c codes
library(countrycode)
```

## Setting up notebook
  
```{r}
#Base folder for project
base_folder <- "/rd/gem/private/users/camillan"

#Defining location of notebook outputs
out_folder <- file.path(base_folder, "FAO_Report")
if(!dir.exists(out_folder)){
  dir.create(out_folder)
}

country_level_folder <- file.path(out_folder, "country_level_gage")
```
  
## Files with all information per grid cell
  
```{r}
#Getting GFDL grid area
area_gfdl_df <- list.files(out_folder, "gfdl-esm4_60arcmin_global_FAO-EEZ.csv",
                           recursive = T, full.names = T) |> 
  read_csv()

#Getting IPSL grid area
area_ipsl_df <- list.files(out_folder, "ipsl-cm6a_60arcmin_global_FAO-EEZ.csv",
                           recursive = T, full.names = T) |> 
  read_csv()

```

## Adjust admin country names as per FAO short names

```{r}

### GAGE approach ----

#EEZ keys
# eez_keys <- read_csv(file.path(out_folder, "Merged_FAO_EEZ/supporting_files/EEZ_FAO_key.csv")) # CN not neede but if you want to use this, it is now:
# eez_keys<-read_csv("/rd/gem/private/users/camillan/FAO_Report/Merged_FAO_EEZ/FAO-Major-Areas_EEZ_continent_keys.csv")

# admin_keys <- read.csv(here("data/country_admin_key.csv")) %>%
#   filter(continent != "") %>% # there was blank space at bottom for some reason
#   mutate(iso3c_admin = countrycode(sourcevar = admin_name, origin = "country.name", destination = "iso3c")) %>% ## ok cool
#   filter(!is.na(iso3c_admin)) %>% # filter out disputed areas
#   dplyr::select(-continent) # remove continent as it is already in the area_grid below

### CN approach ----
#get info from the masks area_gfdl_df and area_ipsl_df built using EEZ nad FAO info

fix_masks<-function(mask){
  
  # mask = area_gfdl_df
  
  # fix mistakes in file 
  # filter(mask, figure_name == "Comoros") # country == the Union of the Comoros
  mask<-mask %>% 
    mutate(Country_FAO = case_when(figure_name == "Comoros" ~ "the Union of the Comoros",
                                   .default = as.character(Country_FAO)),
          figure_name = ifelse(figure_name == "Prince Edward Island", "Prince Edward Islands", figure_name))

  mask<-mask %>% 
    drop_na(fao_official) # remove all FAO fishing areas 

  # 1. fix names from long to short 

  # # check Congo and Democratic Republic of the Congo as both needs to be there
  # unique(filter(mask, str_detect(Country_FAO, "Congo"))$Country_FAO)
  # unique(filter(mask, str_detect(Country_FAO, "the Republic of the "))$Country_FAO)
  # unique(filter(mask, str_detect(Country_FAO, "the Democratic Republic of "))$Country_FAO)
  # unique(filter(mask, str_detect(Country_FAO, "Korea"))$Country_FAO)
  

  # need to do this manually. Change here and back-adjust below
  mask<-mask %>% 
    mutate(Country_FAO = case_when(Country_FAO == "the Republic of the Congo" ~ "Congo rep",
                                   Country_FAO == "the Democratic Republic of the Congo" ~ "Congo dem",
                                   Country_FAO == "the Democratic People's Republic of Korea" ~ "Korea dem",
                                   Country_FAO == "the Republic of Korea"~ "Korea rep",
                                   .default = as.character(Country_FAO)))

  # remove" long part of" names 
  mask<-mask %>% 
    mutate(Country_FAO = gsub("the Republic of |The Kingdom of |the Commonwealth of |the Democratic Republic of |the Eastern Republic of |the Federal Republic of |the Federated States of |the Federative Republic of |the Islamic Republic of |the Socialist Republic of |the State of |the Sultanate of |the Kingdom of |the |Arab Republic of |Bolivarian Republic of |Democratic Socialist Republic of |People's Democratic Republic of |Principality of |Co-operative Republic of |Hashemite Kingdom of |People's Republic of |Union of |Democratic |Independent State of |the Union of the ", '',Country_FAO))

  # adjust exceptions - These cannot be changed according to FAO rules - e.g. "Iran (Islamic Republic of)"
  mask<-mask %>% 
    mutate(Country_FAO = case_when(Country_FAO == "Argentine Republic" ~ "Argentina",
                                   Country_FAO == "French Republic" ~ "France",
                                   Country_FAO == "Gabonese Republic" ~ "Gabon",
                                   Country_FAO == "Hellenic Republic" ~ "Greece",
                                   Country_FAO == "Iran" ~ "Iran (Islamic Republic of)",
                                   Country_FAO == "Micronesia" ~ "Micronesia (Federated States of)",
                                   Country_FAO == "Netherlands" ~ "Netherlands (Kingdom of the)",
                                   Country_FAO == "Venezuela" ~ "Venezuela (Bolivarian Republic of)",
                                   Country_FAO == "Lebanese Republic" ~ "Lebanon",
                                   Country_FAO == "Portuguese Republic" ~ "Portugal", 
                                   Country_FAO == "Saint Vincent and Grenadines" ~ "Saint Vincent and the Grenadines",
                                   Country_FAO == "Togolese Republic" ~ "Togo",
                                   Country_FAO == "United Mexican States" ~ "Mexico",
                                   # readjust some names that have been changed above 
                                   Country_FAO == "Congo rep" ~ "Congo", 
                                   Country_FAO == "Congo dem" ~ "Democratic Republic of the Congo", 
                                   Country_FAO == "Korea rep" ~ "Republic of Korea", 
                                   Country_FAO == "Korea dem" ~ "Democratic People's Republic of Korea", 
                                   .default = as.character(Country_FAO)))

  setdiff(sort(unique(mask$Country_FAO)), sort(unique(mask$figure_name))) # Kiribati should be the only one

  # 2. add country to missing ones as per Table A4
  # sort(unique(filter(mask, is.na(Country_FAO))$figure_name))

  # remove disputed
  mask<-mask %>% 
    filter(!figure_name == "Juan de Nova Island")

  mask<-mask %>% 
    mutate(Country_FAO = case_when(figure_name == "Ascension" ~ "United Kingdom of Great Britain and Northern Ireland",
                                   figure_name == "Pitcairn" ~ "United Kingdom of Great Britain and Northern Ireland",
                                   figure_name == "Saint Helena" ~ "United Kingdom of Great Britain and Northern Ireland",
                                   figure_name == "Tristan da Cunha" ~ "United Kingdom of Great Britain and Northern Ireland",
                                   figure_name == "Kerguelen Islands" ~ "France",
                                   figure_name == "Norfolk Island" ~ "Australia",
                                   figure_name == "Réunion" ~ "France",
                                   figure_name == "Prince Edward Islands" ~ "South Africa",
                                   figure_name == "Bassas da India" ~ "France", 
                                   figure_name == "British Virgin Islands" ~ "United Kingdom of Great Britain and Northern Ireland",
                                   figure_name == "Cayman Islands" ~ "United Kingdom of Great Britain and Northern Ireland",
                                   figure_name == "Curaçao" ~ "Netherlands (Kingdom of the)",
                                   figure_name == "Faroe Islands" ~ "Denmark", 
                                   figure_name == "Puerto Rico" ~ "United States of America", 
                                   figure_name == "Turks and Caicos Islands" ~ "United Kingdom of Great Britain and Northern Ireland", 
                                   figure_name == "Guernsey" ~ "United Kingdom of Great Britain and Northern Ireland", 
                                   figure_name == "Andaman and Nicobar Islands" ~ "India", 
                                   figure_name == "Wallis and Futuna Islands" ~ "France", 
                                   figure_name == "Svalbard Islands" ~ "Norway", 
                                   figure_name == "Jan Mayen Island" ~ "Norway", 
                                   figure_name == "Niue" ~ "New Zealand",
                                   figure_name == "Cocos (Keeling) Islands" ~ "Australia",
                                   figure_name == "Heard Island and McDonald Islands" ~ "Australia",
                                   # countries accidentally left out in manual merging of FAO EEZ files - not a problem for other analysis
                                   figure_name == "Cabo Verde" ~ "Cabo Verde",
                                   figure_name == "Côte d'Ivoire" ~ "Côte d'Ivoire",
                                   figure_name == "Libya" ~ "Libya", 
                                   figure_name == "Türkiye" ~ "Türkiye", 
                                   figure_name == "Papua New Guinea" ~ "Papua New Guinea", 
                                   .default = as.character(Country_FAO)))

  sort(unique(filter(mask, is.na(Country_FAO))$figure_name))
  # these should all be territories 
  setdiff(sort(unique(mask$figure_name)),sort(unique(mask$Country_FAO))) # 67

  # trial<-mask %>% 
  #   mutate(keep = ifelse(Country_FAO == figure_name, "no", "yes")) %>% 
  #   filter(keep == "yes") %>% 
  #   dplyr::select(Country_FAO, figure_name) %>% 
  #   relocate(figure_name,Country_FAO) %>% 
  #   unique()
  # View(trial)

  return(mask)

}

area_gfdl_df<-fix_masks(area_gfdl_df)
area_ipsl_df<-fix_masks(area_ipsl_df)

```

## Getting FishMIP output files

```{r}
#Getting a list of files containing biomass data
global_files <- list.files(base_folder, 
                           pattern = "2024-06-03.*global_10g-100kg.rds", 
                           full.names = T, recursive = T)

#Getting a list of models
members <- str_extract(global_files, "annual_(.*)_(h|s)", group = 1) |> 
  unique()

```
  
## Calculating percentage change in fish biomass mean per country from global FishMIP models at Admin country level
  
```{r, eval = F}
#Looping through each FishMIP model
for(m in members){
  
 # m = members[[2]]
  
  #Select correct area of grid cell
  if(str_detect(m, "ipsl")){
    area_grid <- area_ipsl_df
    print(paste0("processing '", m, "' file, using ipsl mask"))
  }else if(str_detect(m, "gfdl")){
    area_grid <- area_gfdl_df
    print(paste0("processing '", m, "' file, using gfdl mask"))
  }
  
 ## GAGE appraoch ---- 
  # area_grid <- area_grid %>%
  #   mutate(Country_FAO = case_when(
  #     fao_official == "the Republic of Türkiye" ~ "Türkiye",
  #     fao_official == "the State of Libya" ~ "Libya", 
  #     fao_official == "the Republic of Côte d'Ivoire" ~ "Côte d'Ivoire",
  #     fao_official == "the Independent State of Papua New Guinea" ~ "Papua New Guinea", 
  #     fao_official == "the Commonwealth of Puerto Rico" ~ "United States of America",
  #     fao_official == "the Republic of Cabo Verde" ~ "Cabo Verde",
  #     TRUE ~ Country_FAO
  #   )) %>%
  #   left_join(admin_keys, by = c("figure_name" = "country_territory")) 
  
  # test <- distinct(area_grid, continent, fao_official, Country_FAO, figure_name, admin_name, iso3c_admin) %>%
  #   filter(!is.na(continent))
  # 
  # ## ok we're missing 3, which 3? It is ok actually, I checked the other members and they are in there... so should be fine! 
  # setdiff(admin_keys$country_territory, test$figure_name) # [1] "Iraq"    "Jordan"  "Albania"
  
  # #Load all data available for a single FishMIP model
  # df_model <- str_subset(global_files, m) |>
  #   map(~readRDS(.)) |>
  #   map_df(~bind_rows(.)) |>
  #   #Add area for each grid cell and area ID (FAO and EEZ)
  #   left_join(area_grid, by = join_by(x, y)) |>
  #   #Removing areas not recognised in FAO official name list
  #   drop_na(fao_official) |>
  #   #Calculate mean per ensemble member
  #   group_by(year, mem, esm, scenario, admin_name) |> 
  #   summarise(mean_bio = weighted.mean(biomass, cell_area, na.rm = T)) %>%
  #   ungroup() %>%
  #   rename(Country_FAO = admin_name)
  # 
  # #Create name to save file
  # f_out <- file.path(country_level_folder, str_c(m, "_yearly_perc_bio_change_admin.csv"))
  # 
  # #Saving results for each model
  # df_model |>
  #   fwrite(f_out)
 
  #### 03a code ----
 
  #Keep only information for grid cells with countries recognised by FAO
  #Extracting data from countries calculations
  area_grid <- area_grid |> 
    dplyr::select(x:cell_area, Country_FAO)
  
  #Load all data available for a single FishMIP model
  df_model <- str_subset(global_files, m) |>
    map(~readRDS(.)) |>
    map_df(~bind_rows(.)) |>
    #Add area for each grid cell and country
    left_join(area_grid, by = join_by(x, y)) |>
    drop_na(Country_FAO) |> 
    #Calculate weighted mean for biomass per ensemble member per country
    group_by(year, mem, esm, scenario, Country_FAO) |> 
    summarise(mean_bio = weighted.mean(biomass, cell_area, na.rm = T)) %>% 
    ungroup()
  
  # sort(unique(df_model$Country_FAO))
  # sort(unique(area_grid$Country_FAO))
  
  #Calculating reference period
  ref_bio <- df_model |> 
    filter(year >= 2005 & year <= 2014) |> 
    group_by(Country_FAO) |> 
    #Calculate mean per ensemble member
    summarise(ref_bio = mean(mean_bio, na.rm = T)) %>% 
    ungroup()
  
  #Add reference period to weighted mean data frame
  df_model <- df_model |> 
    left_join(ref_bio, join_by(Country_FAO)) |> 
    #Calculate percentage change
    mutate(perc_change = ((mean_bio-ref_bio)/ref_bio)*100)
  
  #Create name to save file
  f_out <- file.path(country_level_folder, 
                     str_c(m, "_yearly_perc_bio_change_admin.csv"))

  #Saving results for each model
  df_model |>
    fwrite(f_out)
 
}

```
  
## Calculating mean for reference period
  
```{r eval = F}

## CN moved in function above

# #Listing all relevant files to calculate biomass projections
# bio_data <- list.files(country_level_folder, 
#                        pattern = "_yearly_perc_bio_change_admin.csv",
#                        full.names = T) |> 
#   map_df(~fread(.)) %>%
#   filter(Country_FAO != "")
# 
# unique(bio_data$Country_FAO) #149
# 
# #Calculating reference period
# ref_bio <- bio_data |> 
#   filter(year >= 2005 & year <= 2014) |> 
#   #Calculate mean per ensemble member
#   group_by(mem, esm, Country_FAO) |> 
#   dplyr::summarise(ref_bio = mean(mean_bio, na.rm = T)) %>%
#   ungroup()
# 
# # some are repetition if ID_mgr is removed 
# fao_eez <-  eez_keys |>
#   select(!ID_mrgd) |>
#   distinct()
# # nrow(fao_eez)
# 
# fao_eez <- fao_eez %>% 
#   filter((is.na(continent) & is.na(Country_name))|(!is.na(continent) & !is.na(Country_name))) 
# # nrow(fao_eez)
# 
# # calculate trends by model as needed for stat table 
# bio_data_by_model <- bio_data |>
#   #Add reference period to yearly biomass data
#   left_join(ref_bio, by = c("mem", "esm", "Country_FAO")) |>
#   #Calculate percentage change per model/year(scenario)/region
#   mutate(perc_change = ((mean_bio-ref_bio)/ref_bio)*100) %>%
#  select(-mean_bio, -ref_bio) 
# 
# #Check results
# bio_data_by_model
# 
# bio_data <- bio_data |> 
#   #Add reference period to yearly biomass data
#   left_join(ref_bio, by = c("mem", "esm", "Country_FAO")) |> 
#   #Calculate percentage change per model/year(scenario)/region
#   mutate(perc_change = ((mean_bio-ref_bio)/ref_bio)*100) |> 
#   #Calculate mean and standard deviation for percentage change 
#   #per year(scenario)/region
#   group_by(year, scenario, Country_FAO) |> 
#   dplyr::summarise(mean_change = mean(perc_change, na.rm = T),
#             sd_change = sd(perc_change, na.rm = T)) |> 
#  ungroup() 
# 
# #Check results
# bio_data

```
  
## We can now save the results for future reference.  
  
```{r}
## CN moved above/below

# #Create name to save file
# f_out <- file.path(country_level_folder, "ensemble_perc_bio_change_admin_all.csv")
# 
# #Saving results for each model
# bio_data |>
#   write_csv(f_out)
# 
# #Create name to save file
# f_out <- file.path(country_level_folder, "mem_perc_bio_change_admin_all.csv")
# 
# #Saving results for each model
# bio_data_by_model |>
#   write_csv(f_out)

```
  
## Calculating mean ensemble percentage change in biomass for top 7 countries
  
```{r eval = F}
bio_data_by_model <- list.files(country_level_folder, 
                       pattern = "_yearly_perc_bio_change_admin.csv",
                       full.names = T, recursive = T) |> 
  map(~fread(.)) |> 
  bind_rows()

#Create name to save file
f_out <- file.path(country_level_folder, "mem_perc_bio_change_admin_all.csv")

#Saving results for each model
bio_data_by_model |>
  write_csv(f_out)

# bio_data <- bio_data_by_model |>
#   group_by(year, scenario, Country_FAO) |> 
#   summarise(mean_change = mean(perc_change, na.rm = T),
#             sd_change = sd(perc_change, na.rm = T))
# 
# #Create name to save file
# f_out <- file.path(country_level_folder, "ensemble_perc_bio_change_admin_all.csv")
# 
# #Saving results for each model
# bio_data |>
#   write_csv(f_out)

```
    
## Table of stats, model agreement and wilcox.test   

```{r}

#mem percentage change in biomass by admin countries
admin_count_bio_by_model <- list.files(country_level_folder, 
                                 "mem_perc_bio_change_admin_all.csv",
                                 recursive = T, full.names = T) |> 
  read_csv()

admin_count_bio_by_model <- admin_count_bio_by_model %>% 
  filter(year %in% c(2041:2050, 2091:2100)) %>% 
  mutate(decade = ifelse(year %in% c(2041:2050), "2041-2050", "2091-2100")) %>% 
  filter(!is.na(perc_change)) # some models do not output info on some countries (see checks.R) 

### mean, median, sd, min and max, agreement, wilcox ----
stats <- admin_count_bio_by_model %>% 
  group_by(Country_FAO, mem, esm, scenario, decade) %>% 
  summarise(mean_across_decade = mean(perc_change, na.rm = T)) %>%  # across decade to have one (mean) value of change per model and exclude temporal variability from the calculations below 
  ungroup() %>% 
  group_by(Country_FAO, scenario, decade) %>% 
  summarise(
    mean = mean(mean_across_decade), # then across model 
    median = median(mean_across_decade),
    sd = sd(mean_across_decade),
    min = min(mean_across_decade),
    max = max(mean_across_decade),
    n_model = n(), # to calculate agreement again across models
    n_positive = sum(mean_across_decade >= 0),
    n_negative = sum(mean_across_decade < 0), 
    same_sign = max(n_positive,n_negative),
    agreement = (same_sign/n_model)*100) %>% 
  ungroup() %>% 
  dplyr::select(-c(n_positive, n_negative, same_sign))

#### wilcox test ----

w_test <- admin_count_bio_by_model %>% 
  dplyr::select(!c(mean_bio, ref_bio)) %>% 
  spread(scenario, perc_change) %>% 
  group_by(Country_FAO, decade) %>%
  summarise(wilcox_p = wilcox.test(ssp126, ssp585, alternative = "two.sided")$p.value) %>% 
  ungroup()

table <- stats %>% 
  full_join(w_test)

fwrite(table, paste0(country_level_folder,"/table_stats_country_admin.csv"))
fwrite(table, here("data/table_stats_country_admin_CN.csv"))

```
