---
title: "Temporal biomass change at the country level (country and territories combined)"
author: "Gage Clawson and Denisse Fierro Arcos"
date: "2026-05-13"
output: pdf_document
---

## Loading libraries
  
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(data.table)
```

## Setting up notebook
  
```{r}
#Base folder for project
base_folder <- "/rd/gem/private/users/camillan"

#Defining location of notebook outputs
out_folder <- file.path(base_folder, "FAO_Report")

country_level_folder <- file.path(out_folder, "country_level_gage")
if(!dir.exists(country_level_folder)){
  dir.create(country_level_folder)
}
```
  
## Files with all information per grid cell
  
```{r}
#Getting GFDL grid area
area_gfdl_df <- list.files(out_folder, "gfdl-esm4_60arcmin_global_FAO-EEZ.csv",
                           recursive = T, full.names = T) |> 
  fread()

#Getting IPSL grid area
area_ipsl_df <- list.files(out_folder, "ipsl-cm6a_60arcmin_global_FAO-EEZ.csv",
                           recursive = T, full.names = T) |> 
  fread()
```

## Adjusting admin country names as per FAO short names
```{r}
# Get info from the masks area_gfdl_df and area_ipsl_df built using EEZ and 
# FAO info
fix_masks <- function(mask){
  # Editing country names in masks to match FAO short names
  mask <- mask |>
    # remove all grid cells outside EEZs
    drop_na(fao_official) |>
    mutate(Country_FAO = str_remove_all(Country_FAO, "^the "),
           Country_FAO = ifelse(Country_FAO == "Republic of the Congo", "Congo",
                                Country_FAO),
           Country_FAO =
             str_remove(Country_FAO,
                        paste("^Republic of ", "^Kingdom of ", "^State of ",
                              "^H[a-z]+ Kingdom of ", "Independent State of ",
                              "^Democratic Republic of ", "Sultanate of ",
                              "[A|B|E|F|I|P|S][a-z]+ Republic of ",
                              "Commonwealth of ", "^[P,C].* Republic of ",
                              "People's Democratic Republic of ",
                              "^Democratic Socialist Republic of ",
                              sep = "|")),
           Country_FAO = 
             ifelse(Country_FAO == "the Congo", 
                    "Democratic Republic of the Congo", Country_FAO),
           Country_FAO = str_remove_all(Country_FAO, "^the |Union of "),
           Country_FAO = case_when(
             figure_name == "Comoros" ~ "Comoros",
             figure_name == "Prince Edward Island" ~ "South Africa",
             figure_name == "Cabo Verde" ~ "Cabo Verde",
             figure_name == "Côte d'Ivoire" ~ "Côte d'Ivoire",
             # The following names cannot be changed according to FAO rules:
             # e.g. "Iran (Islamic Republic of)"
             Country_FAO == "Argentine Republic" ~ "Argentina",
             Country_FAO == "French Republic" ~ "France",
             Country_FAO == "Gabonese Republic" ~ "Gabon",
             Country_FAO == "Hellenic Republic" ~ "Greece",
             Country_FAO == "Federated States of Micronesia" ~
               "Micronesia (Federated States of)",
             Country_FAO == "Iran" ~ "Iran (Islamic Republic of)",
             Country_FAO == "Korea" ~ "Republic of Korea",
             Country_FAO == "Lebanese Republic" ~ "Lebanon",
             figure_name == "Libya" ~ "Libya",
             Country_FAO == "United Mexican States" ~ "Mexico",
             Country_FAO == "Netherlands" ~ "Netherlands (Kingdom of the)",
             figure_name == "Papua New Guinea" ~ "Papua New Guinea",
             Country_FAO == "Portuguese Republic" ~ "Portugal",
             Country_FAO == "Togolese Republic" ~ "Togo",
             figure_name == "Türkiye" ~ "Türkiye",
             Country_FAO == "Venezuela" ~ "Venezuela (Bolivarian Republic of)",
             # adding country names to territories where no country has been
             # assigned as per Table A4
             figure_name %in% c("Ascension", "Pitcairn", "Saint Helena",
                                "Tristan da Cunha", "British Virgin Islands",
                                "Cayman Islands", "Turks and Caicos Islands",
                                "Guernsey") ~
               "United Kingdom of Great Britain and Northern Ireland",
             figure_name %in% c("Wallis and Futuna Islands", "Réunion",
                                "Bassas da India", "Kerguelen Islands") ~
               "France",
             figure_name %in% c("Norfolk Island", "Cocos (Keeling) Islands",
                                "Heard Island and McDonald Islands") ~
               "Australia",
             figure_name %in% c("Svalbard Islands", "Jan Mayen Island") ~
               "Norway",
             figure_name == "Curaçao" ~ "Netherlands (Kingdom of the)",
             figure_name == "Faroe Islands" ~ "Denmark",
             figure_name == "Puerto Rico" ~ "United States of America",
             figure_name == "Andaman and Nicobar Islands" ~ "India",
             figure_name == "Niue" ~ "New Zealand",
             .default = as.character(Country_FAO))) |>
    # removing disputed territories
    filter(!figure_name == "Juan de Nova Island")

  return(mask)
}

area_gfdl_df <- fix_masks(area_gfdl_df)
area_ipsl_df <- fix_masks(area_ipsl_df)
```

## Getting FishMIP output files

```{r}
#Getting a list of files containing biomass data
global_files <- list.files(base_folder, 
                           pattern = "2024-06-03.*global_10g-100kg.rds", 
                           full.names = T, recursive = T)

#Getting a list of models
members <- str_extract(global_files, "annual_(.*)_(h|s)", group = 1) |> 
  unique()
```
  
## Calculating mean percentage change in fish biomass per country from global FishMIP MEM outputs
  
```{r, eval = F}
#Looping through each FishMIP model
for(m in members){
  
  #Select correct area of grid cell
  if(str_detect(m, "ipsl")){
    area_grid <- area_ipsl_df
    print(paste0("processing '", m, "' file, using ipsl mask"))
  }else if(str_detect(m, "gfdl")){
    area_grid <- area_gfdl_df
    print(paste0("processing '", m, "' file, using gfdl mask"))
  }

  #Keep only information for grid cells with countries recognised by FAO
  #Extracting data from countries calculations
  area_grid <- area_grid |> 
    dplyr::select(x:cell_area, Country_FAO)
  
  
  #Load all data available for a single FishMIP model
  df_model <- str_subset(global_files, m) |>
    map(~readRDS(.)) |>
    map_df(~bind_rows(.)) |>
    #Add area for each grid cell and country
    left_join(area_grid, by = join_by(x, y)) |>
    drop_na(Country_FAO) |> 
    #Calculate weighted mean for biomass per ensemble member per country
    group_by(year, mem, esm, scenario, Country_FAO) |> 
    summarise(mean_bio = weighted.mean(biomass, cell_area, na.rm = T)) |> 
    ungroup()
  
  #Calculating reference period
  ref_bio <- df_model |> 
    filter(year >= 2005 & year <= 2014) |> 
    #Calculate mean per ensemble member
    summarise(ref_bio = mean(mean_bio, na.rm = T), .by = Country_FAO)
  
  #Add reference period to weighted mean data frame
  df_model <- df_model |> 
    left_join(ref_bio, join_by(Country_FAO)) |> 
    #Calculate percentage change
    mutate(perc_change = ((mean_bio-ref_bio)/ref_bio)*100)
  
  #Create name to save file
  f_out <- file.path(country_level_folder, 
                     str_c(m, "_yearly_perc_bio_change_admin.csv"))

  #Saving results for each model
  df_model |>
    fwrite(f_out)
}
```
  
## Calculating mean ensemble percentage change in biomass per country
  
```{r eval = F}
bio_data_by_model <- list.files(country_level_folder, 
                                pattern = "_yearly_perc_bio_change_admin.csv",
                                full.names = T, recursive = T) |> 
  map(~fread(.)) |> 
  bind_rows()

#Create name to save file
f_out <- file.path(country_level_folder, "mem_perc_bio_change_admin_all.csv")

#Saving results for each model
bio_data_by_model |>
  fwrite(f_out)
```
    
## Table of stats, model agreement and wilcox.test   

```{r}
# mem percentage change in biomass by admin countries
admin_count_bio_by_model <- list.files(country_level_folder, 
                                       "mem_perc_bio_change_admin_all.csv",
                                       recursive = T, full.names = T) |> 
  fread() |> 
  filter(year %in% c(2041:2050, 2091:2100)) |> 
  mutate(decade = ifelse(year %in% c(2041:2050), "2041-2050", "2091-2100")) |> 
  # some models do not output info on some countries (see checks.R) 
  filter(!is.na(perc_change)) 

# mean, median, sd, min and max, agreement
stats <- admin_count_bio_by_model |> 
  group_by(Country_FAO, mem, esm, scenario, decade) |> 
  summarise(mean_across_decade = mean(perc_change, na.rm = T)) |> 
  # across decade to calculate mean change per model and exclude temporal 
  # variability from the calculations below 
  ungroup() |> 
  group_by(Country_FAO, scenario, decade) |> 
  # then across model 
  summarise(
    mean = mean(mean_across_decade), 
    median = median(mean_across_decade),
    sd = sd(mean_across_decade),
    min = min(mean_across_decade),
    max = max(mean_across_decade),
    n_model = n(), # to calculate agreement again across models
    n_positive = sum(mean_across_decade >= 0),
    n_negative = sum(mean_across_decade < 0), 
    same_sign = max(n_positive,n_negative),
    agreement = (same_sign/n_model)*100) |> 
  ungroup() |> 
  dplyr::select(-c(n_positive, n_negative, same_sign))

# wilcox test 
w_test <- admin_count_bio_by_model |> 
  dplyr::select(!c(mean_bio, ref_bio)) |> 
  pivot_wider(names_from = scenario, values_from = perc_change) |> 
  group_by(Country_FAO, decade) |>
  summarise(wilcox_p = wilcox.test(ssp126, ssp585, alternative = "two.sided",
                                   exact = F)$p.value) |> 
  ungroup()

table_stats <- stats |> 
  full_join(w_test) |> 
  mutate(spatial_scale = "countries_admin")

# Save results
fwrite(table_stats, file.path(out_folder, "table_stats_country_admin.csv"))
```
