---
title: "Global plots for FISHMIP"
output: html_document
date: "2024-05-24"
author: "Gage" 
---

## Summary

 - In this script we make circular bar charts of change in mean biomass at multiple spatial scales: 
    - Administrative country
    - EEZ
    - Administrative country and FAO high seas regions
    - EEZs and FAO high seas regions
 - Additionally, we create maps of the mean change in those same cateogories

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(sf)
library(glue)
library(cowplot)
library(ggpattern)
library(countrycode)
library(tidyterra)
library(terra)
library(smoothr)
library(rnaturalearth)


options(scipen = 999) ## To disable scientific notation
#Base folder for project
base_folder <- "/rd/gem/private/users/camillan"

#Defining location of notebook outputs
out_folder <- file.path(base_folder, "FAO_Report")

```

Read in data

```{r}

##### CN: OLD FILE - new version: "/rd/gem/private/users/camillan/FAO_Report/table_stats.csv"

#cmip_data <- read.csv(here("data/table_stats_with_continent.csv")) %>%
  
  cmip_data <- read.csv(file.path(base_folder, "FAO_Report/table_stats.csv")) %>%
    mutate(pos_neg = ifelse(mean > 0, "positive", "negative")) %>% # ~200 regions, makes sense
  mutate(uncertain = ifelse(
    agreement < 80, "yes", "no"
  )) %>%
  filter(!is.na(mean)) 

country_continent <- cmip_data %>%
  distinct(figure_name, continent)

high_seas_df <- cmip_data %>%
  filter(spatial_scale == "FAO_area") %>%
  dplyr::select(figure_name, scenario, decade, continent, mean, median, sd, min, max, n_model, agreement, wilcox_p, spatial_scale)


##### CN: OLD FILE - new version: "/rd/gem/private/users/camillan/FAO_Report/Merged_FAO_EEZ/FAO-Major-Areas_EEZ_continent_keys.csv"

regions_df_2 <- read_csv(file.path(out_folder, "Merged_FAO_EEZ/FAO-Major-Areas_EEZ_continent_keys.csv"))

##### CN: NOT SURE ABOUT THIS FILE and how it was created


####### NEW VERSION PRINTED IN biomass_change_adminsitrative_region.Rmd: "data/table_stats_country_admin_CN.csv", some of the steps below might not be needed anymore as country anmes were adjusted there as well as territories with no country specified.  

cmip_data_new <- read.csv(here("data/table_stats_country_admin.csv")) %>%
    dplyr::rename(figure_name = Country_FAO) %>%
  left_join(country_continent) %>%
  mutate(continent = case_when(
    figure_name == "China" ~ "Asia", 
    figure_name == "India" ~ "Asia",
    figure_name == "Indonesia" ~ "Asia", 
    figure_name == "Peru" ~ "Americas", 
    figure_name == "Russian Federation" ~ "Europe",
    figure_name == "United States of America" ~ "Americas",
    figure_name == "Viet Nam" ~ "Asia",
    figure_name == "Kiribati" ~ "Oceania",
    TRUE ~ continent
  )) # fix missing continents

  
length(unique(cmip_data_new$Country_FAO)) # 148


cmip_data_all <- cmip_data_new %>%
  rbind(high_seas_df %>% dplyr::select(colnames(cmip_data_new))) %>%
  rbind(cmip_data %>% filter(spatial_scale == "countries") %>% dplyr::select(colnames(cmip_data_new))) %>% mutate(pos_neg = ifelse(mean > 0, "positive", "negative"))  %>%
  mutate(uncertain = ifelse(
    agreement < 80, "yes", "no"
  ))

## ok so we have 3 spatial scales: countries (eezs), countries_admins (administrative/sovereign nations), and FAO_area (high seas area)
  
```


Make stacked circle bar chart, similar to that of the circle plot presented here: https://www.nature.com/articles/s41893-022-00965-x/figures/7 
 - Make one with categorical fill for <10%, 10-20%, 20-30%, and >30% change, but with a pattern for uncertainty (if sd > mean). Save for High and Low emissions scenarios.
    - NOTE: the pattern takes awhile to run ~10 mins per scenario.


```{r}
scenarios <- c("ssp585", "ssp126")
spatial_scales <- c("countries_high_seas", "countries_admin", "countries_admin_high_seas", "countries")
label_types <- c("country_name") # we removed iso3c, so now only plotting country names


for(label_type in label_types){
for(spatial_type in spatial_scales){ # specify what spatial scales you want to include in plot

# label_type = "country_name"
# spatial_type = "countries_admin"
  
  ## filter for spatial type
  if(spatial_type == "countries_admin_high_seas"){
  
cmip_data_plot <- cmip_data_all %>% 
  filter(spatial_scale %in% c("countries_admin", "FAO_area"))

  }else if(spatial_type == "countries_admin"){
    
  cmip_data_plot <- cmip_data_all %>% 
  filter(spatial_scale %in% c("countries_admin"))
    
  }else if(spatial_type == "countries"){
    
    cmip_data_plot <- cmip_data_all %>% 
  filter(spatial_scale %in% c("countries"))
    
  }else if(spatial_type %in% c("countries_high_seas")){
  
      cmip_data_plot <- cmip_data_all %>% 
  filter(spatial_scale %in% c("countries", "FAO_area"))
}



cmip_data_plot <- cmip_data_plot %>% 
  mutate(mean_outlier = case_when(
    mean >= 69 ~ 69,
    mean <= -69 ~ -69, # limit to max positive or negative change
    TRUE ~ mean
  )) %>%
  mutate(new_index = abs(mean)) %>%
  group_by(figure_name, scenario) %>%
  mutate(stacked_mean_total = sum(new_index),
         arrange_mean = max(new_index)) %>%
  ungroup()
  
  
  max_value <- max(abs(cmip_data_plot$mean_outlier), na.rm = TRUE) # set a max value for plots
  mid_value = max_value/2

  max_value_label <- round(max_value)
  mid_value_label <- round(mid_value)

  
# for(decade_type in decades){ # specify the decade for the plot
  for(scen in scenarios){ # specify the scenario for the plot

# scen = "ssp585"

    # make a cleaned up version of scenario names for plotting
    scen_plot <- data.frame(scenario = scen) %>%
      mutate(scen_clean = case_when(
        scenario == "ssp126" ~ "SSP1-2.6",
        scenario == "ssp585" ~ "SSP5-8.5"
      )) %>%
      pull(scen_clean) %>% 
      unique() # pull a cleaned up version to label the plot
    
    if(scen_plot == "SSP5-8.5"){
      
      scen_plot_name = "High Emissions"
    }else{
      scen_plot_name = "Low Emissions"
    }
    
    
plot_df <- cmip_data_plot %>% ## filter for scenario
  filter(
         scenario == scen) %>%
  data.frame() %>%
  mutate(continent = ifelse(spatial_scale == "FAO_area", "High Seas", continent)) %>% # rename to continent high seas if it is high seas
  arrange(continent, -arrange_mean) # order by continent, increase/decrease, and mean 

get_pos_neg_ranks <- plot_df %>% 
    mutate(new_index = abs(mean_outlier)) %>%
  arrange(continent, -arrange_mean)

rank_pos_neg <- get_pos_neg_ranks %>%
  group_by(continent) %>% 
  summarise(sum_mean = sum(stacked_mean_total),
            count = length(stacked_mean_total)) %>%
  ungroup()

rank_pos_neg$continent <- factor(rank_pos_neg$continent, 
                                         levels=unique(rank_pos_neg$continent))
get_pos_neg_ranks$continent <- factor(get_pos_neg_ranks$continent, 
                                         levels=unique(rank_pos_neg$continent))

## change some names for plot so that they will fit... could probably move this to setup 


### CN not sure I fully understand this, but for Admin country names I would use the Country_FAO column in cmip_data_new and avoid adjusting too much. e.g. USA should be United States of America (maybe United States is OK but not US), Russia must be Russian Federation, and Netherlands should be Netherlands (Kingdom of the). 


rank_rgn <- get_pos_neg_ranks %>%
  arrange(continent, -arrange_mean) %>%
     mutate(rgn_name_short = figure_name)

# , 
#          rgn_name_short = gsub("United Kingdom of Great Britain and Northern Ireland", "United Kingdom", rgn_name_short),
#          rgn_name_short = gsub("Islands|islands", "Isl", rgn_name_short),
#          rgn_name_short = gsub("Island", "Isl", rgn_name_short),
#          rgn_name_short = gsub("Democratic", "Dem", rgn_name_short),
#          rgn_name_short = gsub("Republic", "Rep", rgn_name_short),
#          rgn_name_short = gsub("South", "S", rgn_name_short),
#          rgn_name_short = gsub("American", "Am", rgn_name_short),
#          rgn_name_short = gsub("the United States", "US", rgn_name_short),
#          rgn_name_short = gsub("Territory", "Terr", rgn_name_short),
#          rgn_name_short = gsub("Saint", "St", rgn_name_short),
#          rgn_name_short = gsub(" and ", " & ", rgn_name_short),
#          rgn_name_short = gsub("Republique", "Rep", rgn_name_short),
#          rgn_name_short = gsub("Dem Rep of the", "Dem Rep of", rgn_name_short),
#          rgn_name_short = gsub("Georgia and the", "Georgia and", rgn_name_short),
#          rgn_name_short = gsub("St Vincent and the", "St Vincent and", rgn_name_short),
#          rgn_name_short = gsub("Howland Isl & Baker Isl", "Howland & Baker Isl", rgn_name_short),
#          rgn_name_short = gsub("Northern", "N", rgn_name_short), 
#          rgn_name_short = gsub("Alaska", "AK", rgn_name_short),
#          rgn_name_short = gsub("Hawaii", "HI", rgn_name_short),
#          rgn_name_short = gsub("\\(the\\)", "", rgn_name_short),
#          rgn_name_short = gsub("Kingdom of the Netherlands", "Netherlands", rgn_name_short),
#          rgn_name_short = gsub("Russian Federation", "Russia", rgn_name_short),
#          rgn_name_short = gsub("Federated States", "Fed. States", rgn_name_short),
#          rgn_name_short = gsub("United Rep of Tanzania", "Tanzania", rgn_name_short)) %>%
#   mutate(rgn_name_short = ifelse(rgn_name_short == "Netherlands (Kingdom of the)", "Netherlands", rgn_name_short)) # not sure why gsub wasn't working


rank_rgn_label <- rank_rgn %>% 
    arrange(continent, -arrange_mean) %>%
  distinct(figure_name, rgn_name_short, arrange_mean) 

# add empty space to create space in the circleplot for axes
 empty_bar <- 5
 to_add <- data.frame(matrix(NA, empty_bar, ncol(rank_rgn_label)))
 colnames(to_add) = colnames(rank_rgn_label)
 to_add$figure_name <- as.character(1:empty_bar)
 rank_rgn_label  <- rbind(to_add, rank_rgn_label)
 
empty_bar <- 5
 to_add <- data.frame(matrix(NA, empty_bar, ncol(rank_rgn)) )
 colnames(to_add) = colnames(rank_rgn)
 to_add$figure_name <- as.character(1:empty_bar)
 rank_rgn_adj  <- rbind(to_add, rank_rgn)
 

rank_rgn_adj$end_point <- max_value # point where we will want to place the country labels.
rank_rgn_label$end_point <- max_value # point where we will want to place the country labels.

 # some code to orient the country labels
sequence_length = length(unique(rank_rgn_label$figure_name))
first_sequence = c(1:(sequence_length%/%2)) 
second_sequence = c((sequence_length%/%2+1):sequence_length) 
first_angles = c(90 - 180/length(first_sequence) * first_sequence)
second_angles = c(-90 - 180/length(second_sequence) * second_sequence)
rank_rgn_label$angle <- c(first_angles, second_angles)
rank_rgn_label$hjust <- c(rep(0, length(first_sequence)), 
                    rep(1, length(second_sequence)))


# color for region labels and breaks
rank_rgn_adj <- rank_rgn_adj %>%
  rename(index = mean) %>%
  mutate(figure_name = factor(figure_name, unique(figure_name))) %>%
  mutate(pos_neg = factor(pos_neg, unique(pos_neg))) %>%
  mutate(color = "black") %>%
  mutate(color = ifelse(is.na(index), "white", color)) %>%
  mutate(fill_category = case_when(
    index < -30 ~ "Decrease >30%",
    index <= -20 & index >= -30 ~ "Decrease 20 to 30%",
    index <= -10 & index > -20 ~ "Decrease 10 to 20%",
    index <= 0 & index > -10 ~ "Decrease <10%",
    index >= 0 & index < 10 ~ "Increase <10%",
    index >= 10 & index < 20 ~ "Increase 10 to 20%", 
    index >= 20 & index <= 30 ~ "Increase 20 to 30%",
    index > 30 ~ "Increase >30%"
  ))


# make dataframe for continent text locations. Also specify margins for the plots, as the country names don't always fit on the standard margins (5.5pt)
if(spatial_type == "countries_high_seas"){

  rgn_shift <- data.frame(
                         continent = rank_pos_neg$continent,
                         name_x = c(20, 75, 125, 165, 190, 215),
                         name_y = c(70, 70, 70, 70, 70, 70))
rgn_shift <- rgn_shift %>%
   mutate(continent = as.character(continent))

label_size = 5.5

plot_margin = margin(120, 120, 180, 70, unit = "pt")


}else if(spatial_type == "countries"){

    rgn_shift <- data.frame(
                         continent = rank_pos_neg$continent,
                         name_x = c(20, 75, 125, 165, 200),
                         name_y = c(70, 70, 70, 70, 70))
rgn_shift <- rgn_shift %>%
   mutate(continent = as.character(continent))
  
label_size = 5.5

plot_margin = margin(110, 120, 200, 50, unit = "pt")


}else if(spatial_type == "countries_admin"){
  
    rgn_shift <- data.frame(
                         continent = rank_pos_neg$continent,
                         name_x = c(20, 60, 90, 125, 145),
                         name_y = c(70, 70, 70, 70, 70))
rgn_shift <- rgn_shift %>%
   mutate(continent = as.character(continent))

label_size = 7

plot_margin = margin(160, 60, 180, 60, unit = "pt")   # need to make sure names fit on plot... top, right, bottom, left


}else if(spatial_type == "countries_admin_high_seas"){
  
  rgn_shift <- data.frame(
                         continent = rank_pos_neg$continent,
                         name_x = c(25, 55, 90, 125, 145, 165),
                         name_y = c(70, 70, 70, 70, 70, 70))
  
rgn_shift <- rgn_shift %>%
   mutate(continent = as.character(continent))

label_size = 7

plot_margin = margin(165, 105, 200, 90, unit = "pt")   # need to make sure names fit on plot... 
}


# specify theme
circle_theme <- theme(axis.line=element_blank(),
                      axis.text.y=element_blank(),
                      axis.ticks=element_blank(),
                      axis.title.x=element_blank(),
                      axis.title.y=element_blank(),
                      panel.background=element_blank(),
                      panel.border=element_blank(),
                      panel.grid.major=element_blank(),
                      panel.grid.minor=element_blank(),
                      plot.background=element_blank(),
                      axis.text.x = element_blank(),
                      plot.title = element_text(hjust = 1),
                     plot.caption = element_text(hjust = 1, size = 16),
                    plot.margin = plot_margin)


rank_rgn_adj$uncertain <- factor(rank_rgn_adj$uncertain, c("yes", "no")) # order the sd uncertainty factors

# Define colors for each fill category
fill_colors <- c(
  "Decrease >30%" = "#AC390E",
  "Increase <10%" = "honeydew3",
    "Decrease 20 to 30%" = "#C4603E",
  "Increase 10 to 20%" = "lightblue2",
      "Decrease 10 to 20%" = "darksalmon",
  "Increase 20 to 30%" = "#4297D3",
    "Decrease <10%" = "wheat",
  "Increase >30%" = "#1B194B"
)

# order categories
all_categories <- c("Decrease >30%",  "Decrease 20 to 30%", "Decrease 10 to 20%", "Decrease <10%", "Increase <10%", "Increase 10 to 20%", "Increase 20 to 30%",   "Increase >30%")
rank_rgn_adj$fill_category <- factor(rank_rgn_adj$fill_category, levels = all_categories)

all_categories_labels <- c("Decrease >30%",  "Decrease 20-30%", "Decrease 10-20%", "Decrease <10%", "Increase <10%", "Increase 10-20%", "Increase 20-30%",   "Increase >30%")


# we need to plot separate geoms for the decades because we are stacking the bar plots
data_2041 <- rank_rgn_adj %>%
  filter(decade %in% c("2041-2050", NA))

data_2091 <- rank_rgn_adj %>% 
  filter(decade %in% c("2091-2100", NA))


rank_rgn_label <- rank_rgn_label %>% 
  mutate(rgn_name_short = ifelse(arrange_mean >= 69, glue("{rgn_name_short}*"), rgn_name_short)) # denote if the country mean is technically above the max

 g <- ggplot() + # ok so this works but losing order of data in plots for some reason
  # geom_col(data = data_2091, aes(x = figure_name, y = new_index, fill = fill_category), position = "stack", show.legend = TRUE) + # if you dont want patterns, use this geom instead
      geom_col_pattern(position = "stack",
                    data = data_2091, aes(x = figure_name, y = new_index, fill = fill_category, pattern = uncertain),
    pattern_density = .5,
   pattern_spacing = 0.004,
   pattern_frequency = 5,
    pattern_fill = "white",
    pattern_size = 0.05,
    pattern_color = "white",
   show.legend = TRUE
    ) +
  #geom_col(data = data_2041, aes(x = figure_name, y = new_index, fill = fill_category), position = "stack", show.legend = FALSE) +
   geom_col_pattern(position = "stack",
                    data = data_2041, aes(x = figure_name, y = new_index, fill = fill_category, pattern = uncertain),
    pattern_density = .5,
   pattern_spacing = 0.004,
   pattern_frequency = 5,
    pattern_fill = "white",
    pattern_size = 0.05,
    pattern_color = "white",
   show.legend = FALSE
    ) +
   geom_errorbar(data = data_2041,
                 aes(x = figure_name,
                     ymax = new_index,
                     ymin = new_index), color = "white", size = 0.4) + 
  geom_errorbar(data = data_2091,
     aes(x = 1,
    ymin = -end_point/4,
    ymax = end_point,
    alpha = 0),
    color = "white"
  ) +
    geom_text(
      data = rank_rgn_label,
       aes(x = figure_name,
           label = rgn_name_short,
           angle = angle,
           y = end_point,
           hjust = hjust,
               ),
       inherit.aes = FALSE,
       # fontface = "bold",
       alpha = 0.9,
       size = label_size) +
   geom_segment(data = data_2091,
        x = 0,
         y = 0,
         xend = dim(data_2091)[1]+1,
         yend = 0,
        alpha = 1,
        color = "white",
        size = 0.1) +
    geom_segment(data = data_2091, 
        x = 0,
         y = 10,
         xend = dim(data_2091)[1]+1,
         yend = 10,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
    geom_segment(data = data_2091, 
        x = 0,
         y = 20,
         xend = dim(data_2091)[1]+1,
         yend = 20,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
                   geom_segment(data = data_2091, 
        x = 0,
         y = 30, # incase the scenario doesnt have any values > 30
         xend = dim(data_2091)[1]+1,
         yend = 30,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
           # georegion names
    geom_text(
      data = rgn_shift,
      aes(y = unique(rank_rgn_adj$end_point)/1.5,
          x = name_x,
          label = continent),
      inherit.aes=FALSE,
      size=label_size+1) +
    annotate("text",
        x     = c(3, 3, 3),
        y     = c(0, 30, unique(rank_rgn_adj$end_point)),
        label = c("0", glue("±30%"), glue("±{max_value_label}%")),
        color = "darkgray", angle = -4, size = label_size+1) +
    geom_text(data =data_2091 , x = 0, y = -unique(rank_rgn_adj$end_point)/4, label = glue("{scen_plot_name} \n{scen_plot}"), size = 7) +
  coord_polar(clip = "off") +
      scale_fill_manual(values = fill_colors, breaks = all_categories, labels = all_categories_labels, drop = F) +  # Assign fill colors manually
    labs(fill = "Change in fish biomass (%)") +
         scale_pattern_manual(values = c("circle", "none")) +
    guides(alpha = "none",
            pattern = "none",
            fill = "none") +
   circle_theme

if(spatial_type %in% c("countries", "countries_high_seas")){
  
  spatial_type = glue("{spatial_type}_eez")
  
}

  ggsave(plot = g, glue(here("outputs/circle_figs/{spatial_type}/bar_{scen}_{label_type}_stacked.png")), height=20, width=20, units=c("in"), bg = "white")

    }
  }
}




```

 
Make map of mean change and also save a legend 

```{r}

# Define colors for each fill category
fill_colors <- c(
  "Decrease >30%" = "#AC390E",
  "Increase <10%" = "honeydew3",
    "Decrease 20 to 30%" = "#C4603E",
  "Increase 10 to 20%" = "lightblue2",
      "Decrease 10 to 20%" = "darksalmon",
  "Increase 20 to 30%" = "#4297D3",
    "Decrease <10%" = "wheat",
  "Increase >30%" = "#1B194B"
)

# order categories

all_categories <- c("Decrease >30%",  "Decrease 20 to 30%", "Decrease 10 to 20%", "Decrease <10%", "Increase <10%", "Increase 10 to 20%", "Increase 20 to 30%",   "Increase >30%")

all_categories_labels <- c("Decrease >30%",  "Decrease 20-30%", "Decrease 10-20%", "Decrease <10%", "Increase <10%", "Increase 10-20%", "Increase 20-30%",   "Increase >30%")


globe_bbox <- rbind(c(-180, -90), c(-180, 90), 
                      c(180, 90), c(180, -90), c(-180, -90)) 

globe_border <- st_polygon(list(globe_bbox)) %>%
  st_sfc(crs = 4326) %>%
  st_sf(data.frame(rgn = 'globe', geom = .)) %>%
  smoothr::densify(max_distance = 0.5) %>%
  st_transform(crs =  '+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs')

world_360 <- # read_sf(file.path(base_folder, "data/world_360deg.shp")) %>% ## can't rerun this as this file was deleted....
  # read_sf(here("data/world_360deg.shp")) %>%
  st_transform('+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs')

 ## make map 
plot_df <- cmip_data_all %>%
  filter(spatial_scale %in% c("countries_admin")) %>% ## you can change this if you want to plot high seas or whatever
      mutate(fill_category = case_when(
    mean < -30 ~ "Decrease >30%",
    mean <= -20 & mean >= -30 ~ "Decrease 20 to 30%",
    mean <= -10 & mean > -20 ~ "Decrease 10 to 20%",
    mean <= 0 & mean > -10 ~ "Decrease <10%",
    mean >= 0 & mean < 10 ~ "Increase <10%",
    mean >= 10 & mean < 20 ~ "Increase 10 to 20%", 
    mean >= 20 & mean <= 30 ~ "Increase 20 to 30%",
    mean > 30 ~ "Increase >30%"
  )) %>% 
    mutate(scen_clean = 
           case_when(scenario == "ssp585" ~ "High Emissions: SSP5-8.5",
                     scenario == "ssp126" ~ "Low Emissions: SSP1-2.6"),
           decade_clean = case_when(
             decade == "2041-2050" ~ "Mid-century",
             decade == "2091-2100" ~ "End of century"
           )) 


eez_df <- world_360 %>%
 right_join(plot_df, by = c("adm0_a3" = "ISO3_country")) %>%
  rename(ISO3_country = iso_a3) %>% 
  dplyr::select(colnames(plot_df))


## Let's ditch many of the unnecessary elements
plain <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  panel.background = element_rect(fill = "white"),
  plot.title = element_text(hjust = 0.5),
  legend.position = "right",
  strip.text = element_text(size = 30),
  legend.text = element_text(size = 30),
  legend.key.height = unit(30, "mm"), # for vertical
  legend.key.width = unit(0.5, "cm"), # for vertical
  strip.background = element_rect(fill = "white"),
  legend.title = element_text(size = 30)
)

eez_df <- eez_df %>%
  mutate(mean = case_when(
    mean <= -69 ~ 69,
    mean >= 69 ~ 69, 
    TRUE ~ mean
  ))

eez_df$scen_clean <- factor(eez_df$scen_clean, levels = c("Low Emissions: SSP1-2.6", "High Emissions: SSP5-8.5"))
eez_df$decade_clean <- factor(eez_df$decade_clean, levels = c("Mid-century", "End of century"))


        p1 <- ggplot(data = eez_df) +
          geom_sf(aes(fill = fill_category), 
                  key_glyph = "rect", color = "white") + 
                scale_fill_manual(values = fill_colors, breaks = all_categories, labels = all_categories_labels, drop = F) +  # Assign fill colors manually
                 plain +
               labs(fill = "Change in fish biomass (%)")  +
                    coord_sf(crs =  '+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs') +
                           geom_sf(data = world_360, fill = NA, colour = "grey", alpha = 0.2)  +
                           geom_sf(data = globe_border, fill = NA, colour = "grey") +
          facet_grid(decade_clean~scen_clean, switch = "y") +
          guides(fill = guide_legend(byrow = TRUE,
                              title.position = "top",
               title.hjust = 1,
               title.vjust = 1))

        
        ggsave(plot = p1, glue(here("outputs/map_admin_countries.png")), bg = "white", height = 30, width = 30, units = "in", dpi = 300)
        ggsave(plot = p1, glue(here("outputs/map_admin_countries.pdf")), bg = "white", height = 30, width = 30, units = "in", dpi = 300)


 ## only saved this once

g_legend <- cowplot::get_plot_component(p1, "guide-box", return_all = TRUE)[[1]] # for vertical

ggsave(plot = g_legend, glue(here("outputs/circle_figs/legend_vertical.png")), height=20, width=20, units=c("in"), bg = "white") # for vertical
  
 


```



Code for saving non-stacked bar plots: 

```{r}

decades <- c("2091-2100", "2041-2050")
scenarios <- c("ssp585", "ssp126")
spatial_scales <- c("countries_high_seas", "countries_admin", "countries_admin_high_seas", "countries")


for(spatial_type in spatial_scales){ # specify what spatial scales you want to include in plot

  # spatial_type = "countries_admin"
  
  if(spatial_type == "countries_admin_high_seas"){
  
cmip_data_plot <- cmip_data_all %>% 
  filter(spatial_scale %in% c("countries_admin", "FAO_area"))

  }else if(spatial_type == "countries_admin"){
  cmip_data_plot <- cmip_data_all %>% 
  filter(spatial_scale %in% c("countries_admin"))
    
  }else if(spatial_type == "countries"){
    cmip_data_plot <- cmip_data_all %>% 
  filter(spatial_scale %in% c("countries"))
    
  }else if(spatial_type %in% c("countries_high_seas")){
  
      cmip_data_plot <- cmip_data_all %>% 
  filter(spatial_scale %in% c("countries", "FAO_area"))
}

  ## limit to maximum biomass loss (Kuwait -69%, this is the same regardless of spatial scale)

cmip_data_plot <- cmip_data_plot %>% 
  mutate(mean_outlier = case_when(
    mean >= 69 ~ 69,
    mean <= -69 ~ -69, 
    TRUE ~ mean
  ))
  

## grab max and midpoint values for the plot  
  max_value <- max(abs(cmip_data_plot$mean_outlier), na.rm = TRUE)
  mid_value = max_value/2

  max_value_label <- round(max_value)
  mid_value_label <- round(mid_value)

  
for(decade_type in decades){ # specify the decade for the plot
  for(scen in scenarios){ # specify the scenario for the plot

# scen = "ssp585"
# decade_type = "2091-2100"
    
    scen_plot <- data.frame(scenario = scen) %>%
      mutate(scen_clean = case_when(
        scenario == "ssp126" ~ "SSP1-2.6",
        scenario == "ssp585" ~ "SSP5-8.5"
      )) %>%
      pull(scen_clean) %>% 
      unique() # pull a cleaned up version to label the plot
    
    ## pull values for the global average mean and sd 
scen_decade_global_vals <- cmip_data %>%
  filter(spatial_scale %in% c("global")) %>%
  filter(decade == decade_type,
         scenario == scen) 

scen_decade_global_sd <- scen_decade_global_vals %>%
  pull(sd) %>% unique() %>% round
scen_decade_global_mean <- scen_decade_global_vals %>%
  pull(mean) %>% unique() %>% round()
    
plot_df <- cmip_data_plot %>% ## filter for scenario and decade
  filter(decade == decade_type,
         scenario == scen) %>%
  data.frame() %>%
  mutate(continent = ifelse(spatial_scale == "FAO_area", "High Seas", continent)) %>%
  arrange(continent, pos_neg, abs(mean_outlier)) # order by continent, increase/decrease, and mean

get_pos_neg_ranks <- plot_df %>%
  group_by(continent, pos_neg, figure_name) %>%
  summarise(index = sum(mean_outlier, na.rm = TRUE),
            sd = sum(sd, na.rm = TRUE)) %>% # these are just distinct numbers, so group by is unnecessary i think?
    mutate(new_index = abs(index)) %>%
  arrange(continent, pos_neg, -new_index)

rank_pos_neg <- get_pos_neg_ranks %>%
  group_by(continent) %>% 
  summarise(sum_mean = sum(new_index),
            count = length(new_index)) %>%
  ungroup()

rank_pos_neg$continent <- factor(rank_pos_neg$continent, 
                                         levels=unique(rank_pos_neg$continent))
get_pos_neg_ranks$continent <- factor(get_pos_neg_ranks$continent, 
                                         levels=unique(rank_pos_neg$continent))

## change some names for plot so that they will fit better...
rank_rgn <- get_pos_neg_ranks %>%
  arrange(continent, pos_neg, -new_index) %>%
     mutate(rgn_name_short = figure_name)

# , 
#          rgn_name_short = gsub("United Kingdom of Great Britain and Northern Ireland", "United Kingdom", rgn_name_short),
#          rgn_name_short = gsub("Islands|islands", "Isl", rgn_name_short),
#          rgn_name_short = gsub("Island", "Isl", rgn_name_short),
#          rgn_name_short = gsub("Democratic", "Dem", rgn_name_short),
#          rgn_name_short = gsub("Republic", "Rep", rgn_name_short),
#          rgn_name_short = gsub("South", "S", rgn_name_short),
#          rgn_name_short = gsub("American", "Am", rgn_name_short),
#          rgn_name_short = gsub("the United States", "US", rgn_name_short),
#          rgn_name_short = gsub("Territory", "Terr", rgn_name_short),
#          rgn_name_short = gsub("Saint", "St", rgn_name_short),
#          rgn_name_short = gsub(" and ", " & ", rgn_name_short),
#          rgn_name_short = gsub("Republique", "Rep", rgn_name_short),
#          rgn_name_short = gsub("Dem Rep of the", "Dem Rep of", rgn_name_short),
#          rgn_name_short = gsub("Georgia and the", "Georgia and", rgn_name_short),
#          rgn_name_short = gsub("St Vincent and the", "St Vincent and", rgn_name_short),
#          rgn_name_short = gsub("Howland Isl & Baker Isl", "Howland & Baker Isl", rgn_name_short),
#          rgn_name_short = gsub("Northern", "N", rgn_name_short), 
#          rgn_name_short = gsub("Alaska", "AK", rgn_name_short),
#          rgn_name_short = gsub("Hawaii", "HI", rgn_name_short),
#          rgn_name_short = gsub("\\(the\\)", "", rgn_name_short),
#          rgn_name_short = gsub("Kingdom of the Netherlands", "Netherlands", rgn_name_short),
#          rgn_name_short = gsub("Russian Federation", "Russia", rgn_name_short),
#          rgn_name_short = gsub("Federated States", "Fed. States", rgn_name_short),
#          rgn_name_short = gsub("United Rep of Tanzania", "Tanzania", rgn_name_short)) %>%
#   mutate(rgn_name_short = ifelse(rgn_name_short == "Netherlands (Kingdom of the)", "Netherlands", rgn_name_short)) # not sure why gsub wasn't working



# add empty space to create space in the circleplot for axes
 empty_bar <- 5
 to_add <- data.frame(matrix(NA, empty_bar, ncol(rank_rgn)) )
 colnames(to_add) = colnames(rank_rgn)
 to_add$figure_name <- as.character(1:empty_bar)
 rank_rgn_adj  <- rbind(to_add, rank_rgn)

rank_rgn_adj$end_point <- max_value # point where we will want to place the country labels.. for this we'll just use 100.

 # some code to orient the country labels
sequence_length = length(unique(rank_rgn_adj$figure_name))
first_sequence = c(1:(sequence_length%/%2)) 
second_sequence = c((sequence_length%/%2+1):sequence_length) 
first_angles = c(90 - 180/length(first_sequence) * first_sequence)
second_angles = c(-90 - 180/length(second_sequence) * second_sequence)
rank_rgn_adj$angle <- c(first_angles, second_angles)
rank_rgn_adj$hjust <- c(rep(0, length(first_sequence)), 
                    rep(1, length(second_sequence)))


# color for region labels and breaks
rank_rgn_adj <- rank_rgn_adj %>%
  mutate(figure_name = factor(figure_name, unique(figure_name))) %>%
  mutate(pos_neg = factor(pos_neg, unique(pos_neg))) %>%
  mutate(color = "black") %>%
  mutate(color = ifelse(is.na(index), "white", color)) %>%
  mutate(fill_category = case_when(
    index < -30 ~ "Decrease >30%",
    index <= -20 & index >= -30 ~ "Decrease 20 to 30%",
    index <= -10 & index > -20 ~ "Decrease 10 to 20%",
    index <= 0 & index > -10 ~ "Decrease <10%",
    index >= 0 & index < 10 ~ "Increase <10%",
    index >= 10 & index < 20 ~ "Increase 10 to 20%", 
    index >= 20 & index <= 30 ~ "Increase 20 to 30%",
    index > 30 ~ "Increase >30%"
  )) %>%
  mutate(uncertain = ifelse(sd > new_index, "yes", "no")) %>%
  mutate(uncertain = ifelse(is.na(uncertain), "no", uncertain))


# make dataframe for continent text locations

# make dataframe for continent text locations. Also specify margins for the plots, as the country names don't always fit on the standard margins (5.5pt)
if(spatial_type == "countries_high_seas"){

  rgn_shift <- data.frame(
                         continent = rank_pos_neg$continent,
                         name_x = c(20, 75, 125, 165, 190, 215),
                         name_y = c(70, 70, 70, 70, 70, 70))
rgn_shift <- rgn_shift %>%
   mutate(continent = as.character(continent))

label_size = 5.5

plot_margin = margin(120, 120, 180, 70, unit = "pt")


}else if(spatial_type == "countries"){

    rgn_shift <- data.frame(
                         continent = rank_pos_neg$continent,
                         name_x = c(20, 75, 125, 165, 200),
                         name_y = c(70, 70, 70, 70, 70))
rgn_shift <- rgn_shift %>%
   mutate(continent = as.character(continent))
  
label_size = 5.5

plot_margin = margin(110, 120, 200, 50, unit = "pt")


}else if(spatial_type == "countries_admin"){
  
    rgn_shift <- data.frame(
                         continent = rank_pos_neg$continent,
                         name_x = c(20, 60, 90, 125, 145),
                         name_y = c(70, 70, 70, 70, 70))
rgn_shift <- rgn_shift %>%
   mutate(continent = as.character(continent))

label_size = 7

plot_margin = margin(140, 10, 160, 10, unit = "pt")   # need to make sure names fit on plot... top, right, bottom, left


}else if(spatial_type == "countries_admin_high_seas"){
  
  rgn_shift <- data.frame(
                         continent = rank_pos_neg$continent,
                         name_x = c(25, 55, 90, 125, 145, 165),
                         name_y = c(70, 70, 70, 70, 70, 70))
  
rgn_shift <- rgn_shift %>%
   mutate(continent = as.character(continent))

label_size = 7

plot_margin = margin(165, 105, 200, 90, unit = "pt")   # need to make sure names fit on plot... 
}

# specify theme
circle_theme <- theme(axis.line=element_blank(),
                      axis.text.y=element_blank(),
                      axis.ticks=element_blank(),
                      axis.title.x=element_blank(),
                      axis.title.y=element_blank(),
                     legend.position="none", # do we want legend?  
                     # legend.position='top', 
                      #legend.justification = "center",
                      panel.background=element_blank(),
                      panel.border=element_blank(),
                      panel.grid.major=element_blank(),
                      panel.grid.minor=element_blank(),
                      plot.background=element_blank(),
                      axis.text.x = element_blank(),
                      legend.key.width = unit(26, "mm"),
                      legend.key.height = unit(0.3, "cm"),
                      plot.title = element_text(hjust = 1),
                     plot.caption = element_text(hjust = 1, size = 16),
                    plot.margin = plot_margin,
                      legend.spacing.y = unit(0, "pt"),
                    legend.text = element_text(size = 16))


rank_rgn_adj$uncertain <- factor(rank_rgn_adj$uncertain, c("yes", "no")) # order the sd uncertainty factors

# Define colors for each fill category
fill_colors <- c(
  "Decrease >30%" = "#AC390E",
  "Increase <10%" = "honeydew3",
    "Decrease 20 to 30%" = "#C4603E",
  "Increase 10 to 20%" = "lightblue2",
      "Decrease 10 to 20%" = "darksalmon",
  "Increase 20 to 30%" = "#4297D3",
    "Decrease <10%" = "wheat",
  "Increase >30%" = "#1B194B"
)

# order categories
all_categories <-  c("Decrease >30%", "Decrease 20 to 30%", "Decrease 10 to 20%", "Decrease <10%", "Increase <10%",  "Increase 10 to 20%", "Increase 20 to 30%", "Increase >30%")
rank_rgn_adj$fill_category <- factor(rank_rgn_adj$fill_category, levels = all_categories)

# this is probably redundant but i dont feel like fixing it
all_categories_new <- c("Decrease >30%", "Decrease 20-30%", "Decrease 10-20%", "Decrease <10%", "Increase <10%",  "Increase 10-20%", "Increase 20-30%", "Increase >30%")

rank_rgn_adj <- rank_rgn_adj %>% 
  mutate(rgn_name_short = ifelse(new_index == 69, glue("{rgn_name_short}*"), rgn_name_short)) %>%
  left_join(rgn_shift)

  
## categorical fill - pattern for uncertainty. Using ggpattern is REALLY finnicky, and it took me a really long time to get the pattern stuff right... so beware. 

  g <- ggplot(data = rank_rgn_adj, aes(x = figure_name, y = new_index, fill = fill_category,
                        pattern = uncertain
                        )) +
   geom_bar_pattern(stat = "identity",
    pattern_density = .5,
   pattern_spacing = 0.004,
   pattern_frequency = 5,
    pattern_fill = "white",
    pattern_size = 0.05,
    pattern_color = "white",
   show.legend = TRUE
    ) +
  geom_errorbar(
     aes(x = 1,
    ymin = -end_point/4,
    ymax = end_point,
    alpha = 0),
    color = "white"
  ) +
    geom_text(
      data = rank_rgn_adj,
       aes(x = figure_name,
           label = rgn_name_short,
           angle = angle,
           y = end_point,
           hjust = hjust,
               ),
       inherit.aes = FALSE,
       fontface = "bold",
       alpha = 0.9,
       size = label_size) +
   geom_segment(
        x = 0,
         y = 0,
         xend = dim(rank_rgn_adj)[1]+1,
         yend = 0,
        alpha = 1,
        color = "white",
        size = 0.1) +
           geom_segment(
        x = 0,
         y = 10,
         xend = dim(rank_rgn_adj)[1]+1,
         yend = 10,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
               geom_segment(
        x = 0,
         y = 20,
         xend = dim(rank_rgn_adj)[1]+1,
         yend = 20,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
                   geom_segment(
        x = 0,
         y = 30, # incase the scenario doesnt have any values > 30
         xend = dim(rank_rgn_adj)[1]+1,
         yend = 30,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
           # georegion names
    geom_text(
     data = rgn_shift,
      aes(y = 69/1.5,
          x = name_x,
          label = continent),
      inherit.aes=FALSE,
      size=label_size + 1 ) +
    annotate("text",
        x     = c(3, 3, 3),
        y     = c(0, 69/2, 69),
        label = c("0", "±50%", "±100%"),
        color = "darkgray", angle = -4, size = label_size + 1) +
      geom_text(x = 0, y = -69/4, label = glue("{scen_plot} \n{decade_type} \nGlobal Average \n{scen_decade_global_mean}% ±{scen_decade_global_sd}%"), size = 7) + ## place this in the middle of the plot
      coord_polar(clip = "off") +
      scale_pattern_manual(values = c("circle", "none")) +
      scale_fill_manual(values = fill_colors, breaks = all_categories, labels = all_categories_new, drop = F,
                        guide = guide_legend(override.aes = list(pattern = "none") # might be redundant (see below)
                                             )) +  # Assign fill colors manually
    circle_theme +
    labs(
      #caption = glue("{scen_plot}, {decade_type}"),
         fill = "Average biomass change") +
    guides(alpha = "none",
           pattern = "none",
           fill = guide_legend(
               title.position = "bottom",
               title.hjust = 0.5,
               title.vjust = 1,
               nrow = 1,
               label.position = "bottom",
               label.hjust = 0.5,
               label.vjust = 1,
               override.aes = list(pattern = "none")
    ))

if(spatial_type %in% c("countries", "countries_high_seas")){
  
  spatial_type = glue("{spatial_type}_eez")
  
}

  ggsave(glue(here("outputs/circle_figs/{spatial_type}/non_stacked/bar_{scen}_{decade_type}_pattern.png")), height=20, width=20, units=c("in"), bg = "white") # save as png


 g <- ggplot(data = rank_rgn_adj, aes(x = figure_name, y = new_index, fill = fill_category)) +
   geom_bar(stat = "identity", show.legend = TRUE) +
  geom_errorbar(
     aes(x = 1,
    ymin = -end_point/4,
    ymax = end_point,
    alpha = 0),
    color = "white"
  ) +
    geom_text(
     # data = rank_rgn_adj,
       aes(x = figure_name,
           label = rgn_name_short,
           angle = angle,
           y = end_point,
           hjust = hjust,
               ),
       inherit.aes = FALSE,
       fontface = "bold",
       alpha = 0.9,
       size = label_size) +
   geom_segment(
        x = 0,
         y = 0,
         xend = dim(rank_rgn_adj)[1]+1,
         yend = 0,
        alpha = 1,
        color = "white",
        size = 0.1) +
           geom_segment(
        x = 0,
         y = 10,
         xend = dim(rank_rgn_adj)[1]+1,
         yend = 10,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
               geom_segment(
        x = 0,
         y = 20,
         xend = dim(rank_rgn_adj)[1]+1,
         yend = 20,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
                   geom_segment(
        x = 0,
         y = 30,
         xend = dim(rank_rgn_adj)[1]+1,
         yend = 30,
        alpha = 1,
        color = "grey30",
        size = 0.1,
        linetype = "dashed") +
   
geom_text(
      # data = rgn_shift,
      aes(y = 69/1.5,
          x = name_x,
          label = continent),
      inherit.aes=FALSE,
      size=label_size + 1) +
    annotate("text",
        x     = c(3, 3, 3),
        y     = c(0, 69/2, 69),
        label = c("0", glue("±{mid_value_label}%"), glue("±{max_value_label}%")),
        color = "darkgray", angle = -4, size = label_size) +
           # georegion names
      geom_text(x = 0, y = -69/4, label = glue("{scen_plot} \n{decade_type} \nGlobal Average \n{scen_decade_global_mean}% ±{scen_decade_global_sd}%"), size = 7) + ## place this in the middle of the plot
  coord_polar(clip = "off") +
      scale_fill_manual(values = fill_colors, breaks = all_categories, labels = all_categories_new, drop = F) +  # Assign fill colors manually
  circle_theme +
    labs(
         fill = "Average biomass change") +
    guides(alpha = "none",

           fill = guide_legend(
               title.position = "bottom",
               title.hjust = 0.5,
               title.vjust = 1,
               nrow = 1,
               label.position = "bottom",
               label.hjust = 0.5,
               label.vjust = 1
    )
    )



  ggsave(glue(here("outputs/circle_figs/{spatial_type}/non_stacked/bar_{scen}_{decade_type}.png")), height=20, width=20, units=c("in"), bg = "white")

    }
  }
}

```

