---
title: "recreating_ipcc_plot"
author: "Julia Blanchard"
date: "2024-05-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load libraries 

```{r}
rm(list = ls())
library(reshape2)
library(ggplot2)
library(patchwork)
library(cowplot)
```

## Get data from IPCC figure S4 Summary for policy makers

Figure SPM.4 in IPCC, 2021: Summary for Policymakers. In: Climate Change 2021: The Physical Science Basis. Contribution of Working Group I to the Sixth Assessment Report of the Intergovernmental Panel on Climate Change [Masson-Delmotte, V., P. Zhai, A. Pirani, S.L. Connors, C. Péan, S. Berger, N. Caud, Y. Chen, L. Goldfarb, M.I. Gomis, M. Huang, K. Leitzell, E. Lonnoy, J.B.R. Matthews, T.K. Maycock, T. Waterfield, O. Yelekçi, R. Yu, and B. Zhou (eds.)]. Cambridge University Press, Cambridge, UK and New York, NY,USA, pp. 3−32, doi: 10.1017/9781009157896.001.]

```{r}
co2 <- read.csv("data/Carbon_dioxide_Gt_CO2_yr.csv") 
co2 <- melt(co2, id = c("years"))
names(co2) <- c("year", "scenario", "Co2")
```

## Define style and plot panel a

```{r}
base_gg <- list(geom_line(),
                #Adding no change line at 0 for reference 
                geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, 
                           linetype = 1),
                #Manually setting colours to be used in plots
                scale_color_manual(values = c("ssp119" = "#00ADCF",
                                              "ssp126" = "#173C66",
                                              "ssp245" = "#F79420",
                                              "ssp370" = "#E71D25",
                                              "ssp585" = "#951B1E"),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9", "SSP1-2.6", 
                                              "SSP2-4.5", "SSP3-7.0", 
                                              "SSP5-8.5")),
                scale_linetype_manual(values = c("ssp119" = 2, "ssp126" = 1,
                                                 "ssp245" = 2, "ssp370" = 2,
                                                 "ssp585" = 1),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9", "SSP1-2.6", 
                                              "SSP2-4.5", "SSP3-7.0", 
                                              "SSP5-8.5")),
                                scale_size_manual(values = c("ssp119" = 0.5,
                                                             "ssp126" = 1,
                                                             "ssp245" = 0.5,
                                                             "ssp370" = 0.5,
                                                             "ssp585" = 1),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9", "SSP1-2.6", 
                                              "SSP2-4.5", "SSP3-7.0", 
                                              "SSP5-8.5")),
                 guides(color = guide_legend(nrow = 1, title.position = "left")),
                theme_classic(),
                labs(y = bquote(CO[2]~"emissions Gt/year"), x = "Year"),
                 theme(legend.position = "none", 
                       panel.grid.minor.y = element_blank(), 
                       axis.title.x = element_text(size = 13),
                       axis.title.y = element_text(size = 13),
                       axis.text.y = element_text(size = 11),
                       axis.text.x = element_text(size = 11, angle = 45, 
                                                  vjust = 0.765, hjust = 0.65)))

a <- ggplot(data = co2, aes(x = year, y = Co2, color = scenario, 
                            linetype = scenario, size = scenario)) +
  base_gg +
  ggtitle("a") + 
  theme(plot.title = element_text(face = "bold", size = 13))
```

## Get data from IPCC figure S8 Summary for policy makers

Citation:

Figure SPM.8 in IPCC, 2021: Summary for Policymakers. In: Climate Change 2021: The Physical Science Basis. Contribution of Working Group I to the Sixth Assessment Report of the Intergovernmental Panel on Climate Change [Masson-Delmotte, V., P. Zhai, A. Pirani, S.L. Connors, C. Péan, S. Berger, N. Caud, Y. Chen, L. Goldfarb, M.I. Gomis, M. Huang, K. Leitzell, E. Lonnoy, J.B.R. Matthews, T.K. Maycock, T. Waterfield, O. Yelekçi, R. Yu, and B. Zhou (eds.)]. Cambridge University Press, Cambridge, UK and New York, NY,USA, pp. 3−32, doi: 10.1017/9781009157896.001 .]

```{r}
#historical Global mean surface air temperature (GSAT) anomalies relative to 
#1850-1900
hist <- read.csv("data/tas_global_Historical.csv") 
ssp119 <- read.csv("data/tas_global_SSP1_1_9.csv")
ssp126 <- read.csv("data/tas_global_SSP1_2_6.csv") 
ssp245 <- read.csv("data/tas_global_SSP2_4_5.csv")
ssp370 <- read.csv("data/tas_global_SSP3_7_0.csv") 
ssp585 <- read.csv("data/tas_global_SSP5_8_5.csv")

hist$scenario <-"hist"
ssp119$scenario <-"ssp119"
ssp126$scenario <-"ssp126"
ssp245$scenario <-"ssp245"
ssp370$scenario <-"ssp370"
ssp585$scenario <-"ssp585"

tas_global <- rbind(hist, ssp119, ssp126, ssp245, ssp370, ssp585)
# summary(tas_global)
tas_global$scenario <- as.factor(tas_global$scenario)
```

## Define style and plot panel b

```{r, echo=FALSE}
base_gg <- list(geom_line(),
                #Manually setting colours to be used in plots
                 scale_color_manual(values = c("hist" = "black",
                                               "ssp119" = "#00ADCF",
                                               "ssp126" = "#173C66",
                                               "ssp245"="#F79420",
                                               "ssp370"="#E71D25",
                                               "ssp585" = "#951B1E"
                                               ),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9", "SSP1-2.6", 
                                              "SSP2-4.5", "SSP3-7.0", 
                                              "SSP5-8.5")),
                scale_linetype_manual(values = c("hist" = 1, "ssp119" = 2,
                                                 "ssp126" = 1, "ssp245" = 2,
                                                 "ssp370" = 2, "ssp585" = 1),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9", "SSP1-2.6", 
                                              "SSP2-4.5", "SSP3-7.0", 
                                              "SSP5-8.5")),
                                scale_size_manual(values = c("hist" = 1,
                                                             "ssp119" = 0.5,
                                                             "ssp126" = 1,
                                                             "ssp245" = 0.5, 
                                                             "ssp370" = 0.5,
                                                             "ssp585" = 1),
                                                  name = "Scenarios",
                                                  labels = c("SSP1-1.9", 
                                                             "SSP1-2.6", 
                                                             "SSP2-4.5", 
                                                             "SSP3-7.0", 
                                                             "SSP5-8.5")),
                #Adding SD as shading, here 5% and 95% range
                geom_ribbon(aes(ymin = X5., ymax = X95., fill = scenario),
                             alpha = 0.1, color = NA),
                 scale_fill_manual(values = c("hist" = "black", 
                                              "ssp119" = "#00ADCF",
                                              "ssp126" = "#173C66",
                                              "ssp245"="#F79420",
                                              "ssp370"="#E71D25",
                                              "ssp585" = "#951B1E"),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9", "SSP1-2.6", 
                                              "SSP2-4.5", "SSP3-7.0", 
                                              "SSP5-8.5")),
                guides(color = guide_legend(nrow = 1, title.position = "left")),
                theme_classic(),
                labs(y = "Global surface temperature change", x = "Year"),
                theme(legend.position = "top", legend.justification = "center", 
                      legend.title = element_text(size = 12),
                      legend.text = element_text(size = 12),
                      panel.grid.minor.y = element_blank(), 
                      axis.title.x = element_text(size = 13),
                      axis.title.y = element_text(size = 13),
                      axis.text.y = element_text(size = 11),
                      axis.text.x = element_text(size = 11, angle = 45,
                                                 vjust = 0.765, hjust = 0.65)))

b <- ggplot(data = tas_global, aes(x = Year, y = Mean, color = scenario, 
                                 linetype = scenario, size = scenario))+
  base_gg + 
  scale_y_continuous(labels = ~ paste0(.x, "\u00B0"))+
  ggtitle("b")+ 
  theme(plot.title = element_text(face = "bold", size = 13))
```

## Combine panels into a single plot 

```{r}
# # problems with patchwork
# combined <- a + b +
#   plot_layout(guides = "collect") & theme(legend.position = 'top') 

# # work around with cowplot 

# extract legend from b
leg <- cowplot::get_plot_component(b, 'guide-box-top', return_all = TRUE)
# cowplot::ggdraw(leg) 

# remove legend from b
b <- b+
  theme(legend.position = "none")

# combine all in one figure
fig_scen <- plot_grid(plot_grid(a, b, ncol = 2, nrow = 1), 
                      plot_grid(NULL, leg, NULL, nrow = 1, 
                                rel_widths = c(0.5, 1, 0.5)),
                      ncol = 1, nrow = 2, rel_heights = c(1, 0.1))

# save file
# ggsave(fig_scen,filename="/Users/juliab6/Dropbox/FishMIP MAIN/FishMIP Repos/FAO_Report/figures/fig_scen.png")

pdf("figures/fig_scen.pdf", width = 9, height = 5)
fig_scen
dev.off()
```

# Cascading Uncertainty Plot

```{r}
# library(devtools)
# install_github(c("SantanderMetGroup/loadeR.java",
#                  "SantanderMetGroup/climate4R.UDG",
#                  "SantanderMetGroup/loadeR",
#                  "SantanderMetGroup/transformeR",
#                  "SantanderMetGroup/visualizeR",
#                  "SantanderMetGroup/downscaleR"))
# library(visualizeR)
# 
# load(url("http://meteo.unican.es/work/visualizeR/data/multimip_sdelta_es.rda"), verbose=T)
# cat(multimip.README)
# filter.members <- multimip.factors[["Method"]]=="GCM" & ! multimip.factors[["GCM"]] %in% c("CNRM-CM5-r8", "EC-EARTH-r3")
# sdeltas <- multimip.prtas.sdeltas["tas",c("GFDL-ESM2M-r1","IPSL-CM5A-MR-r1")]
# factors <- lapply(multimip.factors, function(x) factor(x[filter.members]))
# title <- "Delta T (K)"
# cascadePlot(sdeltas, factors, multimip.stages, title=title, density.offset=2)

```

