---
title: "recreating_ipcc_plot"
author: "Julia Blanchard"
date: "2024-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Get data from IPCC figure S4 Summary for policy makers

Figure SPM.4 in IPCC, 2021: Summary for Policymakers. In: Climate Change 2021: The Physical Science Basis. Contribution of Working Group I to the Sixth Assessment Report of the Intergovernmental Panel on Climate Change [Masson-Delmotte, V., P. Zhai, A. Pirani, S.L. Connors, C. Péan, S. Berger, N. Caud, Y. Chen, L. Goldfarb, M.I. Gomis, M. Huang, K. Leitzell, E. Lonnoy, J.B.R. Matthews, T.K. Maycock, T. Waterfield, O. Yelekçi, R. Yu, and B. Zhou (eds.)]. Cambridge University Press, Cambridge, UK and New York, NY,USA, pp. 3−32, doi: 10.1017/9781009157896.001 .]


```{r}

co2<-read.csv("../data/Carbon_dioxide_Gt_CO2_yr.csv") 


co2<-melt(co2,id=c("years"))
names(co2)<-c("year","scenario","Co2")

library(reshape2)


base_gg <- list(geom_line(),
                #Adding no change line at 0 for reference 
                geom_hline(yintercept = 0, color = "grey", linewidth = 0.5, 
                           linetype = 1),
                #Adding line dividing historical period and future projections
                #geom_vline(xintercept = 2015, color = "#709fcc", 
                #           linewidth = 0.65),
                #Manually setting colours to be used in plots
                 scale_color_manual(values = c(
                                               "ssp119" = "#00ADCF",
                                               "ssp126" = "#173C66",
                                               "ssp245"="#F79420",
                                               "ssp370"="#E71D25",
                                               "ssp585" = "#951B1E"
                                               ),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9","SSP1-2.6","SSP2-4.5","SSP3-7.0","SSP5-8.5")),
                scale_linetype_manual(values = c(
                                               "ssp119" = 2,
                                               "ssp126" = 1,
                                               "ssp245"=2,
                                               "ssp370"=2,
                                               "ssp585" = 1
                                               ),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9","SSP1-2.6","SSP2-4.5","SSP3-7.0","SSP5-8.5")),
                                scale_size_manual(values = c(
                                               "ssp119" = 0.5,
                                               "ssp126" = 1,
                                               "ssp245"=0.5,
                                               "ssp370"=0.5,
                                               "ssp585" = 1
                                               ),
                                   name = "Scenarios",
                                   labels = c("SSP1-1.9","SSP1-2.6","SSP2-4.5","SSP3-7.0","SSP5-8.5")),
                # # #Adding SD as shading, here 5% and 95% range
                 # geom_ribbon(aes(ymin = X5.,
                 #                ymax = X95., fill = scenario),
                 #             alpha = 0.1,color=NA),
                 # scale_fill_manual(values = c("hist" = "black",
                 #                               "ssp119" = "#00ADCF",
                 #                               "ssp126" = "#173C66",
                 #                               "ssp245"="#F79420",
                 #                               "ssp370"="#E71D25",
                 #                               "ssp585" = "#951B1E"
                 #                              
                 #                               ),
                 #                   name = "Scenarios",
                 #                  labels = c("Historical", "SSP1-1.9","SSP1-2.6","SSP2-4.5","SSP3-7.0","SSP5-8.5")),
                 guides(color = guide_legend(nrow = 7, title.position = "left")),
                theme_bw(),
                labs(y = bquote(CO[2]~"emissions Gt/year")),
                 theme(legend.position = "right", legend.justification = "center", 
                       legend.text = element_text(size = 8),
                       panel.grid.minor.y = element_blank(), 
                       axis.title.x = element_blank(),
                       axis.title.y = element_text(size = 10),
                       axis.text.x = element_text(angle = 45, vjust = 0.765, 
                                                  hjust = 0.65)))

 a<-ggplot(data = co2, aes(x = year, y = Co2, color = scenario,linetype=scenario,size=scenario))+
 base_gg  + theme(legend.position="none")



```




## Get data from IPCC figure S8 Summary for policy makers

Citation:

Figure SPM.8 in IPCC, 2021: Summary for Policymakers. In: Climate Change 2021: The Physical Science Basis. Contribution of Working Group I to the Sixth Assessment Report of the Intergovernmental Panel on Climate Change [Masson-Delmotte, V., P. Zhai, A. Pirani, S.L. Connors, C. Péan, S. Berger, N. Caud, Y. Chen, L. Goldfarb, M.I. Gomis, M. Huang, K. Leitzell, E. Lonnoy, J.B.R. Matthews, T.K. Maycock, T. Waterfield, O. Yelekçi, R. Yu, and B. Zhou (eds.)]. Cambridge University Press, Cambridge, UK and New York, NY,USA, pp. 3−32, doi: 10.1017/9781009157896.001 .]



```{r}
library(tidyverse)

#historical Global mean surface air temperature (GSAT) anomalies relative to 1850-1900

hist<-read.csv("../data/tas_global_Historical.csv") 
ssp119<-read.csv("../data/tas_global_SSP1_1_9.csv")
ssp126<-read.csv("../data/tas_global_SSP1_2_6.csv")
ssp245<-read.csv("../data/tas_global_SSP2_4_5.csv")
ssp370<-read.csv("../data/tas_global_SSP3_7_0.csv")
ssp585<-read.csv("../data/tas_global_SSP5_8_5.csv")

hist$scenario <-"hist"
ssp119$scenario <-"ssp119"
ssp126$scenario <-"ssp126"
ssp245$scenario <-"ssp245"
ssp370$scenario <-"ssp370"
ssp585$scenario <-"ssp585"

tas_global<-rbind(hist,ssp119,ssp126,ssp245,ssp370,ssp585)
summary(tas_global)
tas_global$scenario<-as.factor(tas_global$scenario)

```

## Plot trends

First get base ggplot style

```{r, echo=FALSE}
base_gg <- list(geom_line(),
                #Adding no change line at 0 for reference 
                #geom_hline(yintercept = 1.5, color = "#709fcc", linewidth = 0.65, 
                 #          linetype = 2),
                #Adding line dividing historical period and future projections
                #geom_vline(xintercept = 2015, color = "#709fcc", 
                #           linewidth = 0.65),
                #Manually setting colours to be used in plots
                 scale_color_manual(values = c("hist" = "black",
                                               "ssp119" = "#00ADCF",
                                               "ssp126" = "#173C66",
                                               "ssp245"="#F79420",
                                               "ssp370"="#E71D25",
                                               "ssp585" = "#951B1E"
                                               ),
                                   name = "Scenarios",
                                   labels = c("Historical","SSP1-1.9","SSP1-2.6","SSP2-4.5","SSP3-7.0","SSP5-8.5")),
                scale_linetype_manual(values = c("hist" = 1,
                                               "ssp119" = 2,
                                               "ssp126" = 1,
                                               "ssp245"=2,
                                               "ssp370"=2,
                                               "ssp585" = 1
                                               ),
                                   name = "Scenarios",
                                   labels = c("Historical","SSP1-1.9","SSP1-2.6","SSP2-4.5","SSP3-7.0","SSP5-8.5")),
                                scale_size_manual(values = c("hist" = 1,
                                               "ssp119" = 0.5,
                                               "ssp126" = 1,
                                               "ssp245"=0.5,
                                               "ssp370"=0.5,
                                               "ssp585" = 1
                                               ),
                                   name = "Scenarios",
                                   labels = c("Historical","SSP1-1.9","SSP1-2.6","SSP2-4.5","SSP3-7.0","SSP5-8.5")),
                # # #Adding SD as shading, here 5% and 95% range
                 geom_ribbon(aes(ymin = X5.,
                                ymax = X95., fill = scenario),
                             alpha = 0.1,color=NA),
                 scale_fill_manual(values = c("hist" = "black",
                                               "ssp119" = "#00ADCF",
                                               "ssp126" = "#173C66",
                                               "ssp245"="#F79420",
                                               "ssp370"="#E71D25",
                                               "ssp585" = "#951B1E"
                                              
                                               ),
                                   name = "Scenarios",
                                  labels = c("Historical", "SSP1-1.9","SSP1-2.6","SSP2-4.5","SSP3-7.0","SSP5-8.5")),
                 guides(color = guide_legend(nrow = 7, title.position = "left")),
                theme_bw(),
                labs(y = "Global surface temperature change"),
                 theme(legend.position = "bottom", legend.justification = "center", 
                       legend.text = element_text(size = 8),
                       panel.grid.minor.y = element_blank(), 
                       axis.title.x = element_blank(),
                       axis.title.y = element_text(size = 10),
                       axis.text.x = element_text(angle = 45, vjust = 0.765, 
                                                  hjust = 0.65)))

 b<-ggplot(data = tas_global, aes(x = Year, y = Mean, color = scenario,linetype=scenario,size=scenario))+
 base_gg + 
 scale_y_continuous(labels = ~ paste0(.x, "\u00B0"))
 
 

library(patchwork)
combined <- a + b 
fig_scen<-combined + plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

ggsave(fig_scen,filename="/Users/juliab6/Dropbox/FishMIP MAIN/FishMIP Repos/FAO_Report/figures/fig_scen.png")


```

# Cascading Uncertainty Plot


```{r}
library(devtools)
install_github(c("SantanderMetGroup/loadeR.java",
                 "SantanderMetGroup/climate4R.UDG",
                 "SantanderMetGroup/loadeR",
                 "SantanderMetGroup/transformeR",
                 "SantanderMetGroup/visualizeR",
                 "SantanderMetGroup/downscaleR"))
library(visualizeR)

load(url("http://meteo.unican.es/work/visualizeR/data/multimip_sdelta_es.rda"), verbose=T)
cat(multimip.README)
filter.members <- multimip.factors[["Method"]]=="GCM" & ! multimip.factors[["GCM"]] %in% c("CNRM-CM5-r8", "EC-EARTH-r3")
sdeltas <- multimip.prtas.sdeltas["tas",c("GFDL-ESM2M-r1","IPSL-CM5A-MR-r1")]
factors <- lapply(multimip.factors, function(x) factor(x[filter.members]))
title <- "Delta T (K)"
cascadePlot(sdeltas, factors, multimip.stages, title=title, density.offset=2)

```

